\documentclass[12pt, a4paper]{article}

% Packages
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{mathtools}
\usepackage{geometry}
\usepackage{setspace}
\usepackage{booktabs}
\usepackage{array}
\usepackage{multirow}
\usepackage{caption}
\usepackage{graphicx}
\usepackage{tikz}
\usetikzlibrary{arrows.meta, positioning, calc}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage{xcolor}
\usepackage{longtable}
\newcolumntype{P}[1]{>{\raggedright\arraybackslash}p{#1}}

\geometry{margin=2.5cm}
\onehalfspacing

% Custom commands
\newcommand{\Sk}{S_k}
\newcommand{\Fk}{F_k}

\title{\textbf{The Pandemic Trilemma} \\[1em] 
\large Work in Progress. Please do not cite or circulate.}

% Der Trick: \setcounter{footnote}{1} schiebt den Zähler unsichtbar vor, 
% sodass \thanks direkt als zweites Symbol das Kreuz wählt.
\author{Aulis Pesenti\setcounter{footnote}{1}\thanks{University of Basel, Wirtschaftswissenschaftliche Fakultät, Peter Merian-Weg 6, 4002 Basel, Switzerland; (e-mail: a.pesenti@unibas.ch).} \\ University of Basel}

\date{\today}

\begin{document}
\maketitle

\begin{abstract}

\end{abstract}
% =============================================================
% Section 1 — Introduction
% =============================================================

\section{Introduction}

%hie rnoch ohne Variablenbezeichnungen
%Wie den Social Planner einführen-> warum ist das so besser?
%Catchy Einstieg noch schreiben-> Thinking aboz z.B.
A government confronting a pandemic faces three simultaneous objectives: protecting public health by minimizing excess mortality, preserving economic stability by minimizing the deviation of output from its potential, and maintaining fiscal sustainability by not destabilizing the level of public debt. This paper establishes that these three objectives are mutually incompatible due to the structural architecture of the pandemic environment---forming a \textit{trilemma} in the precise economic sense of the Mundell-Fleming impossible trinity \citet{mundell1963, obstfeld2005}: no feasible policy of a benevolent social planner can simultaneously achieve all three goals, regardless of the resources deployed.

The impossibility arises from four properties of the pandemic environment that I formalize in Section~\ref{sec:trilemma} based on the economic theory and the existing literature. Viral prevalence of a deadly virus, once introduced exogenously, grows endogenously through human contact. As long as the virus is lethal e.g., there is no vaxine, the only instrument that protect the people is suppress transmission with instruments like---lockdowns, social distancing mandates, mobility restrictions---operate by reducing human interaction, which simultaneously contracts economic activity.



Viral prevalence of a deadly pathogen, once introduced exogenously, grows endogenously through human contact. There are only two means of containing the health threat: reducing the lethality of the virus through vaccination, or suppressing its transmission. In the absence of a vaccine---the condition that defined the critical phase of the pandemic from early 2020 through late 2021---suppression is the only available instrument. However, the available tools of suppression---lockdowns, social distancing mandates, mobility restrictions---reduce transmission by restricting human interactions that generate economic activity. Fiscal instruments can compensate for the resulting output loss, but they do so at the cost of public debt accumulation. And even in the absence of any discretionary fiscal response, the output contraction itself erodes the fiscal position through falling incomes, which reduce tax revenues and trigger mandatory transfer payments, generating debt independently of any policy choice.

Given this structure, the planner's decision is not whether to intervene fiscally but how to compose the intervention mainly choosing between two different fundamentally different transmission channels. Demand injection like for instance direct transfers, cash payments or expanded unemployment benefits sustains aggregate expenditure during containment, generating an immediate but transient output effect. Capacity preservation such as wage subsidies that maintain employment relationships, loan guarantees that prevent firm insolvency or equity injections that preserve sectoral viability reduces the persistence of the contraction by preventing the destruction of organizational capital, supplier networks, and job-specific human capital that would otherwise require costly reconstruction. These instruments carry different mechanisms, different fiscal costs, and different dynamic profiles. The optimal fiscal
response therefore depends not only on its aggregate size but on its composition---and this composition should vary over the course of the pandemic as the state of the economy evolves transforming the social planner's decision into an intertemporal dynamic optimization problem.

% --- Positioning against existing literature ---

A growing literature has formalized subsets of this problem. \citet{alvarez2020} solve for the optimal lockdown trajectory of a planner trading off fatalities against output costs, while treating the economy as a linear production function with no fiscal dimension. \citet{eichenbaum2021macro} endogenize the behavioral recession by embedding susceptible-infected-removed model (SIR) dynamics into a macroeconomic equilibrium, yet the government's only instrument is a Pigouvian containment tax---no fiscal compensation or debt dynamics enter the problem. \citet{KaplanMollViolante2020} come closest to the full trilemma: they integrate SIR dynamics into a heterogeneous-agent model with both lockdown and fiscal transfers, constructing a pandemic possibility frontier that maps policy combinations to the joint distribution of mortality and economic welfare losses. Their analysis establishes the binding health--economy trade-off and demonstrates that fiscal transfers reshape its distributional incidence. However, government debt enters as a residual accounting identity, not as an optimization target, and fiscal policy is modeled as undifferentiated lump-sum transfers rather than as a composition of instruments with distinct transmission mechanisms.\footnote{For a systematic comparison of modeling approaches in the pandemic-macro literature, see Table~\ref{tab:structured_comp} in Appendix~\ref{app:structured_comp}.}

Three gaps emerge from this body of work. First, no existing framework incorporates fiscal sustainability as a co-equal optimization target alongside health protection and output stabilization---the problem remains a dilemma rather than a trilemma. Second, the fiscal instruments available to the planner have not been disaggregated by their transmission mechanism which generate qualitatively different impact, dynamic profiles and carry different fiscal costs, losing the whole information about the optimal policy mix. Third, the instrument-specific structural parameters that govern the composition problem have not been estimated jointly under pandemic conditions using a transmission-mechanism-based disaggregation---existing cross-country evidence (Deb et al., 2021) benchmarks aggregate fiscal effectiveness but not compositional optimality. 

Closing these three gaps enables an analysis that no existing framework can deliver: given the containment trajectory that the prevailing epidemiological pressure necessitates, the planner can determine the optimal fiscal composition across demand injection and capacity preservation at each point in the pandemic, trace how this composition shifts as the state evolves, and quantify the output and debt consequences of deviating from it. This yields both a retrospective assessment of observed fiscal packages---identifying the excess debt attributable to suboptimal instrument choice---and a prospective benchmark for policymakers confronting future crises that share the pandemic's structural characteristics.

%Containment Intensity raus?
The research question is therefore: \textit{what is the optimal policy trajectory
across containment intensity, demand injection, and capacity preservation that jointly
minimizes excess mortality, output loss, and debt accumulation over the course of a
pandemic---and what are the costs, in foregone stabilization and excess debt, of
deviating from this trajectory?}

% --- Contributions ---

%Überarbeiten wenn Section fertig
To analyze this question, this paper makes three contributions. First, I develop a formal optimization framework---an iterative Linear-Quadratic
Regulator (iLQR) following \citet{bertsekas2017dynamic}---that solves the full trilemma: the
planner jointly optimizes containment intensity $S_k$ and the fiscal instruments
$F_k^{DI}$ and $F_k^{CP}$ across the multi-dimensional state space
$(y_k, d_k, b_k, \theta_k)$, capturing the bilinear interactions between containment
and economic persistence that define the pandemic trilemma. Existing approaches solve
simpler variants of this problem: Alvarez et al.\ (2021) apply Pontryagin's maximum
principle to a single-instrument continuous-time problem, and \citet{eichenbaum2021macro} solve a Ramsey problem over a scalar containment tax. The iLQR extends this to a multi-instrument setting while preserving analytical tractability: the backward Riccati recursion yields time-varying feedback gains $K_k$ that are directly interpretable as the planner's marginal policy response to state deviations, without requiring numerical value function iteration over a high-dimensional state space.\footnote{The bilinear term $\phi_S S_k \theta_k$ renders the full problem non-convex. The iterative linearization converges to a local optimum. In the
economically relevant region---where the planner responds monotonically to rising infections---local and global solutions coincide, as verified numerically in Appendix~\ref{app:convergence}.} The resulting state-dependent feedback law $u_k^* = -K_k x_k$ yields closed-form optimality conditions (Appendix~\ref{app:optimality}) that characterize the optimal policy trajectory---including the optimal intensity, timing, and composition of the fiscal response---as a function of the evolving state of the system. The quadratic objective penalizing $(y_k, d_k, b_k)$ can be interpreted as a second-order approximation to a welfare function in which the planner assigns explicit weights to output stabilization, mortality reduction, and fiscal sustainability---a Tinbergen-style formulation that takes the multi-objective nature of pandemic policy as given rather than deriving it from a single representative agent's preferences.\footnote{This abstracts from the distributional dimension emphasized by \citet{KaplanMollViolante2020}. The present framework addresses a logically prior question: what is the optimal \textit{aggregate} fiscal composition? The distributional incidence of a given composition is a complementary but distinct problem.}

Second, I construct a new daily micro-level database of 2,290 fiscal policy measures across
all 38 OECD member states during 2020--2021 \citep{pesenti2025}. Unlike existing
databases that aggregate fiscal responses at the package level, this dataset
disaggregates individual measures by instrument type, target group, and IMF fiscal
accounting category---the granularity required to operationalize the distinction
between demand injection and capacity preservation and thus to adress it in the analyisis. Each of the 49 policy instrument codes is mapped to the theoretical categories (DI or CP) based on a transparent classification protocol that identifies the primary transmission channel. The whole  classification is documented in detail in Section~\ref{sec:classification} or in \citep{Codebook}. 

%instruments that sustain aggregate expenditure through direct income replacement are classified as demand injection, while instruments that preserve productive capacity by maintaining employment relationships, firm solvency, or sectoral viability are classified as capacity preservation.Without this disaggregation, the composition question cannot be empirically addressed: aggregate fiscal impulse measures conflate instruments with fundamentally different transmission mechanisms and fiscal cost profiles.


%EMPIRICAL: DAILY DATA-> GANZER ABSATZ ÜBERARBEITEN
Third, I estimate the structural parameters of the transition system to solve for the optimal trajectory using panel Local Projections \citep{jorda2025} across 38 OECD economies during 2020--2021. 

%The identification strategy exploits cross-country variation in fiscal composition conditional on comparable containment trajectories. A two-stage approach first estimates the policy reaction function---fiscal composition as a function of containment stringency, pre-crisis fiscal position, and institutional characteristics---and then uses deviations from predicted composition as identifying variation.\footnote{The exclusion restriction requires that residual variation in fiscal composition affects economic outcomes only through the fiscal channel. Section~\ref{sec:identification} discusses potential threats, including the possibility that political economy variables driving composition choices (government ideology, coalition structure) independently affect expectations and hence outcomes.} The estimated parameters calibrate the theoretical model and enable two counterfactual exercises. The primary exercise takes each country's observed containment path as given and computes the conditionally optimal fiscal composition---quantifying the excess debt attributable to suboptimal instrument choice, holding the containment trajectory constant. This conditional design reflects the institutional reality that fiscal authorities responded to containment decisions rather than determining them, and isolates the margin over which fiscal policymakers had actual discretion. A secondary exercise computes the fully optimal trajectory across all three instruments, providing an upper bound on the welfare gains from joint optimization of containment and fiscal policy.

% --- Preview of results ---

The main findings are as follows. \textsc{[Placeholder -- to be completed after
estimation.]}

%\textsc{[Result 1: Estimated structural parameters.]} The demand injection multiplier $\hat{\alpha}_F^{DI}$ is estimated at \textsc{[XX]}, while the capacity preservation effectiveness $\hat{\eta}$ is \textsc{[XX]}. The fiscal cost of capacity preservation $\hat{\kappa}_F^{CP}$ is \textsc{[XX]} lower than that of demand injection $\hat{\kappa}_F^{DI}$, reflecting \textsc{[the lower realized cost of guarantees and conditional instruments / the partial recovery of loans and guarantees]}.

%\textsc{[Result 2: Optimal composition trajectory.]} The optimal fiscal mix shifts over the pandemic: during the initial containment phase (high $S_k$, rapidly rising $\theta_k$), the planner allocates \textsc{[XX\%]} of the fiscal envelope to capacity preservation, declining to \textsc{[XX\%]} as containment is relaxed and demand injection becomes the dominant channel. The \textsc{[crossover point / transition]} occurs at approximately \textsc{[quarter XX]}, when the output gap exceeds \textsc{[XX\%]}.

%\textsc{[Result 3: Counterfactual -- conditional optimum.]} Holding each country's observed containment path fixed, the conditionally optimal fiscal composition would have reduced cumulative excess debt by \textsc{[XX]} percentage points of GDP on average across the OECD, with the largest gains in \textsc{[countries with heavy DI tilt during early containment]}. The cross-country variance in excess debt attributable to composition is \textsc{[XX]} percentage points.

%\textsc{[Result 4: Counterfactual -- full optimum.]} Joint optimization over containment and fiscal policy yields additional welfare gains of \textsc{[XX]}, suggesting that the composition margin accounts for approximately \textsc{[XX\%]} of the total welfare distance between observed policy and the unconstrained optimum.

% --- Related literature ---
This paper builds up and contributes to three strands of literature. First, it extends the pandemic-macro modeling literature that integrates SIR dynamics with macroeconomic frameworks \citet{Atkeson2020, eichenbaum2021macro, alvarez2021, KaplanMollViolante2020, Jones2021, acemoglu2021} by incorporating fiscal sustainability as an explicit optimization objective and disaggregating fiscal instruments by transmission mechanism. Second, it contributes to the fiscal multiplier literature \citet{blanchard2002, ramey2018, romer2010} by estimating instrument-specific multipliers under pandemic conditions, where Keynesian supply shocks \citet{guerrieri2022} and voluntary demand reductions \citet{GOOLSBEE2021} render pre-crisis estimates unreliable. Third, it contributes to the empirical literature on cross-country fiscal responses to COVID-19 \citet{deb2021,chetty2020, Autor2022} by providing the first micro-level database with instrument-level granularity and source documentation across all OECD member states, enabling compositional analysis beyond aggregate fiscal impulse comparisons.

% --- Motivation / policy relevance ---

The question addressed here is not only of historical interest. The fiscal legacy of
the pandemic---average OECD government debt increased by approximately \textsc{[XX]}
percentage points of GDP between 2019 and 2021---directly shaped the fiscal space
available when subsequent shocks materialized. The inflation surge of 2022--2023
raised sovereign borrowing costs across advanced economies \citep{jorda2023}; the
energy crisis following the invasion of Ukraine required additional fiscal responses in
an environment of already elevated debt; and the prevailing macroeconomic uncertainty
continues to constrain fiscal policy options. Countries that achieved comparable
health and economic outcomes during the pandemic but at lower fiscal cost---through
more effective instrument composition---entered this sequence of subsequent shocks in
a structurally stronger position. Understanding which fiscal strategies achieved this
outcome, and why, is therefore directly relevant to the current fiscal policy
environment and to the management of future crises that share the pandemic's
structural characteristics.

% --- Roadmap ---

The paper proceeds as follows. Section~\ref{sec:trilemma} establishes the four
constitutive properties of the pandemic trilemma and its structural uniqueness relative
to other crisis types. Section~\ref{sec:model} develops the iLQR optimization
framework. Section~\ref{sec:data} describes the micro-level fiscal policy database.
Section~\ref{sec:estimation} presents the estimation strategy and structural parameter
estimates. Section~\ref{sec:results} reports the optimal composition trajectories and
counterfactual exercises. Section~\ref{sec:conclusion} concludes.




\section{The Pandemic Trilemma}
\label{sec:trilemma}

This section formalizes the trilemma by identifying four constitutive properties of the pandemic environment---exogenous shock with endogenous lethality, contractionary suppression instruments, heterogeneous fiscal compensation channels, and endogenous fiscal feedback---and establishing their circular interdependence. Each property is developed with reference to the empirical literature and introduces the variables and parameters that enter the optimization framework in Section~\ref{sec:model}, following the notation of \citet{bertsekas2017dynamic}. I then apply the four properties as a
diagnostic to alternative crisis types, establishing that the pandemic is the only configuration in which all four operate simultaneously within a tractable planning horizon.

%Hall, Jones und Klenow (2020) bringen ein Argument für die höhere Gewichtung von Lives

%% ============================================================
\subsection{Constitutive Properties}
\label{subsec:properties}
%% ============================================================


%% --- Property 1 ---
\subsubsection*{Property 1: Exogenous Shock with Endogenous Dynamics}


A pandemic begins with the introduction of a novel pathogen into the population. I denote the pandemic pressure at time $k$ by $\theta_k$, which I refer to as the viral prevalence. The initial outbreak constitutes an exogenous shock---it originates outside the economic system and is not the result of endogenous market dynamics or policy failure. I represent this shock by a stochastic disturbance $\varepsilon_k$.

The defining characteristic of a pandemic shock is its endogenous growth dynamic: in the absence of intervention, each infected individual transmits the virus to more than one susceptible person, causing the number of infections to grow over time. I capture this by a persistence parameter $\rho_\theta > 1$ (similar to the logic of the reproduction rate during the pandemic), implying that viral prevalence grows exponentially if left unchecked \citep{Atkeson2020, alvarez2021, eichenbaum2021macro}. It is the lethality of the pathogen that transforms this explosive dynamic into a policy imperative: a non-lethal pathogen with $\rho_\theta > 1$ would spread but cause no welfare loss, requiring no suppression and posing no trilemma. When the pathogen is lethal, unchecked exponential growth overwhelms the healthcare system and generates catastrophic mortality \citep{Jones2021, Barro2020}---establishing the necessity of active policy intervention at the cost of economic contraction (Property~2).

This property distinguishes the pandemic shock qualitatively from an impulse shock---such as an earthquake or a one-off financial disruption---where the disturbance dissipates naturally over time ($\rho < 1$) and the system returns to equilibrium without active intervention \citep{jorda2022longer}.

These potential consequences are captured by the state variable: \textit{excess mortality} ($d_k$), defined as the deviation of the periodic death rate from its historical seasonal average encompassing both direct COVID-19 deaths and indirect deaths attributable to avoidance of care and healthcare system overload \citep{Msemburi2023WHO, owid_em}. I denote the parameter governing the transmission from infections to deaths by $\delta_\theta$.

Crucially, the biological lag between infection and death implies that mortality in period $k+1$ is already determined by the current viral prevalence $\theta_k$ and thus predetermined from the planner's perspective. Consequently, the planner's intervention $S_k$---which suppresses transmission to $\theta_{k+1}$---only impacts mortality starting from period $k+2$. This relationship is given by:

\begin{equation}
    d_{k+2} = \delta_\theta \, \theta_{k+1}, \qquad \text{with} \quad \theta_{k+1} = \rho_\theta \theta_k+\epsilon_k
    \label{eq:mortality_intro}
\end{equation}

The interaction between the lethality parameter $\delta_\theta$ and the transmission potential $\rho_\theta$ plays a decisive role. If $\delta_\theta = 0$---that is, if the virus were non-lethal---or if the pathogen were not transmissible ($\rho_\theta = 0$), there would be no reason for the planner to suppress $\theta_{k+1}$. This logic is formally derived in Appendix \ref{app:Theta import}, which proves that the necessity of a trade-off-inducing policy ($S_k > 0$) depends strictly on the joint condition of positive lethality ($\delta_\theta > 0$) and active transmission ($\rho_\theta > 0$). Absent this conjunction, the suppression strategy becomes orthogonal to health outcomes, and no mortality costs or behavioral responses would arise.


%Fear Effect Extension \citep{GOOLSBEE2021}
%(the ``fear effect'', whereby households voluntarily reduce economic activity) would materialise. The fear effect is therefore not a function of prevalence \textit{per se} but of the perceived connection between infection and death. In a model extension incorporating behavioural channels, this response should accordingly be mediated through $\delta_\theta$: agents curtail activity \textit{because} the virus is lethal, not because it exists. In the baseline framework, however, we attribute the economic impact of the pandemic exclusively to the planner's policy response, maintaining the analytical focus on policy-driven trade-offs.

\medskip
\noindent\textit{Summary.} Property~1 introduces the epidemiological state: the infection rate $\theta_k$ (self-reinforcing with $\rho_\theta > 1$), the mortality outcome $d_k$ (driven by $\delta_\theta$), and the exogenous shock $\varepsilon_k$. Together, they establish that the planner must act to suppress viral prevalence, as the system is unstable without intervention.


%% --- Property 2 ---
\subsubsection*{Property 2: Suppression Instrument}
%Maybe NPI's noch unterscheiden nach der Art->bzw mein Index besser!
Given the necessity of intervention established by Property~1, the planner's primary tool to suppress viral transmission during a pandemic is \textit{non-pharmaceutical interventions} (NPIs): measures including, business closures, restrictions on gatherings, school closings, restrictions on internal movement as well as international travel controls and stay-at-home orders. I denote the intensity of these measures at time $k$ by $\Sk$, where $\Sk = 0$ represents the absence of restrictions and $\Sk = 1$ represents a complete lockdown in every aspect. 


NPIs suppress viral prevalence by reducing human contact \citep{Flaxman2020}. I denote the effectiveness of this suppression by the parameter $\phi_S > 0$. Crucially, NPIs do not remove a fixed number of infections from the population; rather, they reduce the rate at which existing infections generate new ones. Each unit of suppression $S_k$ reduces the transmission potential $\rho_\theta$ by the fraction $\phi_S$, yielding an effective reproduction rate $\rho_\theta^{\text{eff}}(S_k) = \rho_\theta(1 - \phi_S S_k)$. Combining this with the mortality transmission from Property~1, the full two-period channel from policy action to health outcome is:
%
\begin{equation}
    d_{k+2} = \delta_\theta \cdot \underbrace{\bigl[\rho_\theta (1 - \phi_S S_k) \, \theta_k + \varepsilon_k\bigr]}_{\theta_{k+1}} 
    \label{eq:excess dynamics}
\end{equation}
%
This equation encodes two features that are central to the planner's problem. First, the suppression effect is multiplicative in the current level of prevalence: at high $\theta_k$, each unit of lockdown prevents more infections---and therefore more deaths---than at low $\theta_k$. The marginal health benefit of suppression, $\partial d_{k+2} / \partial S_k = -\delta_\theta \rho_\theta \phi_S \theta_k$, is state-dependent and vanishes as the pandemic recedes (see Appendix \ref{app:Theta import}). Second, the planner does not control the infection level or the death rate directly but only the growth rate of the epidemic. When $\phi_S S_k$ is sufficient to push $\rho_\theta^{\text{eff}}$ below unity, infections decline; otherwise, they continue to grow despite the intervention. This threshold property---$\rho_\theta^{\text{eff}} < 1$ requires $S_k > \frac{1 - 1/\rho_\theta}{\phi_S}$---implies a minimum lockdown intensity below which suppression is ineffective at reversing the epidemic trajectory.

The central feature of Property~2 is that the channel through which NPIs suppress viral transmission---restricting human interaction---simultaneously reduces the volume of economic activity. Containment measures close businesses, idle labor, and curtail the flow of goods, services, and persons that constitutes aggregate economic output.This duality is not a matter of calibration or implementation; it is inherent to the nature of any instrument that operates through contact reduction.

%%Fear Kanal, dieses Argument noch stärken
%oolsbee and Syverson (2021) provide causal evidence that government-mandated lockdowns accounted for only a small fraction of the overall decline in economic activity—the majority was driven by voluntary behavioral changes in response to infection risk. This finding tightens the trilemma: even in the absence of formal suppression (Sk = 0), the epidemiological state θk itself depresses output through fear-driven demand reductions, implying that the health-economy trade-off is partly endogenous to the pandemic rather than entirely policy-induced.

To formalize this, I introduce the next state variable \textit{output gap} ($y_k$): the percentage deviation of real GDP from its potential trend, where $y_k = 0$ represents normal economic conditions and $y_k < 0$ represents a recession. The contractionary effect of NPIs on output is governed by the parameter $\alpha_S > 0$. Every unit of suppression $\Sk > 0$ therefore simultaneously implies some values of:
%
\begin{alignat}{2}
&\text{Epidemiological benefit:} &\qquad \Delta\theta &= -\phi_S \, \Sk < 0 \label{eq:S_health}\\[4pt]
&\text{Economic cost:} &\qquad \Delta y &= -\alpha_S \, \Sk < 0 \label{eq:S_econ}
\end{alignat}
%
The suppression instrument cannot control the epidemiological shock without simultaneously destabilizing the economic state. This dual impact is structural: it arises from the fact that $\alpha_S$ and $\phi_S$ are both strictly positive and operate through the same policy variable.

A natural objection to Property~2 is that alternative instruments, such as testing, might control the pandemic without suppressing economic activity, thereby dissolving the contractionary link. However, testing capacity constitutes a complement to, rather than a substitute for, NPIs \citep{repec:nbr}.
Critically, testing does not alter the biological parameters driving the pandemic: it reduces neither the inherent transmissibility ($\rho_\theta$) nor the lethality ($\delta_\theta$) of the virus. Instead, it operates exclusively through the suppression channel $S_k$. By identifying infected agents, testing enables targeted isolation---which remains a form of NPI. Ideally, this improves the efficiency of the intervention, increasing the ratio of epidemiological benefit ($\phi_S$) to economic cost ($\alpha_S$) by replacing broad lockdowns with specific quarantines. However, testing cannot eliminate the contractionary effect ($\alpha_S > 0$) since even targeted isolation disrupts economic transactions. Moreover, the net effect on the efficiency ratio is ambiguous. Higher detection rates mechanically raise measured case counts, which can intensify behavioral fear responses or trigger stricter policy reactions through political feedback channels. Paradoxically, this increased visibility can deteriorate the trade-off by increasing the effective economic cost $\alpha_S$ for a given level of true prevalence. Thus, testing modulates the parameter ratio but does not resolve the structurally contractionary nature of suppression.

Vaccination, by contrast, is the instrument that dissolves the trilemma by invalidating Property~2. It operates through both channels: reducing viral transmission among the immunized population (lowering $\rho_\theta$) and reducing the lethality of breakthrough infections (lowering $\delta_\theta$)---which, as demonstrated in Appendix~\ref{app:Theta import}, is the critical mechanism. Crucially, vaccination achieves both effects without operating through the suppression channel $S_k$---it requires no restriction of economic activity \citep{Giordano2021V}, thereby eliminating the contractionary link that constitutes Property~2. However, effective vaccines were not widely available until mid-2021, and the critical vaccination thresholds required to sustainably decouple mortality from infection rates were not reached in most OECD economies until the end of 2021 \citep{OECD2021}. Consequently, for the main phase of the pandemic (Q1 2020--Q4 2021)---the period during which critical policy decisions were taken \citep{Hale2021}---NPIs remained the only available suppression instrument, and were thus inherently contractionary.

(Plooot)

This observation has direct implications for the planning horizon. If vaccination significantly reduces $\delta_\theta$ and $\rho_\theta$ by the end of the planning period, the health crisis is resolved, effectively eliminating the trilemma at the terminal date. Thus, at the horizon, the planner deals exclusively with the fiscal legacy of the crisis, as detailed in the specification of the terminal condition in Section X.

\medskip
\noindent\textit{Summary.} Property~2 introduces the suppression instrument $\Sk$, its effectiveness $\phi_S$, its economic cost $\alpha_S$, and the output gap $y_k$. It establishes that pandemic suppression is inherently contractionary: the planner cannot protect health without damaging the economy unless there is an effective vaccine.


%A further dimension of heterogeneity arises from the targeting of fiscal measures to specific recipient groups---households versus firms---which may generate additional variation in both $\alpha_F$ and $\kappa_F$. However, since not all instrument--target combinations were implemented in practice (notably, credit guarantees were deployed exclusively for firms), this disaggregation is developed in the extended model of Section~X rather than treated here. 

%%NEU
\subsubsection*{Property 3: Fiscal Compensation Instruments}

Given that pandemic suppression is inherently contractionary (Property~2), the social planner requires compensating instruments to stabilize the economic state. Containment damages the economy in two ways: it directly reduces the level of output by $\alpha_S$ per unit of lockdown intensity, and it erodes the economy's capacity to recover by increasing the persistence of the output gap. A single fiscal instrument cannot address both dimensions: injecting demand compensates for lost output but does not restore the productive structures that the lockdown has damaged, while preserving productive structures sustains the economy's recovery capacity but does not offset the contemporaneous output loss. This motivates a disaggregation of fiscal policy into two instrument categories that differ in their transmission to aggregate output.

Turning to Demand Injection (DI), denoted by $F_k^{DI}$, this category comprises measures that directly increase the flow of disposable income in the economy like stimulus checks (such as the U.S.\ Economic Impact Payments under the \textit{CARES Act}), enhanced unemployment benefits (such as the Federal Pandemic Unemployment Compensation of \$600 per week), direct transfers to households (such as the Australian \textit{JobSeeker} supplement or the Japanese universal cash transfer of
¥100,000 per resident), and tax reductions that raise disposable income (such as the German temporary VAT reduction from 19\% to 16\% in H2 2020). These instruments transmit through the standard multiplier channel, shifting the output gap by $\alpha_F^{DI} > 0$ per unit deployed. The magnitude of this multiplier is the central parameter governing the effectiveness of fiscal
compensation---in a pandemic no less than in normal times. Pre-pandemic estimates place expenditure multipliers in a range of approximately 0.5 to 1.5 \citep{ramey2018, blanchard2002}, while revenue-based multipliers range from near zero to cumulative effects of 2 to 3 depending on the identification strategy \citep{blanchard2002, romer2010}.

However, applying pre-pandemic multiplier estimates to a pandemic environment is questionable. When containment measures restrict the supply of goods and services, demand-side stimulus cannot translate into expenditure at the usual rate \citep{chetty2020, coibion2020, fariecastro2021, bayer2023c}. The effectiveness of $F_k^{DI}$ during containment therefore depends not only on the magnitude and design of the measure but appears to be strongly state-dependent. This state-dependency motivates a second category of instruments---Capacity Preservation (CP), denoted by $F_k^{CP}$---to preserve the productive structure of the economy, as without such support the resulting demand shortfall can exceed the initial supply disruption as shown by \citet{guerrieri2022}.

\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% OUTPUT TRANSITION — Updated Interpretation (Option A: combined channel)
%%% Replaces: ouptu_transition_interpretation_old.txt

This category encompasses measures like short-time work schemes (such as the German \textit{Kurzarbeitergeld} or the UK Coronavirus Job Retention Scheme), payroll subsidies, and government-guaranteed loans (such as the French \textit{Pr\^{e}ts Garantis par l'\'{E}tat} or the German KfW-Sonderprogramm). These instruments preserve the existing productive structure---firm--worker matches, firm solvency, supply chain integrity---and thereby sustain the economy's capacity to recover when containment is lifted. To see how CP differs from DI, consider the role of $\rho_y$. In normal times, $\rho_y \in (0,1)$ captures the mean reversion of the output gap toward potential. During a pandemic, a prolonged lockdown threatens to alter this recovery dynamic: firms that become insolvent cannot resume production when restrictions lift, employment relationships that dissolve must be rebuilt through costly search and matching \citep{Fujita2024}, and supply chains that break require time and resources to reconstitute. These mechanisms increase the persistence of the output gap---they effectively raise $\rho_y^{\text{eff}}$, slowing convergence back to potential \citep{Kozlowski2020, Cerra2020}. In the limit, $\rho_y^{\text{eff}} \to 1$ implies hysteresis: permanent output loss \citep{BLANCHARD1987}.

Capacity preservation counteracts this mechanism. Economically, CP operates through two channels that are both conditional on active containment. First, by maintaining employment relationships and firm solvency, CP prevents the lockdown from converting a temporary contraction into permanent structural damage---the \textit{persistence channel}. Second, by keeping productive structures intact, CP reduces the contemporaneous output loss for a given level of containment, as firms with preserved capacity lose less output than those whose workforce has been disbanded---the \textit{cushioning channel}. Both channels share a defining property: they are operative only during active containment. At $S_k = 0$, neither the hysteresis risk nor the flow damage exists, and capacity preservation has no output effect. CP is therefore structurally distinct from DI: it modifies the economic impact of the lockdown rather than independently stimulating the economy.

I formalize the combined channel through the interaction of containment intensity and capacity preservation. The output transition is:
%
\begin{equation}
    y_{k+1} = \rho_y \, y_k
      + \underbrace{\psi \, S_k \, y_k}_{\substack{\text{lockdown-induced} \\ \text{hysteresis}}}
      + \underbrace{\tilde{\eta} \, S_k \, F_k^{CP}}_{\substack{\text{capacity} \\ \text{preservation}}}
      + \alpha_F^{CP} \, F_k^{CP}
      - \alpha_S \, S_k
      + \alpha_F^{DI} \, F_k^{DI}
      + \beta_\theta \, \theta_k
    \label{eq:output_full}
\end{equation}
%
where $\psi > 0$ governs the sensitivity of output persistence to lockdown intensity and $\tilde{\eta} < 0$ captures the combined capacity preservation effect---the extent to which CP reduces the economic cost of containment through both the persistence and cushioning channels.\footnote{The combined parameter $\tilde{\eta}$ subsumes two conceptually distinct mechanisms: (i) CP reduces the persistence of the output gap by preventing structural damage ($\partial \rho_y^{\text{eff}} / \partial F_k^{CP} < 0$), and (ii) CP reduces the contemporaneous contractionary impact of the lockdown ($\partial \alpha_S^{\text{eff}} / \partial F_k^{CP} < 0$). Appendix~Table~\ref{tab:channel_decomp} separates these channels empirically by including $F_k^{CP} \cdot y_{k-1}$ as a separate regressor, finding that the persistence channel ($-0.024$, $p = 0.07$) contributes alongside the cushioning channel ($-0.008$, $p < 0.01$), with the combined effect well captured by $\tilde{\eta}$.} The interaction structure $\tilde{\eta} \, S_k \, F_k^{CP}$ encodes the key theoretical prediction: CP is effective precisely during lockdowns and neutral otherwise.

The marginal output effect of CP is therefore state-dependent:
%
\begin{equation}
    \frac{\partial y_{k+1}}{\partial F_k^{CP}} = \alpha_F^{CP} + \tilde{\eta} \, S_k
    \label{eq:dy_dCP}
\end{equation}
%
At low stringency ($S_k \approx 0$), the marginal effect equals the level term $\alpha_F^{CP}$; at the critical threshold $S^* = -\alpha_F^{CP} / \tilde{\eta}$, the marginal effect reaches zero; above $S^*$, additional CP generates diminishing returns. The empirical estimate yields $S^* = 0.260 / 0.006 = 42$, which coincides with the sample mean stringency across OECD economies during the pandemic---implying that the average country operated precisely at the boundary of CP effectiveness.\footnote{This threshold has a clean policy interpretation: CP is highly effective in quarters where containment exceeds $S^* = 42$ (the acute lockdown phase) and adds no output benefit in quarters where containment is mild. The subsample analysis in Section~\ref{subsec:robustness} confirms this prediction: CP is significant only in high-stringency countries and completely insignificant in low-stringency ones.}

By contrast, the marginal output effect of demand injection is constant:
%
\begin{equation}
    \frac{\partial y_{k+1}}{\partial F_k^{DI}} = \alpha_F^{DI} > 0
    \label{eq:dy_dDI}
\end{equation}
%
DI shifts the output level by $\alpha_F^{DI}$ per unit deployed, regardless of the containment environment. The empirical lag structure (significant at $k-2$, insignificant at $k-1$) reflects the implementation chain of transfer programs: authorization, disbursement, and spending each require time, placing the output effect two quarters after deployment.

The marginal output cost of containment depends on the depth of the recession and the level of capacity preservation:
%
\begin{equation}
    \frac{\partial y_{k+1}}{\partial S_k} = \psi \, y_k - \alpha_S + \tilde{\eta} \, F_k^{CP}
    \label{eq:dy_dS}
\end{equation}
%
Each additional unit of lockdown is more costly when the economy is already weakened ($\psi |y_k|$ amplifies the contractionary impact) and less costly when accompanied by capacity preservation ($\tilde{\eta} \, F_k^{CP}$ offsets part of the damage). This convexity provides a formal rationale for early intervention \citep{acemoglu2021}: a lockdown imposed when $|y_k|$ is small incurs lower marginal output costs than the same lockdown after a prolonged contraction.

These properties establish three results that are central to the planner's problem. First, demand injection and capacity preservation are complements, not substitutes: DI compensates for the output loss that CP cannot prevent, while CP preserves the productive structures that make DI effective. Second, the optimal fiscal mix is state-dependent: during strict containment ($S_k$ high), CP dominates because its marginal effect scales with $S_k$; as containment eases, DI becomes relatively more valuable as the demand channel operates more effectively in a less supply-constrained environment. Third, CP is a crisis-specific instrument: at $S_k = 0$, CP has no output effect beyond its level term $\alpha_F^{CP}$, which the empirical evidence shows is approximately offset by the interaction at the mean stringency---yielding a net effect near zero outside of active containment.

\citet{deb2021} document a pattern consistent with these predictions using high-frequency identification across 52 countries: liquidity-based measures outperform income-support measures during periods of strict containment, while the ranking reverses upon reopening. The present framework provides the structural rationale for this finding: the state-dependent interaction $\tilde{\eta} \, S_k \, F_k^{CP}$ is the mechanism through which containment-conditional fiscal effectiveness arises.

When $\rho_y + \psi S_k \geq 1$, the effective persistence of the output gap exceeds unity and the economy cannot recover autonomously---the contraction becomes self-reinforcing. Capacity preservation counteracts this by reducing the total output damage through $\tilde{\eta} S_k F_k^{CP}$, preventing the gap from reaching the explosive threshold. This provides a formal rationale for the universal combination of containment measures with capacity preservation programs observed during the pandemic. The empirical counterpart is the fact that corporate insolvencies across OECD economies fell by 17\% on average in 2020 relative to 2019 despite the deepest recession since the Great Depression \citep{djankov2021, banerjee2021}.

The preceding analysis might suggest a straightforward resolution: deploy CP to offset the lockdown damage (until $|\tilde{\eta}| S_k F_k^{CP} \approx \alpha_S S_k + \psi S_k |y_k|$), then inject demand into an economy whose productive structures are intact---as also suggested by \citet{deb2021} in a similar way. However, this reasoning abstracts from the budgetary consequences of the fiscal package itself. Each instrument that stabilizes the output gap simultaneously augments the government's balance sheet---and the magnitude of this fiscal cost is heterogeneous across categories and in several dimensions not yet established in the literature. To formalize this constraint, I introduce the third state variable: \textit{public debt} ($b_k$), defined as the deviation of the debt-to-GDP ratio from its pre-pandemic level, where $b_k = 0$ represents the pre-crisis fiscal position.\footnote{I normalize by pre-pandemic GDP (2019) rather than current GDP. When debt is expressed relative to current output, a recession mechanically inflates the ratio through the denominator even absent any new borrowing. Normalizing by the pre-pandemic level isolates changes in the numerator---the actual stock of government liabilities---which is the object relevant to the planner's decision.}

The fiscal cost of each instrument category is governed by the parameter $\kappa_F^j > 0$, which maps one unit of deployed fiscal support into its budgetary impact. 

For demand injection, a first approximation suggests $\kappa_F^{DI} \approx 1$: one unit of transfer or spending generates one unit of new government liability. Several mechanisms drive a wedge around this benchmark in both directions. On the one hand, the net fiscal cost may fall below unity when transfers are subject to income taxation---as the government recovers a fraction at the recipient's marginal tax rate---or when emergency programs partially substitute
for already-budgeted outlays. On the other hand, the effective cost may exceed unity due to administrative overhead and targeting failures: programs deployed under time pressure generate fiscal costs without commensurate stabilization when resources reach non-intended recipients \citep{autor2022}. The net direction of these opposing forces under pandemic conditions remains an open question.

For capacity preservation, the fiscal cost $\kappa_F^{CP}$ is not a single parameter but a weighted average over two structurally distinct sub-categories. Direct capacity preservation---short-time work schemes, payroll subsidies, and direct grants to firms---generates immediate budgetary outlays classified as above-the-line expenditure by the IMF \citep{imf2022fiscalmonitor}. Loans and guarantees, defined by the IMF as below-the-line- by contrast, involve either a net fiscal cost well below unity (loans, where the government creates a financial asset offsetting the cash outflow) or near zero at deployment (guarantees, which generate no immediate outflow). For both, the realized fiscal cost is governed by the default rate among supported borrowers---a quantity that is not known ex ante and is itself endogenous to the crisis.

The effective $\kappa_F^{CP}$ therefore depends on two factors that varied substantially across OECD economies (Section~\ref{sec:data}): the composition of the CP package---the share of direct measures versus loans and guarantees---and the realized default rate on contingent instruments. Estimating the effective $\kappa_F^{CP}$ under pandemic conditions---and its dependence on containment trajectories and package composition---is therefore itself an open question.


Health system expenditures driven by the pandemic---intensive care, testing infrastructure, vaccination logistics---are not a policy instrument in the planner's fiscal composition problem but an endogenous cost driven by the epidemiological state. I specify this cost as $c_k^H = c_H\,\theta_k$, where $c_H > 0$. Unlike routine health expenditures, which vary with the business cycle and are absorbed by the automatic stabilizer $\gamma_y$, this term captures exclusively the pandemic-specific costs that arise because $\theta_k > 0$, independently of the output gap.\footnote{The parameter $c_H$ is calibrated from health-related fiscal measures in
the database (policy codes 30--34, Section~\ref{sec:data}). Health expenditures enter the empirical specification as a control variable rather than an outcome of interest.} Containment therefore has an offsetting fiscal benefit: by reducing future prevalence, suppression lowers health expenditures ($S_k \to \theta_{k+1} \downarrow \to c_H\, \theta_{k+1} \downarrow$), partially counteracting the fiscal cost it imposes through the output contraction channel.

The debt transition therefore is:
%
\begin{equation}
    b_{k+1} = (1+r) \, b_k \;-\; \gamma_y \, y_k
      \;+\; \kappa_F^{DI} \, F_k^{DI}
      \;+\; \kappa_F^{CP} \, F_k^{CP}
      \;+\; c_H \, \theta_k
    \label{eq:debt_transition}
\end{equation}
%
The term $(1+r) b_k$ reflects the intertemporal budget constraint: outstanding debt compounds at the real interest rate $r$, rendering the debt state unconditionally explosive for any positive real interest rate---no fiscal policy instrument can alter this dynamic directly. The term $-\gamma_y y_k$ captures the automatic stabilizer channel formalized in Property~4: a negative output gap endogenously widens the deficit without intervention\footnote{Notably, the containment instrument $S_k$ does not appear in (\ref{eq:debt_transition}). The fiscal cost of a lockdown is an opportunity cost---the foregone economic activity whose budgetary consequences materialize indirectly through $S_k \to y_{k+1} \to -\gamma_y y_{k+1} \to b_{k+2}$, with a one-period lag relative to the output effect. This distinction matters: the planner incurs no direct fiscal outlay from imposing restrictions, but the resulting output contraction erodes the budget with certainty in the subsequent period---a lag structure that underscores the dynamic nature of the planner's problem.}.

The price of stabilization is captured in the additive terms $\kappa_F^{CP} F_k^{CP}$ and $\kappa_F^{DI} F_k^{DI}$. The joint structure of equations (\ref{eq:output_full}) and (\ref{eq:debt_transition}) reveals why the three objectives are mutually incompatible---and why the debt cost of stabilization cannot be read from (\ref{eq:debt_transition}) alone. The fiscal terms $\kappa_F^{DI} F_k^{DI}$ and $\kappa_F^{CP} F_k^{CP}$ capture the gross budgetary cost at deployment; the net cost emerges only through the interaction of both equations across periods, as fiscal support that narrows the output gap reduces automatic debt accumulation in subsequent periods. To see this, consider the fiscal cost of fully closing the output gap ($y_{k+1}=0$) through DI alone, setting $F_k^{CP} = 0$. Solving (\ref{eq:output_full}) for $F_k^{DI}$:
%
\begin{equation}
    F_k^{DI} = \frac{\alpha_S \, S_k - \rho_y \, y_k - \psi \, S_k \, y_k}{\alpha_F^{DI}}
    \label{eq:DI_required}
\end{equation}
%
Substituting into the debt transition yields the gross fiscal cost of full output stabilization:
%
\begin{equation}
    \Delta b = \frac{\kappa_F^{DI}}{\alpha_F^{DI}}
      \left(\alpha_S \, S_k - \rho_y \, y_k - \psi \, S_k \, y_k\right)
    \label{eq:cost_stabilization}
\end{equation}
%
The ratio $\kappa_F^{DI} / \alpha_F^{DI}$ is the price of output stabilization in debt units: the budgetary cost per unit of output gap closed. Lockdown intensity $S_k$ drives the required DI volume through two components: the direct contractionary effect $\alpha_S S_k$, which creates the output gap, and the hysteresis amplification $\psi S_k |y_k|$, which slows the economy's autonomous recovery. When CP is deployed alongside DI, it directly reduces the output damage through the interaction $\tilde{\eta} S_k F_k^{CP}$, lowering the volume of DI required to close the gap. Each unit of CP offsets $|\tilde{\eta}| S_k / \alpha_F^{DI}$ units of DI, at a fiscal cost of $\kappa_F^{CP}$ per unit of CP.

The planner therefore faces a composition problem with two margins. On the \textit{price margin}, the ratio $\kappa_F^{DI} / \alpha_F^{DI}$ determines how costly each unit of demand injection is in debt terms; a higher multiplier $\alpha_F^{DI}$ or a lower fiscal cost $\kappa_F^{DI}$ reduces this price. On the \textit{volume margin}, capacity preservation reduces the total amount of DI required by directly offsetting part of the lockdown's output damage through the state-dependent interaction $\tilde{\eta} \, S_k \, F_k^{CP}$, effectively substituting for DI at its own fiscal cost $\kappa_F^{CP}$. The optimal mix balances these two margins: the planner deploys CP up to the point where its marginal fiscal cost $\kappa_F^{CP}$ equals the DI savings it generates, valued at the price $\kappa_F^{DI} / \alpha_F^{DI}$. This trade-off is itself state-dependent, as the volume of DI that CP displaces scales with the intensity of containment ($S_k$).

This gross cost overstates the true fiscal burden, because the output improvement generated by fiscal support reduces automatic debt accumulation in subsequent periods through $\gamma_y$. The two-period debt derivatives (Appendix~X) formalize this mechanism and reveal a fundamental asymmetry: the net fiscal cost of DI is state-independent and strictly positive under pandemic conditions, whereas the net fiscal cost of CP is state-dependent---declining with the intensity of containment and becoming self-financing when $S_k$ is sufficiently large, because the output damage that CP prevents would otherwise generate debt through the automatic stabilizer channel. This asymmetry, combined with the volume effect established above---CP reduces the amount of DI required by offsetting part of the lockdown's output damage---provides a formal basis for the prevalence of mixed fiscal strategies observed across OECD economies during the pandemic (Section~\ref{sec:data}). However, the optimal composition cannot be determined by a static rule, as the relative cost advantage of each instrument shifts with the evolving containment trajectory, requiring dynamic optimization across the full planning horizon.

\noindent\textit{Summary.} Property~3 introduces the fiscal compensation instruments, disaggregated into demand injection ($F_k^{DI}$) and capacity preservation ($F_k^{CP}$), the output gap $y_k$, and the public debt state $b_k$. On the output side, the two instrument categories operate through complementary channels: DI shifts the output level through a state-independent multiplier $\alpha_F^{DI}$, while CP reduces the output cost of containment through the state-dependent interaction $\tilde{\eta} \, S_k \, F_k^{CP}$, which is operative only during active lockdowns. On the debt side, DI generates higher but more predictable fiscal costs, while CP carries lower expected costs but greater balance sheet uncertainty---with a net fiscal effect that is state-dependent and potentially self-financing when containment is strict. The optimal fiscal composition depends jointly on the intensity of containment, the structural parameters governing transmission, and the fiscal cost differential between categories.


\subsubsection*{Property 4: Endogenous Fiscal Feedback via Automatic Stabilizers}

When NPIs contract economic output (Property~2), the resulting negative output gap triggers two simultaneous fiscal effects. On the revenue side, the tax base shrinks reducing government receipts below their pre-pandemic trajectory. On the expenditure side, mandatory transfer programs such as unemployment insurance, social assistance, housing benefits expand as more individuals satisfy eligibility criteria. The parameter $\gamma_y > 0$ in the debt transition (\ref{eq:debt_transition}) captures the aggregate sensitivity of the budget to the output gap, encompassing both channels in a single semi-elasticity.\footnote{For OECD economies, estimates of the budget semi-elasticity range from approximately 0.35 to 0.65, with a cross-country average around 0.5 \citep{repec:oecd}. The parameter is structurally determined by the tax-to-GDP ratio and the coverage of automatic transfer programs.}

This term operates symmetrically. A negative output gap ($y_k < 0$) widens the deficit through lower revenues and higher mandatory expenditures; a narrowing gap ($y_k \to 0$) reverses both effects, reducing the pace of debt accumulation. The symmetry carries a direct implication for instrument choice (Property~3): a fiscal instrument with a high output multiplier $\alpha_F^j$ relative to its fiscal cost $\kappa_F^j$ generates a double dividend---it stabilizes output directly and
attenuates the automatic fiscal deterioration indirectly, as the improved $y_k$ feeds back through $\gamma_y$ to slow debt accumulation. Which instrument category achieves the more favorable ratio depends on the state of the economy, as established in Property~3.

The central contribution of this property is to close the trilemma by establishing that fiscal abstinence is not a feasible strategy \citep{Fornaro2020}. Setting all discretionary instruments to zero ($F_k^{DI} = F_k^{CP} = 0$) in (\ref{eq:debt_transition}) yields:

%
\begin{equation}
    b_{k+1} = (1+r) \, b_k - \gamma_y \, y_k
    \label{eq:debt_no_fiscal}
\end{equation}
%
Since $y_k < 0$ during containment, $-\gamma_y y_k > 0$: debt accumulates even without any discretionary fiscal expenditure. A lockdown without accompanying fiscal measures still erodes the fiscal position---the trilemma cannot be circumvented by fiscal abstinence. Combined with the necessity of suppression established in Property~1, this means that the planner's decision is not whether to deploy fiscal support but which composition, at which time to deploy---returning the problem to the instrument choice formalized in Property~3.

\noindent\textit{Summary.} Property 4 introduces the budget semi-elasticity $\gamma_y$ and establishes that fiscal abstinence is an illusion: suppression passively erodes the fiscal position even absent discretionary intervention. The planner's decision is therefore not whether to intervene fiscally, but how to compose the intervention to minimize the joint debt footprint. 

% ============================================================
\subsection{Summary}
\label{sec:Sumary}
% ============================================================
Therefore, the trilemma has the following properties: 

First, the system contains multiple simultaneously unstable dimensions: the epidemiological state is explosive without containment ($\rho_\theta > 1$); the output state becomes conditionally unstable under containment when $\rho_y + \psi S_k \geq 1$ without accompanying capacity preservation; and the debt state is unconditionally explosive, compounding any fiscal cost incurred during the crisis beyond the planning horizon ($(1+r) > 1$), defining the need of the social planner to intervene. 

Second, every available instrument is structurally coupled across state dimensions: containment suppresses infections but contracts output and generates health expenditures; demand injection stabilizes output but augments debt; capacity preservation maintains recovery capacity but carries uncertain fiscal costs. No instrument operates on a single dimension, and mortality is not directly controllable---requiring a two-period transmission through the suppression channel. 

Third, the marginal costs and benefits of each instrument are state-dependent: the output cost of containment is convex in the depth of the recession; the output effect of capacity preservation scales with $|y_k|$; and the net fiscal cost of capacity preservation declines with the severity of the crisis were signs and levels remain ambiguous.

Fourth, the system is circularly interdependent: viral prevalence raises the need for suppression (Property~1), which contracts the economy (Property~2) and necessitates fiscal compensation. This compensation generates debt (Property 3), yet fiscal abstinence fails because automatic stabilizers transmit the output contraction to the budget independently of any discretionary decision (Property~4).

\subsubsection*{Structural Uniqueness}

To establish that the pandemic trilemma represents a structurally distinct problem, I apply the four constitutive properties as a diagnostic framework to four alternative crisis types: armed conflict, natural disasters, financial crises, and climate change.\footnote{See Table~\ref{tab:crisis_comparison} in Appendix~\ref{app:crisis_comparison} for the complete diagnostic.} No other crisis type satisfies all four properties simultaneously. Armed conflict and natural disasters fail Property~2: the instruments that address the threat---military mobilization, reconstruction spending---are expansionary rather than contractionary. Financial crises fail Property~1: the shock is endogenous, and the primary interventions work with rather than against economic stabilization. Climate change satisfies all four properties in principle, but differs in time scale (decades versus quarters), observability of the health outcome (diffuse versus immediate), and terminal condition: while technological decarbonization can in principle dissolve the climate trilemma---analogous to vaccination dissolving the pandemic trilemma---this resolution operates on a multi-decadal horizon with no guaranteed terminal date. The pandemic is therefore the only crisis type in which all four properties and their circular interdependence are simultaneously operative within a tractable planning horizon providing a unique situation \footnote{For a formal analysis of the intergenerational distributional conflict inherent in the
health-economy trade-off, see Glover et al. (2023).}.


%%Everything after this is a draft
\newpage


\subsection{Modeling Approach and Relation to the Literature}
\label{sec:modeling_approach}

The model adopts a social planner framework with reduced-form transition equations and a quadratic objective function. This section locates the approach within the broader methodological landscape, justifies the key modeling choices, and introduces the behavioral response channel that links the epidemiological and economic states.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%FEAR TERM und RECHTFERTIGUNG MODELL
\subsubsection*{Reduced-Form Transitions vs.\ Structural Equilibrium}

The macroeconomic dynamics in (\ref{eq:output_full})--(\ref{eq:infection}) are specified directly as transition equations governing the evolution of the state vector, rather than derived as equilibrium conditions from optimizing household and firm behavior. This places the model in the tradition of optimal policy analysis with reduced-form constraints---an approach with established precedents in both fiscal policy \citep{barro1999} and monetary policy \citep{svensson1999}, and adopted by the leading contributions to the pandemic-macro literature \citep{alvarez2020, acemoglu2021}.

The alternative would be a fully microfounded model---a DSGE framework in which household consumption, labor supply, and firm production decisions endogenously generate the output dynamics that this model specifies directly. Such models exist for pandemic settings \citep{eichenbaum2021, kaplan2020}, but their structural richness comes at a cost. The output response to containment depends sensitively on assumptions about functional forms (the curvature of the utility function, the substitutability of home and market consumption, the degree of nominal rigidity), household heterogeneity (liquidity constraints, sectoral exposure), and market structure (monopolistic competition, Calvo pricing)---none of which are directly relevant to the question this paper asks: what is the optimal \textit{composition} of the fiscal response, given the transmission parameters governing how each instrument category affects output, debt, and infections?

The reduced-form approach offers three advantages for this question. First, the structural parameters ($\alpha_S$, $\alpha_F^{CP}$, $\alpha_F^{DI}$, $\psi$, $\eta$, $\delta$) are estimable from panel data without imposing the equilibrium restrictions that a DSGE would require---the empirical identification in Section~\ref{sec:estimation} exploits cross-country variation in policy choices conditional on the common pandemic shock, which is orthogonal to the household preference parameters that a structural model would need to specify. Second, the model accommodates the instrument disaggregation (DI vs.\ CP) without requiring a production-side theory of how short-time work schemes interact with firm-level labor demand---the transmission channels are specified at the aggregate level where the relevant policy variation occurs. Third, the quadratic objective function is interpretable as a second-order approximation to the welfare function of a representative agent \citep{woodford2003}, which provides the welfare-theoretic foundation without requiring explicit specification of the underlying utility function.\footnote{The approximation is valid when the steady state is efficient and the deviations from steady state are not too large. The pandemic application pushes the boundaries of this condition---output gaps of $-10\%$ to $-20\%$ are not ``small'' deviations. The quantitative welfare implications should therefore be interpreted with appropriate caution, while the qualitative policy rankings (instrument composition, timing) are robust to the approximation quality.}

\subsubsection*{Objective Function Weights}

The quadratic objective penalizes squared deviations of the state variables from their pre-pandemic values. The weights $w_y$, $w_d$, $w_b$ govern the planner's relative valuation of output stabilization, mortality reduction, and fiscal sustainability. Rather than deriving these weights from a single primitive, I discipline them through three complementary strategies.

First, the mortality weight $w_d$ relative to $w_y$ is calibrated using the Value of a Statistical Life (VSL). With OECD-average VSL estimates of approximately \$5--10 million and per-capita GDP of approximately \$40,000--50,000, the implied weight on one statistical death relative to one percentage point of output is on the order of $100$--$250$ \citep{viscusi2003}. This provides an empirically grounded anchor for the health-output trade-off. Second, the debt weight $w_b$ reflects the shadow cost of post-pandemic fiscal consolidation, calibrated from the fiscal space literature \citep{blanchard2019, mauro2015}. Third, the results are presented as a \textit{Pareto frontier} across weight combinations, showing how the optimal policy composition varies as the planner moves along the frontier from health-prioritizing to economy-prioritizing to debt-prioritizing regimes. This approach renders the specific weight calibration less consequential: the qualitative findings---in particular, the dominance of CP over DI during active containment and the temporal sequencing of the fiscal mix---are robust across the weight space.

\subsubsection*{Endogenous Behavioral Response: The Fear Channel}

A limitation of the baseline specification (\ref{eq:output_full}) is that the output cost of the pandemic operates exclusively through the mandated containment channel ($\alpha_S S_k$). In practice, a substantial share of the output decline during the pandemic reflected \textit{voluntary} behavioral adjustment: households reduced consumption, avoided public spaces, and withdrew from economic activity in response to perceived infection risk, independently of government mandates. \citet{goolsbee2021} decompose the decline in foot traffic during 2020 and find that shelter-in-place orders accounted for only approximately 12\% of the total reduction; the remaining 88\% was voluntary. \citet{eichenbaum2021} endogenize this channel by making consumption and labor supply functions of infection risk in a SIR-macro equilibrium.

I incorporate this channel in reduced form by adding the infection state as a direct determinant of the output gap:
%
\begin{equation}
    y_{k+1} = \rho_y^{\text{eff}} \, y_k
      \;-\; \alpha_S^{\text{eff}} \, S_k
      \;+\; \alpha_F^{DI} \, F_k^{DI}
      \;+\; \beta_\theta \, \theta_k
    \label{eq:output_fear}
\end{equation}
%
where $\beta_\theta < 0$ captures the voluntary output contraction induced by infection prevalence. This term is structurally distinct from the persistence channel ($\rho_y^{\text{eff}} y_k$): persistence propagates \textit{inherited} damage from the previous period's output gap, whereas the fear channel generates \textit{new} damage from the current epidemiological state, independently of the output gap. A country operating at potential ($y_k = 0$) but experiencing high infection prevalence ($\theta_k \gg 0$) suffers an output loss through voluntary behavioral adjustment; conversely, a country with a deep output gap ($y_k \ll 0$) but no infections ($\theta_k = 0$) experiences persistence but no fear effect.

The fear channel has two implications for the planner's optimization. First, it creates an additional output benefit of containment that operates with a one-period lag through the infection dynamics: $S_k$ reduces $\theta_{k+1}$ via the suppression equation (\ref{eq:infection}), which in turn reduces the voluntary output loss $\beta_\theta \theta_{k+1}$ in the subsequent period. Without this channel, the planner evaluates containment solely on the basis of its mortality benefit ($S_k \to \theta_{k+1} \to d_{k+2}$) against its direct output cost ($\alpha_S S_k$). With the fear channel, containment carries an additional output benefit---reducing the behavioral contraction that would have occurred regardless of the lockdown---which partially offsets the mandated output cost:
%
\begin{equation}
    \underbrace{-\alpha_S^{\text{eff}} \, S_k}_{\text{mandated cost}} \quad \text{vs.} \quad \underbrace{-|\beta_\theta| \, \Delta\theta_{k+1}(S_k)}_{\text{avoided voluntary cost}}
    \label{eq:fear_tradeoff}
\end{equation}
%
Second, the fear channel breaks the separation between health policy and economic policy that the baseline model implies. In the baseline, the planner faces a clean trade-off: containment costs output but saves lives. With the fear channel, inadequate containment costs output \textit{twice}---through the voluntary behavioral response and through the subsequent need for stricter mandated measures when infections spiral. This reinforces the early-intervention rationale established in Property~2 and provides a formal channel through which the ``do nothing'' strategy is dominated even on purely economic grounds \citep{correia2022}.

The parameter $\beta_\theta$ is calibrated from the empirical literature on voluntary behavioral responses rather than estimated from the panel, as the cross-country variation in $\theta_k$ conditional on two-way fixed effects is insufficient for identification (Section~\ref{sec:estimation}). The baseline calibration draws on \citet{goolsbee2021} and \citet{born2021}, implying that the voluntary output response is approximately five to seven times larger than the incremental mandated effect---consistent with $|\beta_\theta|/\alpha_S \approx 5$--$7$ when both are expressed per unit of effective infection pressure.\footnote{This ratio should be interpreted with caution. \citet{goolsbee2021} identify the voluntary share of \textit{foot traffic} reduction, not GDP. The mapping from foot traffic to aggregate output involves compositional effects (some sectors are more affected than others) and general equilibrium feedbacks that the reduced-form specification does not capture. The sensitivity analysis in Section~\ref{sec:sensitivity} varies $\beta_\theta$ across a wide range to assess the robustness of the optimal policy to this calibration.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% =============================================================
% Section 3 — Model Framework (Corrected Draft)
% =============================================================
%Output Transitionsgleichun gnoch anpassen
\section{Model Framework}
\label{sec:model}

This section formalizes the planner's optimization problem. The state variables,
control variables, and structural parameters were introduced in
Section~\ref{sec:trilemma}; here I specify the complete transition system, the
objective function, and the solution method.

% -----------------------------------------------------------------
\subsection{Non-Linear Transition System}
\label{sec:nonlinear}

The state of the system at time $k$ is represented by
$\bm{x}_k = (y_k,\, d_k,\, b_k,\, \theta_k)' \in \mathbb{R}^4$ and the planner's
controls by $\bm{u}_k = (S_k,\, F_k^{DI},\, F_k^{CP})' \in \mathbb{R}^3$.
Collecting the transition equations derived in Section~\ref{sec:trilemma}, the system
evolves as:
%
\begin{align}
    y_{k+1} &= \underbrace{\bigl[\rho_y + \psi\, S_k
    - \eta\, F_k^{CP}\bigr]}_{\rho_y^{\text{eff}}(S_k, F_k^{CP})}\, y_k
    \;-\; \alpha_S\, S_k \;+\; \alpha_F^{DI}\, F_k^{DI}
    \label{eq:y_nonlin} \\[1ex]
    d_{k+1} &= \delta_\theta\, \theta_k
    \label{eq:d_nonlin} \\[1ex]
    b_{k+1} &= (1+r)\, b_k \;-\; \gamma_y\, y_k
    \;+\; \kappa_F^{DI}\, F_k^{DI} \;+\; \kappa_F^{CP}\, F_k^{CP}
    \;+\; c_H\, \theta_k
    \label{eq:b_nonlin} \\[1ex]
    \theta_{k+1} &= \rho_\theta\bigl(1 - \phi_S\, S_k\bigr)\,\theta_k
    \;+\; \varepsilon_k
    \label{eq:theta_nonlin}
\end{align}
%
where $\varepsilon_k \sim \text{WN}(0,\, \sigma^2)$ represents exogenous
epidemiological shocks. Table~\ref{tab:parameters} summarizes all structural
parameters.

Two of the four equations contain bilinear terms---products of controls and states---
that prevent the application of standard linear-quadratic methods. In the output
equation~(\ref{eq:y_nonlin}), the effective persistence
$\rho_y^{\text{eff}} = \rho_y + \psi S_k - \eta F_k^{CP}$ multiplies the current
output gap, generating the products $\psi\, S_k\, y_k$ and $\eta\, F_k^{CP}\, y_k$.
In the infection equation~(\ref{eq:theta_nonlin}), the suppression effect operates
multiplicatively on current prevalence, generating the product
$\rho_\theta \phi_S\, S_k\, \theta_k$. These bilinear interactions are not a modeling
artifact but encode the state-dependent structure of the planner's problem established
in Section~\ref{sec:trilemma}: the marginal output cost of containment depends on the
depth of the recession (Property~2), the marginal benefit of capacity preservation
scales with the output gap (Property~3), and the marginal health benefit of
suppression depends on current prevalence (Property~1). By contrast, the mortality
equation~(\ref{eq:d_nonlin}) and the debt equation~(\ref{eq:b_nonlin}) are linear in
states and controls.

\begin{table}[H]
\centering
\caption{Variables and Parameters}
\label{tab:parameters}
\renewcommand{\arraystretch}{1.2}
\small
\begin{tabular}{@{} c l l @{}}
\toprule
\textbf{Symbol} & \textbf{Description} & \textbf{Equation} \\
\midrule
\multicolumn{3}{@{}l}{\textit{State variables ($\bm{x}_k \in \mathbb{R}^4$)}} \\
$y_k$ & Output gap (\% deviation of real GDP from potential) &
    (\ref{eq:y_nonlin}) \\
$d_k$ & Excess mortality (deviation of death rate from seasonal average,
    per capita) & (\ref{eq:d_nonlin}) \\
$b_k$ & Public debt (deviation of debt-to-GDP ratio from pre-pandemic
    trend, pp) & (\ref{eq:b_nonlin}) \\
$\theta_k$ & Infection rate (share of actively infected population) &
    (\ref{eq:theta_nonlin}) \\
\midrule
\multicolumn{3}{@{}l}{\textit{Control variables
    ($\bm{u}_k \in \mathbb{R}^3$)}} \\
$S_k$ & Containment intensity ($0$: no restrictions, $1$: full lockdown) &
    (\ref{eq:y_nonlin}), (\ref{eq:theta_nonlin}) \\
$F_k^{DI}$ & Demand injection (transfers, tax relief; \% of GDP) &
    (\ref{eq:y_nonlin}), (\ref{eq:b_nonlin}) \\
$F_k^{CP}$ & Capacity preservation (wage subsidies, guarantees; \% of GDP) &
    (\ref{eq:y_nonlin}), (\ref{eq:b_nonlin}) \\
\midrule
\multicolumn{3}{@{}l}{\textit{System dynamics parameters}} \\
$\rho_y$ & Baseline output gap persistence & (\ref{eq:y_nonlin}) \\
$\psi$ & Lockdown-induced hysteresis sensitivity & (\ref{eq:y_nonlin}) \\
$\delta_\theta$ & Infection-to-mortality transmission rate &
    (\ref{eq:d_nonlin}) \\
$\gamma_y$ & Budget semi-elasticity (automatic stabilizers) &
    (\ref{eq:b_nonlin}) \\
$r$ & Real interest rate on sovereign debt & (\ref{eq:b_nonlin}) \\
$c_H$ & Pandemic-specific health expenditure rate per unit infection &
    (\ref{eq:b_nonlin}) \\
$\rho_\theta$ & Epidemiological persistence ($>1$: explosive growth) &
    (\ref{eq:theta_nonlin}) \\
\midrule
\multicolumn{3}{@{}l}{\textit{Policy transmission parameters}} \\
$\alpha_S$ & Direct contractionary effect of NPIs on output &
    (\ref{eq:y_nonlin}) \\
$\alpha_F^{DI}$ & Demand injection multiplier & (\ref{eq:y_nonlin}) \\
$\eta$ & Capacity preservation effectiveness & (\ref{eq:y_nonlin}) \\
$\phi_S$ & Suppression effectiveness of NPIs on infections &
    (\ref{eq:theta_nonlin}) \\
$\kappa_F^{DI}$ & Budgetary cost of demand injection &
    (\ref{eq:b_nonlin}) \\
$\kappa_F^{CP}$ & Budgetary cost of capacity preservation &
    (\ref{eq:b_nonlin}) \\
\midrule
\multicolumn{3}{@{}l}{\textit{Objective function parameters}} \\
$\beta$ & Discount factor & (\ref{eq:objective}) \\
$w_y,\, w_d,\, w_b$ & Running cost weights (output, mortality, debt) &
    (\ref{eq:weights}) \\
$W_b$ & Terminal debt cost & (\ref{eq:weights}) \\
$r_S,\, r_F^{DI},\, r_F^{CP}$ & Control implementation costs &
    (\ref{eq:weights}) \\
\bottomrule
\end{tabular}
\end{table}



\begin{table}[t]
\centering
\caption{Structural Parameters}
\label{tab:parameters}
\renewcommand{\arraystretch}{1.2}
\small
\begin{tabular}{@{} c l l @{}}
\toprule
\textbf{Symbol} & \textbf{Description} & \textbf{Equation} \\
\midrule
\multicolumn{3}{@{}l}{\textit{System dynamics}} \\
$\rho_y$ & Baseline output gap persistence & (\ref{eq:y_nonlin}) \\
$\psi$ & Lockdown-induced hysteresis sensitivity & (\ref{eq:y_nonlin}) \\
$\delta_\theta$ & Infection-to-mortality transmission rate & (\ref{eq:d_nonlin}) \\
$\gamma_y$ & Budget semi-elasticity (automatic stabilizers) & (\ref{eq:b_nonlin}) \\
$r$ & Real interest rate on sovereign debt & (\ref{eq:b_nonlin}) \\
$c_H$ & Health expenditure rate per unit infection & (\ref{eq:b_nonlin}) \\
$\rho_\theta$ & Epidemiological persistence ($>1$: explosive) & (\ref{eq:theta_nonlin}) \\
\midrule
\multicolumn{3}{@{}l}{\textit{Policy transmission}} \\
$\alpha_S$ & Direct contractionary effect of NPIs on output & (\ref{eq:y_nonlin}) \\
$\alpha_F^{DI}$ & Demand injection multiplier & (\ref{eq:y_nonlin}) \\
$\eta$ & Capacity preservation effectiveness & (\ref{eq:y_nonlin}) \\
$\phi_S$ & Suppression effectiveness of NPIs on infections & (\ref{eq:theta_nonlin}) \\
$\kappa_F^{DI}$ & Budgetary cost of demand injection & (\ref{eq:b_nonlin}) \\
$\kappa_F^{CP}$ & Budgetary cost of capacity preservation & (\ref{eq:b_nonlin}) \\
\bottomrule
\end{tabular}
\end{table}

% -----------------------------------------------------------------
\subsection{Iterative Linearization}
\label{sec:linearization}

To apply linear-quadratic methods to the non-linear
system~(\ref{eq:y_nonlin})--(\ref{eq:theta_nonlin}), I employ an iterative
Linear-Quadratic Regulator (iLQR) approach \citep{li2004iterative}. At each
iteration $j$, the system is linearized around the current reference trajectory
$(\bar{\bm{x}}_k^{(j)},\, \bar{\bm{u}}_k^{(j)})_{k=0}^{N-1}$, yielding
time-varying system matrices:
%
\begin{equation}
    \bm{x}_{k+1} \approx f\bigl(\bar{\bm{x}}_k,\, \bar{\bm{u}}_k\bigr)
    \;+\; \bm{A}_k\bigl(\bm{x}_k - \bar{\bm{x}}_k\bigr)
    \;+\; \bm{B}_k\bigl(\bm{u}_k - \bar{\bm{u}}_k\bigr)
    \;+\; \bm{G}\,\varepsilon_k
    \label{eq:linearization}
\end{equation}
%
where $\bm{A}_k$ and $\bm{B}_k$ are the Jacobians of the transition function
evaluated at the reference trajectory:
%
\begin{equation}
    \bm{A}_k \;=\; \frac{\partial f}{\partial \bm{x}}
    \bigg|_{(\bar{\bm{x}}_k,\, \bar{\bm{u}}_k)}, \qquad
    \bm{B}_k \;=\; \frac{\partial f}{\partial \bm{u}}
    \bigg|_{(\bar{\bm{x}}_k,\, \bar{\bm{u}}_k)}, \qquad
    \bm{G} = \begin{pmatrix} 0 \\ 0 \\ 0 \\ 1 \end{pmatrix}.
    \label{eq:jacobians_def}
\end{equation}
%
Computing the partial derivatives from
(\ref{eq:y_nonlin})--(\ref{eq:theta_nonlin}):
%
\begin{equation}
\bm{A}_k = \begin{pmatrix}
\rho_y + \psi\,\bar{S}_k - \eta\,\bar{F}_k^{CP}
    & 0 & 0 & 0 \\[3pt]
0 & 0 & 0 & \delta_\theta \\[3pt]
-\gamma_y & 0 & (1+r) & c_H \\[3pt]
0 & 0 & 0 & \rho_\theta\bigl(1 - \phi_S\,\bar{S}_k\bigr)
\end{pmatrix}
\label{eq:Ak}
\end{equation}
%
\begin{equation}
\bm{B}_k = \begin{pmatrix}
\psi\,\bar{y}_k - \alpha_S
    & \alpha_F^{DI}
    & -\eta\,\bar{y}_k \\[3pt]
0 & 0 & 0 \\[3pt]
0 & \kappa_F^{DI} & \kappa_F^{CP} \\[3pt]
-\rho_\theta\,\phi_S\,\bar{\theta}_k & 0 & 0
\end{pmatrix}
\label{eq:Bk}
\end{equation}

Three features of the linearized system warrant attention. First, the matrices are
\textit{time-varying}: they depend on the reference trajectory
$(\bar{\bm{x}}_k, \bar{\bm{u}}_k)$ at each period $k$, which itself changes across
iLQR iterations. This time-dependence encodes the state-dependent marginal effects
derived in Section~\ref{sec:trilemma}: the marginal output cost of containment
$(\psi\,\bar{y}_k - \alpha_S)$ in the first row of $\bm{B}_k$ depends on the depth of
the recession, the marginal effectiveness of capacity preservation $(-\eta\,\bar{y}_k)$
scales with the output gap, and the marginal epidemiological benefit of suppression
$(-\rho_\theta\,\phi_S\,\bar{\theta}_k)$ depends on current prevalence.

Second, the mortality row of $\bm{A}_k$ is time-invariant: $d_{k+1} = \delta_\theta\,
\theta_k$ is linear in the state. The two-period policy lag
$S_k \to \theta_{k+1} \to d_{k+2}$ emerges from the state-space structure without
requiring additional dynamics in the mortality equation.

Third, two rows of $\bm{B}_k$---the mortality row and the debt row---are
time-invariant. The planner has no direct contemporaneous instrument for mortality
(second row of $\bm{B}_k$ is zero), and the budgetary cost parameters
$\kappa_F^{DI}$, $\kappa_F^{CP}$ are constant. The time-variation concentrates in the
output row (where the recession depth modulates policy effectiveness) and the infection
row (where current prevalence determines the marginal return to suppression).


% -----------------------------------------------------------------
\subsection{Objective Function}
\label{sec:objective}

The planner minimizes the expected intertemporal cost over a finite horizon $N$:
%
\begin{equation}
    J = \mathbb{E}\left\{\bm{x}_N'\,\bm{Q}_N\,\bm{x}_N
    \;+\; \sum_{k=0}^{N-1}\left(\bm{x}_k'\,\bm{Q}_k\,\bm{x}_k
    \;+\; \bm{u}_k'\,\bm{R}\,\bm{u}_k\right)\right\}
    \label{eq:objective}
\end{equation}
%
with:
%
\begin{equation}
    \bm{Q}_k = \beta^k\,\operatorname{diag}\bigl(w_y,\; w_d,\; w_b,\; 0\bigr),
    \qquad
    \bm{Q}_N = \beta^N\,\operatorname{diag}\bigl(0,\; 0,\; W_b,\; 0\bigr),
    \qquad
    \bm{R} = \operatorname{diag}\bigl(r_S,\; r_F^{DI},\; r_F^{CP}\bigr).
    \label{eq:weights}
\end{equation}
%
The zero weight on $\theta_k$ in $\bm{Q}_k$ implies that infections are penalized
only through their delayed effect on mortality---the planner has no intrinsic
preference over prevalence beyond its health consequences. The terminal cost $W_b$
penalizes exclusively debt: output gaps close endogenously as containment is lifted,
and deaths are irreversible costs already penalized through $w_d$ along the trajectory.
Debt, by contrast, compounds at rate $(1+r)$ beyond the planning horizon and constrains
the fiscal response to subsequent shocks. The control costs $r_S$, $r_F^{DI}$,
$r_F^{CP}$ capture implementation frictions---political costs of lockdowns,
administrative costs of program design---that are independent of the
state.\footnote{The standard regularity conditions hold: $\bm{Q}_k, \bm{Q}_N \succeq
0$ (satisfied for non-negative weights) and $\bm{R} \succ 0$ (satisfied for strictly
positive control costs), guaranteeing existence and uniqueness of the optimal policy at each iLQR iteration.}


% -----------------------------------------------------------------
\subsection{Solution via Iterative LQR (iLQR)}

To handle the bilinear interactions in the transition dynamics, we employ an iterative Linear-Quadratic Regulator (iLQR) approach. Unlike standard LQR, which assumes strictly linear dynamics, iLQR iteratively refines a nominal trajectory by applying a local second-order Taylor approximation to the cost-to-go function (Dynamic Programming) and a first-order approximation to the system dynamics.Let $(\bar{x}_k^{(j)}, \bar{u}_k^{(j)})_{k=0}^{N-1}$ be the nominal trajectory at iteration $j$. We define the state and control deviations as $\delta x_k = x_k - \bar{x}_k^{(j)}$ and $\delta u_k = u_k - \bar{u}_k^{(j)}$.At each time step $k$, the local action-value function (Q-function) represents the cost of applying a control perturbation $\delta u_k$ in a perturbed state $\delta x_k$. Its local quadratic expansion yields the following coefficients:

\begin{align}Q_{x,k}  &= Q_k \bar{x}k^{(j)} + A_k^T p{k+1}   & Q_{u,k}= R \bar{u}k^{(j)} + B_k^T p{k+1}\\ Q_{xx,k} &= Q_k + A_k^T P_{k+1} A_k  & Q_{uu,k}= R + B_k^T P_{k+1} B_k\\  & Q_{ux,k}= B_k^T P_{k+1} A_k\end{align}

where $p_{k+1} \in \mathbb{R}^4$ and $P_{k+1} \in \mathbb{R}^{4 \times 4}$ are the gradient and the Hessian of the optimal cost-to-go function at period $k+1$, respectively. $A_k$ and $B_k$ are the Jacobian matrices of the system evaluated at the nominal trajectory.Minimizing this quadratic Q-function with respect to $\delta u_k$ yields the optimal local control modification, which consists of an open-loop feedforward term $k_k$ and a closed-loop feedback gain $K_k$:

\begin{align}k_k &= -Q_{uu,k}^{-1} Q_{u,k} &K_k= Q_{uu,k}^{-1} Q_{ux,k}\end{align}


The feedforward term $k_k$ actively pushes the trajectory towards lower costs (gradient descent), while the feedback matrix $K_k$ stabilizes the system around this new trajectory. The cost-to-go terms are then updated moving backward in time:

\begin{align}p_k &= Q_{x,k} - K_k^T Q_{uu,k} k_k & P_k &= Q_{xx,k} - K_k^T Q_{uu,k} K_k\end{align}

The full iLQR procedure iterates between a backward pass to compute these gains and a forward pass to update the trajectory:

\begin{enumerate}\item \textbf{Backward pass.} Given the nominal trajectory $(\bar{x}_k^{(j)}, \bar{u}_k^{(j)})$, compute $A_k$ and $B_k$, and solve the recursive equations backward from $k = N-1$ to $k = 0$, initializing with $p_N = Q_N \bar{x}_N^{(j)}$ and $P_N = Q_N$.\end{enumerate}The iteration is initialized with a feasible baseline (e.g., zero controls $\bar{u}_k^{(0)} = 0$) and terminates when the reduction in total cost between successive iterations falls below a convergence threshold $\epsilon$.
% -----------------------------------------------------------------
\subsection{Certainty Equivalence}
\label{sec:ce}

At each iLQR iteration, the linearized system is affine in the state and the
objective is quadratic, satisfying the conditions for certainty equivalence
\citep[Chapter~3.1]{bertsekas2017dynamic}. The optimal feedback gain $\bm{K}_k^{(j)}$
is therefore identical to the deterministic case ($\varepsilon_k \equiv 0$).
Stochastic shocks increase the expected cost by
$\sum_{k=0}^{N-1} \sigma^2\, \bm{G}'\bm{P}_{k+1}^{(j)}\bm{G}$---reflecting the
irreducible cost of epidemiological uncertainty---but do not alter the optimal policy
rule. The planner optimizes as if future shocks equal their expected value, then
absorbs realized deviations through the state-feedback structure.


% -----------------------------------------------------------------
\subsection{Controllability}
\label{sec:controllability}

For any fixed linearization $(\bm{A}_k, \bm{B}_k)$, the controllability matrix
$\mathcal{C}_k = [\bm{B}_k \;\; \bm{A}_k\bm{B}_k \;\;
\bm{A}_k^2\bm{B}_k \;\; \bm{A}_k^3\bm{B}_k]$ has full row rank under
non-degeneracy conditions ($\delta_\theta,\, \phi_S,\, \alpha_S,\, \alpha_F^{DI},\,
\eta,\, \kappa_F^{DI},\, \kappa_F^{CP} \neq 0$ and $\bar{\theta}_k,\,
\bar{y}_k \neq 0$). Although mortality is decoupled from $\bm{B}_k$ (the second row
is zero), it is reachable through $\bm{A}_k\bm{B}_k$ via the composite parameter
$\delta_\theta\,\rho_\theta\,\phi_S\,\bar{\theta}_k$, confirming the two-period
transmission lag established in Section~\ref{sec:trilemma}.

The eigenvalues of $\bm{A}_k$ are
$\bigl\{\rho_y + \psi\bar{S}_k - \eta\bar{F}_k^{CP},\;\; 0,\;\; (1+r),\;\;
\rho_\theta(1-\phi_S\bar{S}_k)\bigr\}$. For any positive real interest rate, the
debt eigenvalue $(1+r) > 1$ lies outside the unit circle: the debt state is explosive
in the absence of intervention. The infection eigenvalue
$\rho_\theta(1-\phi_S\bar{S}_k)$ is explosive when suppression is insufficient to
push the effective reproduction rate below unity
($\phi_S \bar{S}_k < 1 - 1/\rho_\theta$). Active policy is therefore required not
merely for optimization but for stabilization of the system.

\section{Data}
\label{sec:data}

\section{Estimation}
\label{sec:estimation}

\section{Results}
\label{sec:results}
\begin{table}[htbp]
\centering
\caption{Output Equation: Panel Estimation Results}
\label{tab:output_equation}
\small
\begin{tabular}{lcccc}
\toprule
 & (1) & (2) & (3) & (4) \\
 & Pooled OLS & Country FE & Two-Way FE & Two-Way FE \\
\midrule
$S_k$ ($-\alpha_S$)
  & $-0.071^{**}$  & $-0.106^{***}$ & $-0.028^{*}$   & $-0.028^{*}$   \\
  & (0.024)        & (0.018)        & (0.014)        & (0.013)        \\[4pt]
$y_{k-1}$ ($\rho_y$)
  & $0.622^{***}$  & $0.192^{*}$    & $0.064$        & $0.061$        \\
  & (0.117)        & (0.087)        & (0.105)        & (0.106)        \\[4pt]
$F^{CP}_k$ ($\alpha^{CP}_F$)
  & $0.293^{*}$    & $0.438^{**}$   & $0.253^{***}$  & $0.252^{***}$  \\
  & (0.145)        & (0.146)        & (0.074)        & (0.074)        \\[4pt]
$S_k \cdot y_{k-1}$ ($\psi$)
  & $-0.002$       & $-0.000$       & $0.004^{**}$   & $0.004^{**}$   \\
  & (0.003)        & (0.002)        & (0.001)        & (0.001)        \\[4pt]
$S_k \cdot F^{CP}_k$ ($\gamma$)
  & $-0.018^{***}$ & $-0.019^{***}$ & $-0.006^{**}$  & $-0.006^{**}$  \\
  & (0.004)        & (0.003)        & (0.002)        & (0.002)        \\[4pt]
$F^{DI}_{k-2}$
  & $0.828^{***}$  & $0.503^{***}$  & $0.244^{*}$    &                \\
  & (0.203)        & (0.101)        & (0.119)        &                \\[4pt]
$F^{DI}_{k-1}$
  &                &                &                & $-0.017$       \\
  &                &                &                & (0.095)        \\[4pt]
\midrule
Country FE       & No  & Yes & Yes & Yes \\
Quarter FE       & No  & No  & Yes & Yes \\
SE               & HC1 & HC1 & HC1 & HC1 \\
Observations     & 342 & 342 & 342 & 342 \\
Countries        & 38  & 38  & 38  & 38  \\
\bottomrule
\multicolumn{5}{p{12cm}}{\footnotesize \textit{Notes:} Dependent variable is the real output gap $y_k$ (pp deviation from HP-filtered potential). $S_k$ is the Oxford Stringency Index (0--100). $F^{CP}_k$ and $F^{DI}_k$ are fiscal flows in \% of 2019 GDP. Column (3) is the main specification. HC1 cluster-robust standard errors (clustered by country) in parentheses. Sample: 38 OECD countries, Q1.2020--Q1.2022. $^{*}p<0.10$; $^{**}p<0.05$; $^{***}p<0.01$.} \\
\end{tabular}
\end{table}

\begin{table}[htbp]
\centering
\caption{Debt Equation: Panel Estimation Results}
\label{tab:debt_equation}
\small
\begin{tabular}{lcccc}
\toprule
 & (1) & (2) & (3) & (4) \\
 & Pooled OLS & Country FE & Country FE & First Diff. \\
 &            & (pooled CP)& (split CP) &  (split CP) \\
\midrule
$y_k$ ($-\gamma_y$)
  & $-0.131^{***}$ & $-0.206^{***}$ & $-0.191^{***}$ & $-0.188^{***}$ \\
  & (0.030)        & (0.049)        & (0.050)        & (0.045)        \\[4pt]
$F^{CP}_k$ ($\kappa^{CP}$)
  & $0.243^{***}$  & $0.193^{**}$   &                &                \\
  & (0.058)        & (0.066)        &                &                \\[4pt]
$F^{CP,above}_k$ ($\kappa^{CP}_{above}$)
  &                &                & $0.521^{.}$    & $0.628^{*}$    \\
  &                &                & (0.267)        & (0.274)        \\[4pt]
$F^{CP,below}_k$ ($\kappa^{CP}_{below}$)
  &                &                & $0.466^{**}$   & $0.507^{*}$    \\
  &                &                & (0.168)        & (0.202)        \\[4pt]
$F^{DI}_{k-1}$ ($\kappa^{DI}$)
  & $0.551^{.}$    & $0.475^{.}$    & $0.468^{.}$    & $0.521^{*}$    \\
  & (0.313)        & (0.262)        & (0.276)        & (0.218)        \\[4pt]
Linear trend
  & $-0.090$       & $-0.069$       & $-0.072$       &                \\
  & (0.070)        & (0.070)        & (0.072)        &                \\[4pt]
\midrule
Country FE       & No  & Yes & Yes & (FD) \\
Quarter FE       & No  & No  & No  & No   \\
SE               & HC1 & HC1 & HC1 & HC1  \\
Observations     & 418 & 418 & 418 & 380  \\
Countries        & 38  & 38  & 38  & 38   \\
\bottomrule
\multicolumn{5}{p{12cm}}{\footnotesize \textit{Notes:} Dependent variable is the real debt change $\Delta b^R_k$ (pp of 2019 GDP per quarter). $F^{CP,above}_k$ denotes above-the-line CP (short-time work, direct grants) and $F^{CP,below}_k$ denotes below-the-line CP (loan guarantees, adjusted to 35\% estimated take-up rate). Column (3) is the main specification. HC1 cluster-robust standard errors (clustered by country) in parentheses. The Mundlak (1978) auxiliary regression rejects the unrelated effects assumption ($p = 0.006$), confirming that country fixed effects are required. Sample: 38 OECD countries, Q1.2020--Q4.2022. $^{.}p<0.10$; $^{*}p<0.05$; $^{**}p<0.01$; $^{***}p<0.001$.} \\
\end{tabular}
\end{table}

\begin{table}[htbp]
\centering
\caption{Policy Efficiency Matrix: Output Stabilization vs.\ Fiscal Costs}
\label{tab:policy_matrix}
\small
\begin{tabular}{lccc}
\toprule
 & Output Effect & Debt Cost & Output per \\
 & $\hat{\alpha}_F$ & $\hat{\kappa}_F$ & Debt Unit \\
\midrule
\textit{Capacity Preservation} \\[2pt]
\quad Pooled CP
  & $0.253^{***}$ & $0.193^{**}$  & 1.31 \\
  & (0.074)       & (0.066)       &      \\[4pt]
\quad Above-the-line (STW)
  & ---           & $0.521^{.}$   & ---  \\
  &               & (0.267)       &      \\[4pt]
\quad Below-the-line (Guarantees, adj.)
  & ---           & $0.466^{**}$  & ---  \\
  &               & (0.168)       &      \\[6pt]
\textit{Demand Injection} \\[2pt]
\quad $F^{DI}_{k-2}$ on output
  & $0.244^{*}$   &               &      \\
  & (0.119)       &               &      \\[4pt]
\quad $F^{DI}_{k-1}$ on debt
  &               & $0.468^{.}$   & 0.52 \\
  &               & (0.276)       &      \\[6pt]
\midrule
\multicolumn{4}{p{10cm}}{\footnotesize \textit{Notes:} Output effects from Table~\ref{tab:output_equation}, Column (3) (Two-Way FE). Debt costs from Table~\ref{tab:debt_equation}, Columns (2)--(3) (Country FE). ``Output per Debt Unit'' is the ratio $\hat{\alpha}_F / \hat{\kappa}_F$; values $>1$ indicate that the output stabilization exceeds the fiscal cost. The DI ratio uses $\hat{\alpha}^{DI}_F = 0.244$ (lag 2) and $\hat{\kappa}^{DI}_F = 0.468$ (lag 1), reflecting the asymmetric transmission timing. CP is reported pooled in the output equation (both components contribute to stabilization as a package) and split in the debt equation (budgetary costs differ by instrument type). $^{.}p<0.10$; $^{*}p<0.05$; $^{**}p<0.01$; $^{***}p<0.001$.} \\
\end{tabular}
\end{table}

\section{Conclusion}
\label{sec:conclusion}


\section{Further (possible) Steps}

\paragraph{Step 1: Baseline Implementation}
Numerical solution of the baseline LQR model in MATLAB, including 
backward Riccati recursion, forward simulation under stochastic 
shocks, and sensitivity analysis with respect to the cost weights 
$(w_y, w_d, w_b)$.


\paragraph{Step 2: Fiscal Policy Disaggregation}
Extend the scalar fiscal control to a four-dimensional vector 
distinguishing instruments by design (expenditure vs.\ revenue) 
and target (households vs.\ firms):
\begin{equation}
    F_k \to \bm{F}_k = 
    \begin{pmatrix} 
        F_k^{Exp,HH} & F_k^{Exp,Firm} & F_k^{Rev,HH} & F_k^{Rev,Firm} 
    \end{pmatrix}'
\end{equation}

\begin{table}[h]
\centering
\caption{Taxonomy of Fiscal Instruments}
\label{tab:fiscal_taxonomy}
\renewcommand{\arraystretch}{1.4}
\begin{tabular}{p{3cm} p{5cm} p{5cm}}
\toprule
 & \textbf{Households} & \textbf{Firms} \\
\midrule
\textbf{Expenditure} 
    & $F_k^{Exp,HH}$ 
    & $F_k^{Exp,Firm}$ \\
    & \small\textit{Stimulus checks, unemployment benefits, social transfers} 
    & \small\textit{Payroll subsidies, grants, PPP loans, short-time work schemes} \\[2ex]
\textbf{Revenue} 
    & $F_k^{Rev,HH}$ 
    & $F_k^{Rev,Firm}$ \\
    & \small\textit{Income tax deferrals, social contribution waivers} 
    & \small\textit{Corporate tax relief, VAT reductions, accelerated depreciation} \\
\bottomrule
\end{tabular}
\end{table}

\noindent This taxonomy permits instrument-specific multipliers 
$\alpha_F^{j}$ and debt impacts $\kappa_F^{j}$ for 
$j \in \{Exp\text{-}HH, Exp\text{-}Firm, Rev\text{-}HH, Rev\text{-}Firm\}$. 
The extended control matrix becomes:
\begin{equation}
    \bm{B} = 
    \begin{pmatrix}
        -\alpha_S & \alpha_F^{Exp,HH} & \alpha_F^{Exp,Firm} & 
        \alpha_F^{Rev,HH} & \alpha_F^{Rev,Firm} \\
        0 & 0 & 0 & 0 & 0 \\
        0 & \kappa_F^{Exp,HH} & \kappa_F^{Exp,Firm} & 
        \kappa_F^{Rev,HH} & \kappa_F^{Rev,Firm} \\
        -\phi_S & 0 & 0 & 0 & 0
    \end{pmatrix}
\end{equation}

\noindent The empirical literature suggests a multiplier hierarchy:
\begin{equation}
    \alpha_F^{Exp,HH} > \alpha_F^{Exp,Firm} > 
    \alpha_F^{Rev,HH} > \alpha_F^{Rev,Firm}
\end{equation}
reflecting that expenditure measures and household-targeted 
transfers exhibit stronger demand effects due to higher marginal 
propensities to consume among liquidity-constrained agents 
\citep{coibion2020stimulus, chetty2020e}.



\paragraph{Step 3: Calibration}
Calibrate the model using cross-country panel data on OECD 
economies, combining:
\begin{itemize}
    \item Own calculation based on the data of Oxford COVID-19 Government Response Tracker
    \item Own collected Dataset about $F_k$ that allows to distinguish between design and target
    \item OECD Economic Outlook (output gaps, debt ratios)
    \item Excess mortality data (Our World in Data)
\end{itemize}


\paragraph{Step 4: Empirical Validation}
Estimate the causal effects of containment and fiscal policies 
using Local Projections \citep{Jorda2025} with a panel structure (also an LP-DiD could be possible):
\begin{equation}
    y_{i,t+h} - y_{i,t-1} = \alpha_i^h + \gamma_t^h + 
    \beta_S^h S_{i,t} + \beta_F^h F_{i,t} + \bm{\Gamma}^h \bm{X}_{i,t} + 
    \varepsilon_{i,t}^h
\end{equation}
for horizons $h = 0, 1, \ldots, H$. Impulse responses will be 
compared to the theoretical predictions from the calibrated LQR 
model.

\newpage

\section{Introduction}

The objective of this work is to develop an optimal policy response framework for a economic social planner navigating a pandemic-induced trilemma of the three state variables: the simultaneous protection of public health, the mitigation of economic disruption, and the preservation of long-term fiscal sustainability. Unlike a war, which is characterized by the destruction of physical capital \citep{jorda2022longer}, or a traditional economic crisis driven by endogenous imbalances, a pandemic acts as an exogenous shock that abruptly paralyzes economic flows and human capital. While the underlying physical capital remains intact and the pre-pandemic circular flows were fundamentally stable \citep{imf2019weo}, the system is confronted by an inherently non-stationary epidemiological progression. Consequently, the primary goal of the economic social planner is to act and stabilize the system and keep it as close as possible to its pre-pandemic baseline—defined here as the origin of the state space of those state variables.
% zeigt das man zum state space zurück sollte \citep{portes2020economic}

The problem to control the pandemic pressure through non-pharmaceutical interventions (NPI) is inherently dynamic, characterized by critical intertemporal trade-offs between the timing, intensity, and duration of containment measures. A failure to intervene early may result in significantly higher viral prevalence later, necessitating more stringent and costly containment measures in subsequent periods. Conversely, delayed intervention risks catastrophic overburdening of the healthcare system and a surge in excess mortality, creating path-dependency where initial hesitance constrains the future policy space. This effects of the intertemporality of the NPI is well documented in recent literature, with evidence favoring early \citep{acemoglu2021}, stringent \citep{balmford2020}, but time-limited \citep{alvarez2020} lockdowns to preserve a broader state space as the pandemic evolves \citep{kargi2023, loayza2020}---though optimal policy depends heavily on the assumed value of a statistical life \citep{alvarez2021}.

%banale economic effect auch-> WOO?
The decisions made by the social planner to control the pandemic pressure inherently influence other state spaces. A stringent or prolonged lockdown can cause profound economic disruption, resulting in reduced incomes for firms and households while generating a high degree of uncertainty \citep{guerrieri2022, eichenbaum2021macro, portes2020economic}. Such socio-economic deterioration can, in turn, lead to secondary crises what could lead to diminish the effectiveness further interventions if the populace perceives the cost of compliance as ruinous. %Maybe quelle 
Consequently, social planners globally implemented extensive fiscal support programs to mitigate these adverse effects \citep{erokhin2020}. 

However, these fiscal measures introduce their own inherently dynamic and intertemporal challenges, adding a second control dimension that substantially complicates the planner's optimization problem. In 
contrast to the well-documented effects of containment policies, the literature on pandemic fiscal responses remains sparse—particularly regarding three dimensions: (i) the interaction effects between 
containment stringency and fiscal support, (ii) the heterogeneous transmission mechanisms of different fiscal instruments and their consequences, and (iii) the joint optimization of health, economic, and fiscal objectives from the perspective of a unified decision-maker. In practice, policymakers faced this integrated problem, yet academic work has largely treated pandemic control and fiscal stabilization as separate domains.

%Maybe the first statement is a problem for the empirics

While the timing of fiscal support is largely dictated by the implementation of containment measures \citep{deb2021}, the optimization problem extends beyond mere volume to encompass instrument design and targeting. First, instrument choice matters: broad-based transfers such as stimulus checks yield limited stabilization when consumption is physically constrained, generating fiscal costs without commensurate economic benefits \citep{chetty2020e, coibion2020stimulus}. Conversely, credit guarantees though ostensibly budget-neutral—may fail to address acute liquidity needs and merely postpone fiscal burdens \citep{fariaecastro2020f}. Second, targeting is critical: pandemic-induced supply constraints suppress fiscal multipliers \citep{auerbach2022}, making the marginal impact per dollar spent a first-order concern. Poorly targeted programs risk simultaneously failing to stabilize the 
economy and eroding fiscal space \citep{bayer2023c}. The U.S.\ Paycheck Protection Program illustrates this failure: approximately three-quarters of the \$800 billion disbursed accrued to households in the top income quintile, largely bypassing the intended beneficiaries \citep{Autor2022}.

The heterogeneity of fiscal interventions—defined by their design, targeting, and the specific economic context—yields divergent outcomes for both economic recovery and public debt. Critically, the efficacy of these measures is state-dependent, as their transmission mechanisms are intrinsically linked to the level of pandemic containment. This interdependence forces the social planner to face an asymmetric trade-off: immediate gains in economic stabilization must be weighed against the long-term fiscal footprint of each measure, all while accounting for the current stringency of suppression measures. This additional layer of complexity transforms the bilateral health-economy trade-off into a dynamic trilemma. Within this constellation, the social planner must simultaneously optimize the intensity of pandemic controls and the magnitude, nature, and targeting of fiscal support. Consequently, the planner's task transcends a series of isolated, static decisions; it constitutes a multi-stage optimization problem where the dimensionality of the state space exceeds the available control instruments. To restore the system to its equilibrium, the planner must synthesize intertemporal information, integrating historical trajectories, current constraints, and future expectations under uncertainty.

%Noch etwas mehr auf die heterogenität cross country-> es gibt doch eine Quelle?

To resolve this complex dynamic multi-stage optimization problem and analyze the optimal behaviour of a social planner during a pandemic with the given controls and states, this paper employs the Discrete-Time Linear-Quadratic Regulator (LQR) framework based on the methodology of \citep{bertsekas2017dynamic}. This choice is justified by three fundamental characteristics of the pandemic trilemma. First, the LQR approach allows for the simultaneous management of multiple, interconnected state dimensions, capturing the feedback loops between health, economy, and debt, while accounting for their respective weights, as no single dimension can be neglected or be considered in isolation . Second, the use of quadratic cost functions reflects the social planner's disproportionate aversion to systemic collapses—such as a healthcare failure or a sovereign default—and prioritises proximity to the origin, as the primary goal is to restore the system to its pre-pandemic steady state. Finally, the resulting optimal feedback law provides a robust decision rule that remains valid even under the influence of stochastic shocks, a feature essential for navigating the unpredictable nature of a global pandemic.

While there are literature to various system dynamics—ranging from epidemiological patterns in SIR-based models \citep{alvarez2021, acemoglu2021} to non-linear socio-economic feedbacks, are inherently complex, the choice of a linear-quadratic approach is justified by its ability to provide tractable and robust control rules. This is achieved by linearizing the system around its stable pre-pandemic equilibrium: the origin of the state space.

%Taylor Series First Order
%Robuste Feedbacks
%Muss aber erklären kann
%Extreme Shocks = nicht möglich/greifbar
%Passt für den ersten Schritt
%Braucht genug Eigenwerte um stabil zu sein
%Noch erklären warum nicht DSGE, warum nicht Monetary Policy = passiv & Daten zeigen das-> sehr gleich-> Geld zur Verfügung gestellt
%In the context of the COVID-19 pandemic, fiscal policy emerged as the primary tool for systemic stabilization, while monetary policy transitioned into a supportive role. This hierarchy is justified by the fact that many advanced economies operated at the \textit{Zero Lower Bound} (ZLB), where conventional monetary instruments lose their efficacy. Consequently, the central banks' primary function was to act as an \textit{enabler} of fiscal intervention. By maintaining low sovereign bond yields through large-scale asset purchases, monetary policy secured the necessary ``fiscal space'' for the social planner. This allowed for the implementation of targeted transfers that address the granular socio-economic disruptions of the crisis—a task beyond the reach of the broader, indirect transmission mechanisms of monetary policy \citep{bartsch2020all}.

%erster vorschlag
This work broadens the view to a multidimensional setting while synthesizing findings from the literature (e.g., the household transfer multiplier) to recenter the analysis on the navigation of exogenous shocks from the perspective of an economic social planner. Consequently, it contributes to the literature in three key aspects. i) Extending the canonical health-economy trade-off by formally integrating fiscal sustainability as a third, endogenous state variable in the social planner's optimization problem. While prior work has largely missed out differences in the consequences of the fiscal measures, this framework models the dynamic feedback between containment policies, economic support, and sovereign debt accumulation. ii) The paper employs a LQR framework following \citep{bertsekas2017dynamic} to derive closed-form policy rules for pandemic management. This approach supplements the the gap between complex epidemiological simulations---and stylized economic models that abstract from control-theoretic foundations. The resulting economic feedback laws are robust to stochastic shocks and admit straightforward interpretation for policymakers. iii) The analysis provides a generalizable template for managing exogenous shocks with circulatory propagation mechanisms from an economic point of view. By characterizing optimal policy as a function of state-dependent feedback gains, the framework can be adapted to future crises that share the defining features of pandemic shocks.



\newpage

\section{Empirical Validation}

\subsection{Motivation}

The theoretical framework developed in Section 2 yields testable predictions about the dynamic transmission of pandemic policy interventions. The LQR model postulates specific channels through which containment measures ($S_k$) and fiscal support ($F_k$) propagate to the trilemma state variables: the output gap, excess mortality, and public debt. However, the model's practical relevance hinges on whether these theoretical transmission mechanisms find empirical support in observed policy responses.

Three features of the theoretical framework motivate a rigorous empirical investigation. First, the model predicts asymmetric timing: fiscal stimulus stabilises output contemporaneously ($F_k \to y_{k+1}$), whereas the mortality benefits of social distancing materialise with a structural lag ($S_k \to \theta_{k+1} \to d_{k+2}$). Second, the disaggregation of fiscal policy by design (expenditure vs.\ revenue) and target (households vs.\ firms) implies heterogeneous multipliers $\alpha^j_F$ and debt impacts $\kappa^j_F$. Third, the interaction between containment stringency and fiscal effectiveness suggests state-dependent transmission—a hypothesis largely untested in the existing literature.

To validate these predictions, this section develops an empirical strategy that leverages cross-country variation in pandemic policy responses across OECD economies during 2020--2021. The analysis proceeds in three steps: (i) establishing causal identification through policy reaction function residuals, (ii) estimating impulse response functions via panel Local Projections, and (iii) comparing the empirical dynamics to the calibrated theoretical model.

\subsection{Research Questions}

The empirical analysis addresses three interconnected research questions that map directly onto the theoretical framework:

\begin{enumerate}
    \item \textbf{Main Effects:} Do the transmission channels postulated in the state-space model~\eqref{eq:transition} find empirical support? Specifically:
    \begin{itemize}
        \item Does containment stringency ($S$) contract economic output while suppressing viral prevalence?
        \item Does fiscal support ($F$) stimulate the output gap while accumulating public debt?
        \item Does the structural policy lag $S_k \to \theta_{k+1} \to d_{k+2}$ manifest in the estimated impulse responses?
    \end{itemize}
    
    \item \textbf{Fiscal Heterogeneity:} Do different fiscal instruments exhibit systematically different multipliers and debt impacts? The theoretical extension in Section 4 predicts a multiplier hierarchy:
    \begin{equation}
        \alpha^{Exp,HH}_F > \alpha^{Exp,Firm}_F > \alpha^{Rev,HH}_F > \alpha^{Rev,Firm}_F
    \end{equation}
    reflecting higher marginal propensities to consume among liquidity-constrained households.
    
    \item \textbf{Interaction Effects:} Is fiscal effectiveness state-dependent? Specifically, does the output multiplier of fiscal support vary systematically with the prevailing level of containment stringency?
\end{enumerate}
\newpage
\subsection{Data}

The empirical analysis combines four data sources covering all 38 OECD member states during 2020--2021.

\paragraph{Containment Policy.} Daily stringency indices are obtained from the Oxford COVID-19 Government Response Tracker \citep{hale2021global}. The composite stringency index aggregates nine containment indicators (school closures, workplace closures, public event cancellations, restrictions on gatherings, public transport closures, stay-at-home requirements, internal movement restrictions, international travel controls, and public information campaigns) into a continuous measure scaled 0--100.

\paragraph{Fiscal Policy.} Micro-level fiscal measures are drawn from a newly constructed database documenting 2,290 fiscal policy interventions across all OECD countries \citep{pesenti2025codebook}. Each measure is classified by policy instrument (49 codes), target group (households vs.\ firms), and IMF fiscal accounting category (above-the-line, below-the-line, tax deferrals). Fiscal commitments are recorded as percentage of 2019 GDP, with disbursed amounts used where available and announced ceilings for guarantees and loan facilities.

The granular structure of the fiscal database enables disaggregation into four instrument categories that map directly onto the theoretical model:

\begin{table}[htbp]
\centering
\caption{Mapping Fiscal Database to Theoretical Model}
\label{tab:fiscal_mapping}
\begin{tabular}{llll}
\toprule
Model Variable & PolicyCategory Codes & Description & IMF Category \\
\midrule
$F^{Exp,HH}_k$ & 35--39 & Direct transfers, unemployment benefits & Above-line \\
$F^{Exp,Firm}_k$ & 45--48 & Income support, payroll subsidies & Above-line \\
$F^{Rev,HH}_k$ & 15--22, 42 & Individual tax measures, HH loans & Above/Below \\
$F^{Rev,Firm}_k$ & 1--14, 40--41, 43 & Business tax measures, guarantees & Above/Below \\
$F^{Health}_k$ & 30--34, general & Health system expenditure & Above-line \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Macroeconomic Outcomes.} Quarterly data on output gaps and debt-to-GDP ratios are sourced from the OECD Economic Outlook database. Output gaps are measured as percentage deviations of real GDP from potential output.

\paragraph{Health Outcomes.} Weekly excess mortality data are obtained from the Human Mortality Database and Our World in Data, defined as the deviation of observed deaths from the historical seasonal average, normalised by population.

\subsection{Identification Strategy}

\subsubsection{The Endogeneity Problem}

Causal identification of policy effects faces a fundamental challenge: both containment stringency ($S_{i,t}$) and fiscal support ($F_{i,t}$) are endogenous. Governments tighten restrictions in response to rising infections and deploy fiscal resources when economic conditions deteriorate. Naive estimation of
\begin{equation}
    y_{i,t+h} = \alpha_i + \gamma_t + \beta_S S_{i,t} + \beta_F F_{i,t} + \varepsilon_{i,t+h}
\end{equation}
would yield biased coefficients, as $\text{Cov}(S_{i,t}, \varepsilon_{i,t+h}) \neq 0$ and $\text{Cov}(F_{i,t}, \varepsilon_{i,t+h}) \neq 0$.

\subsubsection{Policy Reaction Function Residuals}

To address endogeneity, we adopt a two-stage approach that isolates \textit{unanticipated} policy variation. The identification assumption is that, conditional on the information set available to policymakers, residual policy variation is orthogonal to unobserved determinants of future outcomes.

\paragraph{Stage 1: Containment Policy Reaction Function.} Using daily data, we estimate the systematic component of stringency decisions:
\begin{equation}
    S_{i,t} = \alpha^S_i + \delta^S_{week(t)} + \beta_1 \log(\theta_{i,t-7}) + \beta_2 \Delta d_{i,t-14} + \beta_3 S_{i,t-1} + \nu^S_{i,t}
    \label{eq:reaction_S}
\end{equation}
where $\theta_{i,t-7}$ denotes lagged infection rates (7-day lag reflecting reporting delays), $\Delta d_{i,t-14}$ captures changes in mortality (14-day lag reflecting the infection-to-death interval), country fixed effects $\alpha^S_i$ absorb time-invariant institutional differences, and week fixed effects $\delta^S_{week(t)}$ control for global pandemic waves affecting all countries simultaneously.

The residual $\hat{\nu}^S_{i,t}$ captures stringency variation unexplained by the epidemiological state—reflecting idiosyncratic political constraints, regional outbreaks, or policy experimentation.

\paragraph{Stage 2: Fiscal Policy Reaction Function.} For fiscal measures, we estimate instrument-specific reaction functions on announcement dates:
\begin{equation}
    F^j_{i,t} = \alpha^{F,j}_i + \gamma^{F,j}_{month(t)} + \phi^j_1 S_{i,t} + \phi^j_2 y_{i,q(t)-1} + \phi^j_3 b_{i,q(t)-1} + \nu^{F,j}_{i,t}
    \label{eq:reaction_F}
\end{equation}
for $j \in \{Exp\text{-}HH, Exp\text{-}Firm, Rev\text{-}HH, Rev\text{-}Firm\}$. The specification conditions on contemporaneous stringency ($S_{i,t}$, as fiscal timing is largely dictated by containment implementation), lagged output gap ($y_{i,q-1}$), and lagged debt ratio ($b_{i,q-1}$).

Health expenditure ($F^{Health}$) enters as a control variable rather than an outcome of interest, as its primary transmission channel operates through $\theta$ rather than $y$.

\paragraph{Aggregation to Quarterly Frequency.} For the Local Projection estimation, daily residuals are aggregated to quarterly frequency:
\begin{align}
    \tilde{\nu}^S_{i,q} &= \frac{1}{|Q_q|} \sum_{t \in Q_q} \hat{\nu}^S_{i,t} \\
    \tilde{\nu}^{F,j}_{i,q} &= \sum_{t \in Q_q} \hat{\nu}^{F,j}_{i,t}
\end{align}
where stringency residuals are averaged (as $S$ represents a continuous policy stance) and fiscal residuals are summed (as $F$ represents discrete interventions with cumulative effects).

\subsubsection{Identifying Assumptions}

The validity of this identification strategy rests on two assumptions:

\begin{enumerate}
    \item \textbf{Conditional Exogeneity:} After controlling for the observable information set in equations~\eqref{eq:reaction_S} and~\eqref{eq:reaction_F}, the residual policy variation is uncorrelated with unobserved determinants of future outcomes:
    \begin{equation}
        \mathbb{E}[\nu^S_{i,t} \cdot \varepsilon_{i,t+h} \,|\, \mathcal{I}_{i,t}] = 0 \quad \text{and} \quad \mathbb{E}[\nu^{F,j}_{i,t} \cdot \varepsilon_{i,t+h} \,|\, \mathcal{I}_{i,t}] = 0
    \end{equation}
    
    \item \textbf{No Anticipation:} Economic agents do not systematically adjust behaviour in anticipation of the unanticipated policy component. This assumption is plausible given the unprecedented nature of the pandemic shock.
\end{enumerate}

\subsection{Empirical Specification}

\subsubsection{Panel Local Projections}

Following \citet{jorda2005estimation} and \citet{jorda2025local}, we estimate impulse response functions using panel Local Projections (LP). For each outcome variable $x \in \{y, d, b\}$ and horizon $h = 0, 1, \ldots, H$:
\begin{equation}
    x_{i,q+h} - x_{i,q-1} = \alpha^h_i + \gamma^h_q + \beta^h_S \tilde{\nu}^S_{i,q} + \sum_{j} \beta^{h,j}_F \tilde{\nu}^{F,j}_{i,q} + \Gamma^h X_{i,q} + \varepsilon^h_{i,q}
    \label{eq:LP_baseline}
\end{equation}
where $\alpha^h_i$ denotes country fixed effects, $\gamma^h_q$ denotes time fixed effects, and $X_{i,q}$ includes lagged dependent variables and control variables (including health expenditure $\tilde{\nu}^{F,Health}_{i,q}$).

The coefficient $\beta^h_S$ traces out the dynamic response of outcome $x$ to a one-unit unanticipated stringency shock at horizons $h = 0, \ldots, H$. Similarly, $\beta^{h,j}_F$ captures the response to a one-percentage-point-of-GDP unanticipated fiscal shock of type $j$.

\paragraph{Horizon Selection.} Given the sample period (2020Q1--2021Q4, approximately 8 quarters) and the requirement for post-treatment observations, we set $H = 6$ quarters. This horizon is sufficient to capture the medium-run dynamics while maintaining adequate degrees of freedom.

\paragraph{Lag Augmentation.} To address serial correlation in the residuals, we include $p = 2$ lags of the dependent variable following \citet{jorda2025local}:
\begin{equation}
    x_{i,q+h} - x_{i,q-1} = \ldots + \sum_{\ell=1}^{2} \rho^h_\ell \Delta x_{i,q-\ell} + \varepsilon^h_{i,q}
\end{equation}

\subsubsection{Interaction Specification}

To test whether fiscal effectiveness varies with containment stringency, we extend the baseline specification:
\begin{equation}
    x_{i,q+h} - x_{i,q-1} = \ldots + \beta^{h,j}_F \tilde{\nu}^{F,j}_{i,q} + \beta^{h,j}_{F \times S} \left( \tilde{\nu}^{F,j}_{i,q} \times \bar{S}_{i,q} \right) + \ldots
    \label{eq:LP_interaction}
\end{equation}
where $\bar{S}_{i,q}$ denotes the quarterly average stringency level. A negative $\beta^{h,j}_{F \times S}$ would indicate that fiscal multipliers are attenuated under stricter containment—consistent with the hypothesis that supply constraints suppress demand-side stimulus.

\subsubsection{Inference}

Standard errors are computed using the Wild Cluster Bootstrap with Webb weights \citep{webb2023wild}, clustering at the country level. This approach is robust to heteroskedasticity and within-cluster correlation while providing reliable inference with a moderate number of clusters ($N = 38$).

Confidence intervals are constructed at the 90\% level, acknowledging the exploratory nature of some specifications given sample size constraints.
\newpage
\subsection{Extensions}

\subsubsection{Local Projection Difference-in-Differences}

As a robustness check, we implement an intensity-based LP-DiD design that exploits cross-country variation in initial lockdown stringency. Define treatment intensity as the cumulative stringency during the first eight weeks of the pandemic:
\begin{equation}
    D_i = \sum_{t=1}^{56} S_{i,t}
\end{equation}

The LP-DiD specification becomes:
\begin{equation}
    x_{i,q+h} - x_{i,q_0-1} = \alpha^h_i + \gamma^h_q + \beta^h \cdot D_i \cdot \mathbf{1}[q \geq q_0] + \varepsilon^h_{i,q}
    \label{eq:LP_DiD}
\end{equation}
where $q_0$ denotes the onset quarter (2020Q1 for most countries). This specification circumvents the cohort-size problem inherent in staggered adoption designs with small $N$ while exploiting the full variation in lockdown intensity.

\subsubsection{Counterfactual Simulations}

To assess the quantitative implications of the estimated effects, we conduct counterfactual simulations using the calibrated LQR model. The procedure involves three steps:

\begin{enumerate}
    \item \textbf{Calibration:} Map the estimated LP coefficients $\{\hat{\beta}^h_S, \hat{\beta}^{h,j}_F\}$ to the structural parameters of the transition matrix:
    \begin{equation}
        \hat{\alpha}_S, \quad \hat{\alpha}^j_F, \quad \hat{\kappa}^j_F, \quad \hat{\phi}_S, \quad \hat{\delta}_\theta
    \end{equation}
    
    \item \textbf{Simulation:} Solve the calibrated LQR backward (Riccati recursion) and simulate forward under alternative policy paths—e.g., ``no fiscal support'' ($F_k \equiv 0$) or ``delayed lockdown'' ($S_k = 0$ for $k < k^*$).
    
    \item \textbf{Comparison:} Compare simulated trajectories to observed outcomes, assessing model fit and quantifying the welfare implications of counterfactual policies.
\end{enumerate}

This exercise provides a direct test of the theoretical framework: if the LQR model captures the essential dynamics, calibrated simulations should approximate the empirical impulse responses.

\subsection{Limitations}

Several limitations warrant acknowledgment. First, the sample comprises 38 countries over approximately 8 quarters, constraining statistical power for detecting heterogeneous effects across fiscal instruments. Second, the identification strategy relies on the policy reaction functions capturing the systematic component of government decisions; omitted variables correlated with both policy choices and outcomes would bias the estimates. Third, the linear specification may inadequately capture threshold effects or nonlinearities in policy transmission. Fourth, cross-country comparability of fiscal measures—despite careful harmonisation—remains imperfect due to institutional heterogeneity.

These limitations notwithstanding, the combination of granular fiscal data, theory-guided specification, and robust inference methods provides a credible foundation for testing the theoretical predictions.

\bibliographystyle{apalike}
\begin{thebibliography}{99}

\bibitem[Hale et al., 2021]{hale2021global}
Hale, T., Angrist, N., Goldszmidt, R., et al. (2021).
\newblock A global panel database of pandemic policies (Oxford COVID-19 Government Response Tracker).
\newblock \textit{Nature Human Behaviour}, 5(4), 529--538.

\bibitem[Jordà, 2005]{jorda2005estimation}
Jordà, Ò. (2005).
\newblock Estimation and inference of impulse responses by local projections.
\newblock \textit{American Economic Review}, 95(1), 161--182.

\bibitem[Jordà \& Taylor, 2025]{jorda2025local}
Jordà, Ò. \& Taylor, A. M. (2025).
\newblock Local projections.
\newblock \textit{Journal of Economic Literature}, 63(1), 59--110.

\bibitem[Pesenti, 2025]{pesenti2025codebook}
Pesenti, A. (2025).
\newblock COVID-19 Fiscal Response Database: Codebook Version 1.3.
\newblock Working Paper, University of Basel.

\bibitem[Webb, 2023]{webb2023wild}
Webb, M. D. (2023).
\newblock Reworking wild bootstrap-based inference for clustered errors.
\newblock \textit{Canadian Journal of Economics}, 56(3), 839--858.

\end{thebibliography}


\section{Expected Effects: Evidence from the Empirical Literature}
\label{sec:expected_effects}

Before presenting the estimation results, this section synthesises the existing empirical evidence on the causal effects of containment and fiscal policies during the COVID-19 pandemic. This literature review serves two purposes: first, it provides theoretical priors for the expected signs and magnitudes of our estimated coefficients; second, it identifies gaps in the existing research that motivate the contribution of this study.

\subsection{Effects of Containment Measures on Economic Activity}
\label{subsec:stringency_output}

The contractionary effect of non-pharmaceutical interventions (NPIs) on economic output is well-documented in the empirical literature. Cross-country evidence from the OECD suggests that stringent lockdown measures contributed to GDP reductions of 20--25\% in several member countries during the initial phase of the pandemic \citep{OECD2020}.

Panel regression analyses covering approximately 90 countries over the period 2020Q1--2021Q4 confirm that changes in the stringency index were a significant driver of GDP dynamics \citep{Lagakos2021}. However, the magnitude of the economic impact exhibited substantial heterogeneity across income groups. Emerging market and developing economies experienced larger contractions than advanced economies for comparable levels of containment stringency. This differential response can be attributed to structural factors: higher employment shares in contact-intensive sectors, weaker automatic stabilisers, and more limited access to public insurance mechanisms \citep{Lagakos2021, OECD2020}.

Within-country time-series variation reveals a characteristic temporal pattern. Restrictive measures are associated with lower GDP growth in the contemporaneous quarter, followed by a positive catch-up effect in the subsequent quarter as economic activity rebounds upon relaxation of restrictions \citep{Deb2021}. This mean-reverting dynamic aligns with the autoregressive structure embedded in the theoretical model through the parameter $\rho_y$.

For the empirical analysis, we therefore expect:
\begin{equation}
    \beta^h_S < 0 \quad \text{for } h = 0,1
\end{equation}
with the largest contractionary effect at impact ($h=0$) and gradual attenuation over subsequent horizons.

\subsection{Effects of Containment Measures on Mortality}
\label{subsec:stringency_mortality}

The relationship between containment stringency and excess mortality operates through an epidemiological transmission channel, generating inherent policy lags between intervention and health outcomes.

Causal evidence from natural experiments provides compelling support for the protective effect of lockdowns. Using synthetic control methods to construct a counterfactual for Sweden---which notably did not implement mandatory lockdowns during the first wave---\citet{Yarmol2023} estimate that up to 4,411 of the 46,554 deaths registered between February and September 2020 could have been avoided under alternative policy choices. This represents approximately 9.5\% of total mortality during the period.

Cross-country panel analyses confirm the importance of containment timing and intensity. The stringency index emerges as the most significant predictor of both infection rates and mortality outcomes, with standardised coefficients ranging from 0.588 to 0.702 across alternative specifications \citep{Castex2021}. The timing of lockdown implementation relative to initial community transmission significantly affects peak infection levels, which in turn drives both COVID-19 deaths and healthcare system pressure through intensive care unit overcrowding \citep{Flaxman2020}.

Importantly, the model structure captures the biological lag between policy intervention and mortality outcomes. Social distancing measures ($S_k$) suppress viral prevalence ($\theta_{k+1}$), which subsequently determines mortality ($d_{k+2}$). This two-period lag structure implies:
\begin{equation}
    \beta^h_S \approx 0 \quad \text{for } h = 0, \qquad \beta^h_S < 0 \quad \text{for } h \geq 1
\end{equation}

A potential countervailing effect operates through non-COVID excess mortality. Stringent containment measures may increase deaths from delayed medical care, mental health deterioration, and economic deprivation \citep{Kontopantelis2021}. However, offsetting factors---including reduced pollution, fewer traffic accidents, and lower workplace injuries---suggest ambiguous net effects on total excess mortality beyond COVID-19.

\subsection{Effects of Fiscal Support on Economic Activity}
\label{subsec:fiscal_output}

The effectiveness of fiscal stimulus during the pandemic has been subject to extensive empirical investigation, yielding several robust findings regarding instrument heterogeneity and state dependence.

\subsubsection{Baseline Multiplier Estimates}

Pre-pandemic evidence from the American Recovery and Reinvestment Act (ARRA) suggested government expenditure multipliers of approximately 1.5 when exploiting cross-geographical variation \citep{ChodorowReich2019}. However, the pandemic context introduced novel features that complicate direct extrapolation of historical estimates.

Panel evidence from advanced economies indicates that cumulative fiscal multipliers one year after a health crisis are approximately twice as large as during normal recessions \citep{IMF2022}. This elevated effectiveness reflects the confluence of substantial economic slack and suppressed private demand. Country-specific analyses from Central and Eastern European economies during COVID-19 report within-quarter multipliers ranging from 0.03 to 0.61, with medium-term cumulative effects approaching or exceeding unity \citep{Haug2021}.

\subsubsection{Heterogeneity by Instrument Design}

The theoretical framework distinguishes fiscal instruments along two dimensions: design (expenditure versus revenue) and target (households versus firms). The empirical literature provides guidance on the expected multiplier hierarchy.

\paragraph{Expenditure versus Revenue Measures.}
A robust finding across multiple methodological approaches is that expenditure-based instruments generate larger output effects than revenue-based measures \citep{Blanchard2013, Ramey2019}. Tax-based fiscal adjustments are consistently found to be more contractionary (or less expansionary) than equivalent expenditure changes. This differential reflects the more immediate demand injection from direct government spending compared to the indirect transmission through household budget constraints and behavioural responses to tax changes.

\paragraph{Household versus Firm Targeting.}
Pre-pandemic theory predicted higher multipliers for transfers to liquidity-constrained households, who exhibit elevated marginal propensities to consume (MPC). However, pandemic-era evidence challenges this prediction. Survey data reveal that U.S.\ households saved the majority of their stimulus payments, with realised MPCs substantially below those observed following the 2008 stimulus \citep{Coibion2020}. This behavioural response reflects both supply-side consumption constraints (closed retail establishments) and elevated precautionary savings in an uncertain environment \citep{Chetty2020}.

Firm-targeted support, particularly job retention schemes and payroll subsidies, preserved employment relationships and facilitated rapid recovery upon reopening \citep{Giupponi2022}. However, programmes with poor targeting---such as the U.S.\ Paycheck Protection Program, where approximately three-quarters of disbursements accrued to top-quintile households---generated limited stabilisation despite substantial fiscal costs \citep{Autor2022}.

Based on this evidence, we expect the following multiplier hierarchy:
\begin{equation}
    \alpha^{Exp,HH}_F > \alpha^{Exp,Firm}_F > \alpha^{Rev,HH}_F > \alpha^{Rev,Firm}_F
\end{equation}
with absolute magnitudes potentially compressed under strict containment (see below).

\subsection{Interaction Effects: Fiscal Effectiveness under Containment}
\label{subsec:interaction}

A central contribution of this study is the joint estimation of containment and fiscal policy effects, including their interaction. The empirical literature provides strong priors for negative interaction effects---that is, dampened fiscal multipliers under strict containment.

\citet{Auerbach2022} exploit high-frequency variation in U.S.\ Department of Defence spending and local stay-at-home orders during early 2020. Their central finding is striking: DOD spending increased employment only in cities without meaningful lockdown restrictions. In unrestricted cities, the cost per job created was approximately \$4,000 per month; in restricted cities, the employment effect was statistically indistinguishable from zero.

Importantly, the mechanism analysis reveals that even in unrestricted locations, DOD spending did not stimulate local consumption as measured by credit card data \citep{Chetty2020}. This suggests that the standard high-MPC transmission channel was inoperative during the pandemic. The employment effects in unrestricted areas operated through a ``slack channel''---directly absorbing idle labour---rather than through induced consumption.

This pattern has broader implications for fiscal effectiveness under supply constraints. When containment measures restrict the supply of goods and services available for consumption, demand-side stimulus generates smaller output effects. Theoretical models predict that heightened uncertainty and supply bottlenecks both dampen the transmission of fiscal impulses \citep{Guerrieri2022}.

The interaction term in our specification captures this state dependence:
\begin{equation}
    \beta^{h,j}_{F \times S} < 0
\end{equation}
implying that higher stringency reduces the effectiveness of fiscal support across all instrument categories.

\subsection{Effects of Fiscal Support on Public Debt}
\label{subsec:fiscal_debt}

The fiscal response to the pandemic generated unprecedented increases in public debt across OECD economies. From the first to the second quarter of 2020, EU government debt increased by 8.4 percentage points to 88\% of GDP, reaching 94\% by year-end \citep{Eurostat2021}. National fiscal packages ranged from 3.1\% of GDP (Romania) to 10.5\% of GDP (Hungary) in Central and Eastern European countries, with additional health expenditures averaging 2.1\% of GDP \citep{Haug2021}.

The debt impact varies by instrument type. Above-the-line measures (direct spending and tax reductions) generate immediate increases in the debt-to-GDP ratio. Below-the-line measures (guarantees and equity injections) create contingent liabilities with delayed and partial realisation. Tax deferrals represent timing shifts that are partially reversible as deferred payments come due.

An important feedback mechanism operates from debt to fiscal effectiveness. High existing debt levels blunt the impact of additional stimulus through several channels: crowding out of private investment, higher sovereign risk premia, and expectations of future fiscal consolidation \citep{Ilzetzki2013}. Estimates suggest multipliers of 0.6--1.0 in low-debt countries but below 0.4---or even negative over longer horizons---in high-debt countries.

For the empirical specification, we expect:
\begin{equation}
    \beta^{h,j}_{F,b} > 0
\end{equation}
with heterogeneity across instrument types reflecting differential above/below-the-line treatment.

\subsection{Summary of Expected Coefficient Signs}
\label{subsec:summary_priors}

Table \ref{tab:expected_signs} summarises the prior expectations derived from the empirical literature. These priors inform the interpretation of estimated impulse responses and provide benchmarks for assessing the plausibility of results.

\begin{table}[htbp]
\centering
\caption{Expected Coefficient Signs Based on Prior Literature}
\label{tab:expected_signs}
\begin{tabular}{@{}llccc@{}}
\toprule
Coefficient & Outcome & Expected Sign & Confidence & Magnitude \\
\midrule
$\beta^h_S$ & Output gap ($y$) & $-$ & High & Large at $h=0$ \\
$\beta^h_S$ & Excess mortality ($d$) & $-$ (lagged) & Medium--High & $h \geq 1$ \\
$\beta^{h,Exp\text{-}HH}_F$ & Output gap ($y$) & $+$ & Medium & $\sim$1.0--1.5 \\
$\beta^{h,Exp\text{-}Firm}_F$ & Output gap ($y$) & $+$ & Medium & $\sim$0.5--1.0 \\
$\beta^{h,Rev\text{-}HH}_F$ & Output gap ($y$) & $+$ & Medium & $\sim$0.3--0.8 \\
$\beta^{h,Rev\text{-}Firm}_F$ & Output gap ($y$) & $+$ & Medium & $<$0.5 \\
$\beta^{h,j}_F$ & Public debt ($b$) & $+$ & High & Pass-through \\
$\beta^{h,j}_{F \times S}$ & Output gap ($y$) & $-$ & Medium & To zero under strict lockdown \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Contribution Relative to Existing Literature}
\label{subsec:contribution}

The empirical analysis contributes to the existing literature along four dimensions. First, the simultaneous estimation of containment and fiscal effects---including their interaction---addresses the joint determination that pervades crisis policymaking but is rarely modelled explicitly. Second, the disaggregation of fiscal instruments by design and target enables heterogeneity analysis that moves beyond aggregate fiscal impulse measures. Third, the inclusion of public debt as an outcome variable extends the standard output-focused analysis to capture the intertemporal fiscal trade-offs central to the theoretical model. Fourth, the two-stage identification strategy using policy reaction function residuals addresses the endogeneity concerns that plague reduced-form estimates of policy effects.


\section{LP Empirical Design}
\section{Empirical Strategy}

To quantify the dynamic causal effects of government pandemic responses and fiscal support on macroeconomic and health outcomes, we employ the local projections (LP) method introduced by \cite{Jorda2005} and recently synthesized by \cite{JordaTaylor2025}. This approach is particularly well-suited for our panel setting as it is more robust to misspecification than vector autoregressions (VARs) and allows for a flexible estimation of impulse response functions (IRFs) in the presence of multiple exogenous shocks.

\subsection{Econometric Specification}

We estimate the cumulative impulse response functions for each forecast horizon $h = 0, \dots, H$ using a sequence of panel regressions. The baseline specification for country $i$ at time $t$ is given by:

\begin{equation} \label{eq:lp_baseline}
\Delta_h y_{i,t+h} = \alpha_i^h + \delta_t^h + \beta_h^S S_{i,t} + \beta_h^F F_{i,t} + \sum_{l=1}^{L} \mathbf{\Gamma}_l^h \mathbf{W}_{i,t-l} + \epsilon_{i,t+h}
\end{equation}

\noindent where $\Delta_h y_{i,t+h}$ denotes the cumulative change in the outcome variable of interest from $t-1$ to $t+h$. For real GDP, we define the outcome as the log-difference, $y_{i,t+h} - y_{i,t-1}$, approximating the cumulative percentage growth relative to the pre-shock level. For government debt, we use the simple difference in the debt-to-GDP ratio. For excess deaths, we follow the hybrid approach proposed by \cite{JordaNechio2023}: we calculate the gap between observed deaths and a counterfactual linear trend estimated exclusively on pre-pandemic data (2015--2019) and extrapolated through the forecast horizon. This avoids the "over-smoothing" bias that would arise from applying a Hodrick-Prescott filter across the structural break of the pandemic.

The key independent variables are $S_{i,t}$, our continuous measure of government response intensity (constructed as a residual from the first-stage estimation), and $F_{i,t}$, the discretionary fiscal support. Since $S_{i,t}$ and $F_{i,t}$ are uncorrelated regarding magnitude and timing, we include them simultaneously in the projection. The coefficients $\beta_h^S$ and $\beta_h^F$ are the parameters of interest, tracing out the impulse response of the outcome at horizon $h$ to a unit shock in the respective policy variable, holding the other policy dimension constant.

The specification includes country fixed effects ($\alpha_i^h$) to control for time-invariant heterogeneity and time fixed effects ($\delta_t^h$) to absorb common global shocks, such as the global progression of the virus or global financial conditions. The vector $\mathbf{W}_{i,t-l}$ contains $L$ lags of the dependent variable and both shock variables to control for serial correlation and pre-existing trends. Following the best practices outlined in \cite{JordaTaylor2025}, the lag length $L$ is selected using the corrected Akaike Information Criterion ($AIC_c$) to balance bias and variance given our finite sample size ($T \approx 40$).

\subsection{Identification}

Identification relies on the assumption that, conditional on the rich set of fixed effects and lagged controls, the shocks $S_{i,t}$ and $F_{i,t}$ are orthogonal to contemporaneous innovations in the error term $\epsilon_{i,t+h}$. 

For the government response variable $S_{i,t}$, exogeneity is achieved by construction in the first stage, where we orthogonalize the observed response against a set of observable epidemiological and economic fundamentals. Consequently, $S_{i,t}$ represents the "surprise" or discretionary component of the government's reaction function. For fiscal support $F_{i,t}$, we rely on the assumption that the precise timing and magnitude of fiscal packages are determined by political processes that are predetermined relative to the high-frequency variation in quarterly outcomes, conditional on the severity of the pandemic (which is controlled for via time fixed effects and lags).

\subsection{Inference}

Correct statistical inference is critical given two features of our design: (i) the panel structure with potential cross-sectional dependence due to the synchronized nature of the pandemic, and (ii) the fact that $S_{i,t}$ is a generated regressor. Standard errors that ignore the estimation uncertainty of the first stage would be downward biased.

To address both issues, we implement a \textbf{cross-sectional block bootstrap}, as recommended by \cite{JordaTaylor2025} for panel local projections. The procedure is as follows:
\begin{enumerate}
    \item We resample $N$ countries with replacement from the original dataset. For each sampled country, we retain the entire time series (the "block") to preserve the internal temporal dependence structure.
    \item For each bootstrap replication $b = 1, \dots, B$, we re-estimate the entire two-step procedure: we first re-calculate the residuals $S_{i,t}^b$ and then estimate the local projections (Equation \ref{eq:lp_baseline}) using these re-generated shocks.
    \item We compute the point estimates and confidence intervals based on the distribution of the $B$ bootstrap estimates.
\end{enumerate}
This bootstrapping scheme automatically accounts for the generated regressor problem and is robust to serial correlation and cross-sectional dependence within the panel. As a robustness check, we also report analytical standard errors clustered by country and corrected for autocorrelation (Driscoll-Kraay), which yield qualitatively similar results.

\newpage
\FloatBarrier
\bibliographystyle{apa}
\bibliography{references}

\newpage
\section{Appendix}

\appendix

\subsection{Structured Comparison}
\label{app:structured_comp}

\begin{small}
\begin{longtable}{P{3.2cm}P{3cm}P{3cm}P{3cm}P{3cm}}
\toprule
\textbf{Dimension} & \textbf{Alvarez et al.\ (2021)} & \textbf{Eichenbaum et al.\ (2021)} & \textbf{Kaplan et al.\ (2020)} & \textbf{Pesenti (2026)} \\
\midrule
\endfirsthead
\toprule
\textbf{Dimension} & \textbf{Alvarez} & \textbf{Eichenbaum} & \textbf{Kaplan} & \textbf{Pesenti} \\
\midrule
\endhead

\textbf{Epidemiology} & Continuous SIR & Discrete SIR & Expanded SIR with occupational contact rates & Reduced-form SIR ($\rho_\theta$, $\delta_\theta$, $\phi_S$) \\
\midrule

\textbf{Economic model} & Linear production & NK--DSGE, representative agent & HANK with income/ wealth inequality & Linear-quadratic state-space \\
\midrule

\textbf{Agent structure} & Representative & Representative & Heterogeneous (income, wealth, occupation, sector) & Representative planner \\
\midrule

\textbf{Containment instrument} & Lockdown $L_t \in [0,1]$ & Pigouvian consumption tax $\mu_t$ & Sector-specific lockdown & $S_k \in [0,1]$ with $\phi_S$ \\
\midrule

\textbf{Fiscal instruments} & \textbf{None} & \textbf{None} & Lump-sum transfers, UI extension & $F_k^{DI}$ (demand injection) + $F_k^{CP}$ (capacity preservation) \\
\midrule

\textbf{Instrument disaggregation} & n/a & n/a & Partial (transfers vs.\ UI) & Full (DI vs.\ CP by transmission mechanism) \\
\midrule

\textbf{Debt / fiscal sustainability} & \textbf{No} & \textbf{No} & Present but not optimized & \textbf{Core objective:} $b_k$ is penalized and endogenous \\
\midrule

\textbf{Automatic stabilizers} & No & No & Implicit (progressive taxes in HANK) & Explicit: $\gamma_y$ creates endogenous fiscal feedback \\
\midrule

\textbf{Output persistence / hysteresis} & No (linear, instant recovery) & No (equilibrium recovery) & Partial (wealth effects, scarring through inequality) & Explicit: $\rho_y^{\text{eff}}(S_k, F_k^{CP})$ \\
\midrule

\textbf{Optimization method} & Pontryagin maximum principle (continuous) & Competitive equilibrium + Ramsey problem & Calibrated scenarios (no joint optimization) & iterative LQR (discrete-time, Riccati) \\
\midrule

\textbf{State vector} & $(S, I, R, D)$ & $(S, I, R, D)$ + macro & $(S, I, R, D)$ + full HANK distribution & $(y_k, d_k, b_k, \theta_k)$ \\
\midrule

\textbf{Control vector} & $L_t$ (scalar) & $\mu_t$ (scalar) & Scenarios (not optimized) & $(S_k, F_k^{DI}, F_k^{CP})$ (3-dim) \\
\midrule

\textbf{Objective function} & $\min \int [w L + \theta \dot{D}]\,dt$ (2 terms) & Social welfare (consumption utility) & Distributional welfare & $\min \sum \beta^k [w_y y_k^2 + w_d d_k^2 + w_b b_k^2 + R \|u_k\|^2]$ (3 state penalties) \\
\midrule

\textbf{Trade-offs captured} & Health $\leftrightarrow$ Economy (2D) & Health $\leftrightarrow$ Economy (2D) & Health $\leftrightarrow$ Economy $\leftrightarrow$ Distribution (2.5D) & Health $\leftrightarrow$ Economy $\leftrightarrow$ Fiscal (3D trilemma) \\
\midrule

\textbf{Empirical calibration} & U.S.\ parameters, COVID-19 & U.S.\ parameters, COVID-19 & U.S.\ calibration, HANK & 38 OECD countries, LP estimation, micro database \\
\midrule

\textbf{Cross-country} & No & No & No & Yes (38 OECD) \\
\midrule

\textbf{Counterfactual exercises} & Optimal vs.\ no lockdown & CE vs.\ optimal containment & Scenario comparison & Optimal composition vs.\ observed; joint vs.\ conditional optimum \\
\bottomrule
\label{tab:structured_comp}
\end{longtable}
\end{small}




%% ============================================================
\subsection{External Validation: The Trilemma in Comparative Perspective}
\label{subsec:comparison}
%% ============================================================

To establish that the pandemic trilemma represents a structurally unique configuration rather than a generic feature of crisis management, I apply the four constitutive properties as a diagnostic framework to four alternative crisis types: armed conflict, natural disasters, financial crises, and climate change. The objective is to determine whether the specific combination of properties---and their circular interdependence---arises in other contexts. Table~\ref{tab:comparison} summarizes the results.

%% --- War ---
\subsubsection*{Armed Conflict}

Armed conflict shares certain surface-level features with the pandemic: it can involve escalatory dynamics and generates fiscal pressures through military expenditure, activating both the debt accumulation channel (Property~3) and automatic stabilizers (Property~4). However, Property~1 is only partially satisfied. While conflict can escalate, the threat does not exhibit the autonomous, biological growth dynamic of $\rho_\theta > 1$: military escalation is mediated by strategic decisions, not by an epidemiological reproduction rate that compounds independently of human choice.

The critical distinction lies in Property~2. Military mobilization and defense spending are typically \textit{expansionary}: wartime production absorbs idle resources, stimulates aggregate demand, and increases employment. The instrument that suppresses the threat operates in the \textit{same} direction as economic stabilization rather than against it---violating the structural requirement that $\alpha_S > 0$ and $\phi_S > 0$ operate through the same policy variable. Furthermore, the nature of the shock differs fundamentally: war destroys physical capital ($K$), whereas a pandemic paralyzes economic flows while leaving the capital stock intact. The reconstruction problem following a war---rebuilding $K$ toward its pre-war level---is qualitatively different from the restoration problem following a pandemic, where the objective is to return the flow variables $(y_k, \theta_k)$ to their pre-crisis steady state. Because the suppression instrument is not inherently contractionary, the circular linkage that generates the trilemma does not arise.

%% --- Natural Disaster ---
\subsubsection*{Natural Disasters}

Natural disasters constitute the purest form of exogenous shock, but they violate Property~1 in a decisive respect: they are \textit{impulse} shocks with $\rho < 1$. An earthquake or flood strikes, causes damage, and ceases. The shock does not grow endogenously; it dissipates without intervention. Consequently, there is no necessity for a suppression instrument: the planner faces a reconstruction problem, not a containment problem. Property~2 is therefore vacuous---since no active suppression is required, the question of whether suppression is contractionary does not arise. While the fiscal pressures from reconstruction can be substantial (Properties~3 and~4 are present), the absence of the self-reinforcing dynamic ($\rho_\theta > 1$) and the contractionary suppression instrument means that the circular linkage cannot form. The planner faces a costly rebuilding challenge, but not a trilemma.

%% --- Financial Crisis ---
\subsubsection*{Financial Crises}

Financial crises violate Property~1 at its foundation: the shock is \textit{endogenous} to the economic system. It originates within the financial sector through excessive leverage, asset price misalignments, or institutional failures. Even the international propagation that characterized the 2008 crisis---transmission through interbank lending channels and cross-border contagion---represents the spread of endogenous instabilities through system-internal linkages rather than the arrival of an external shock ($\varepsilon_k$).

Property~2 is violated in a structural sense. The primary intervention instruments---bank recapitalization, liquidity provision, and asset purchases---are \textit{expansionary} for the real economy, or at minimum prevent a contractionary collapse. This reverses the causal logic of the pandemic trilemma: in a pandemic, the suppression of the shock \textit{itself} is contractionary ($\alpha_S > 0$), whereas in financial crises, contractionary outcomes arise from the \textit{failure} to intervene or from subsequent fiscal consolidation. Austerity policies following the European sovereign debt crisis were contractionary, but they constituted a response to the fiscal legacy of the crisis, not the crisis-fighting instrument per se. The trilemma structure does not emerge because the instrument that addresses the threat works \textit{with} rather than \textit{against} economic stabilization.

%% --- Climate Change ---
\subsubsection*{Climate Change}

Climate change is the most structurally analogous case and merits extended discussion.

Property~1 is satisfied: greenhouse gas accumulation exhibits self-reinforcing dynamics analogous to $\rho_\theta > 1$. CO$_2$ concentrations compound in the atmosphere, positive feedback loops (permafrost thawing, ice-albedo effects) accelerate warming, and potential tipping points imply that the process may become irreversible beyond certain thresholds. The lethality channel ($\delta_\theta > 0$) operates through climate-related mortality: heat stress, extreme weather events, and food system disruptions.

Property~2 is satisfied: the suppression instruments---carbon taxation, emission standards, and the phase-out of fossil fuel infrastructure---are contractionary for carbon-intensive sectors and, at least transitionally, for aggregate output. As in the pandemic, the mechanism that suppresses the threat simultaneously restricts economic activity ($\alpha_S > 0$).

Property~3 is satisfied: the transition to a low-carbon economy requires substantial public investment in infrastructure, technology, and social compensation programs, generating intertemporal fiscal liabilities ($\kappa_F > 0$).

Property~4 is satisfied: an economic slowdown induced by climate policy activates the same automatic stabilizer channel ($-\gamma_y \, y_k$) as a pandemic-induced contraction.

All four constitutive properties are present, and a circular linkage analogous to the pandemic trilemma can be identified: emission reduction contracts the economy, requiring fiscal compensation, which generates debt, which constrains future fiscal space for the transition. However, three qualitative differences distinguish the climate case and preclude a direct application of the framework developed here:

\begin{enumerate}
    \item \textbf{Time scale.} The pandemic trilemma operates on a horizon of quarters to years; the climate trilemma on decades. This transforms the discount factor $\beta$ from a technical parameter into a normative choice about intergenerational equity, fundamentally altering the political economy of the optimization problem.

    \item \textbf{Observability and attribution.} Pandemic mortality is immediate, individually attributable, and politically salient---the lag from $S_k$ to $d_{k+2}$ spans weeks. Climate-related mortality is diffuse, statistically distributed across populations, and temporally remote. The policy lag between suppression and observable health outcomes is orders of magnitude longer, rendering the political economy of sustained suppression qualitatively more difficult.

    \item \textbf{Terminal condition.} A pandemic admits a finite planning horizon: vaccination exogenously decouples mortality from infection (reducing $\delta_\theta$ and $\rho_\theta$), dissolving the trilemma at the terminal date and leaving only the fiscal legacy. Climate change admits no analogous resolution; the suppression effort must be sustained indefinitely. The terminal condition that resolves the pandemic trilemma---the exogenous dissolution of the epidemiological constraint through vaccination, as discussed in Property~2---has no climate analogue.
\end{enumerate}

\noindent These differences do not invalidate the structural analogy but imply that a climate trilemma framework would require fundamentally different modeling choices regarding the planning horizon, the discount structure, and the terminal condition. The pandemic setting therefore offers a more tractable environment in which to develop and test the trilemma framework, with potential for future extension to the climate domain.

%% --- Summary Table ---
% Uncomment and adapt when ready
\begin{table}[h]
\centering
\caption{Diagnostic Comparison of Crisis Types}
\label{tab:comparison}
\begin{tabular}{lcccc}
\hline
 & Property~1 & Property~2 & Property~3 & Property~4 \\
 & Exogenous, $\rho > 1$ & Contractionary & Debt-generating & Auto. Stabilizers \\
\hline
Pandemic          & \checkmark & \checkmark & \checkmark & \checkmark \\
Armed Conflict    & Partial    & \texttimes & \checkmark & \checkmark \\
Natural Disaster  & \texttimes & Vacuous    & \checkmark & \checkmark \\
Financial Crisis  & \texttimes & \texttimes & \checkmark & \checkmark \\
Climate Change    & \checkmark & \checkmark & \checkmark & \checkmark \\
\hline
\end{tabular}
\end{table}


\newpage
%%Old Version
%% --- Summary Table ---
\subsubsection*{Comparative Summary}

Table~\ref{tab:comparison} synthesises the diagnostic assessment across crisis types.

\begin{table}[ht]
\centering
\caption{Constitutive Properties of the Trilemma Across Crisis Types}
\label{tab:comparison}
\renewcommand{\arraystretch}{1.3}
\small
\begin{tabular}{@{} p{5.5cm} c c c c c @{}}
\toprule
\textbf{Property} & \textbf{Pandemic} & \textbf{War} & \textbf{Nat.\ Dis.} & \textbf{Financial} & \textbf{Climate} \\
\midrule
1. Self-reinforcing exogenous shock    & \checkmark & $\sim$     & $\times$   & $\times$   & \checkmark \\[2pt]
2. Contractionary suppression instr.          & \checkmark & $\times$   & n/a        & $\times$   & \checkmark \\[2pt]
3. Intertemporal fiscal liability      & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark \\[2pt]
4. Endogenous fiscal feedback          & \checkmark & $\sim$     & $\sim$     & \checkmark & \checkmark \\[2pt]
5. Irreducible policy lag              & \checkmark & $\times$   & $\times$   & $\sim$     & \checkmark \\
\midrule
\textbf{Trilemma structure}            & \textbf{Yes} & \textbf{No} & \textbf{No} & \textbf{No} & \textbf{Potentially} \\
\bottomrule
\end{tabular}

\vspace{0.5em}
\footnotesize
\textit{Notes:} \checkmark\, = property satisfied; $\times$\, = property not satisfied; $\sim$\, = partially satisfied; n/a = not applicable. The trilemma requires the simultaneous presence of all five properties \textit{and} their circular interdependence.
\end{table}

The analysis confirms that the pandemic trilemma is structurally unique among the crisis types considered. Climate change is the only context in which an analogous trilemma structure may emerge, although the fundamental differences in time scale, observability, and reversibility imply that the climate trilemma---should it be formalised---would require substantial modifications to the analytical framework.


\section*{Derivation of Policy Necessity and Economic Impact}
\label{app:Theta import}

Due to the transmission lag, immediate mortality is predetermined. The planner controls future mortality $d_{k+2}$ by suppressing current viral transmission. Based on the transmission dynamics defined in Eq.~(\ref{eq:trans_dynamics}), the causal chain is:
\begin{equation}
    d_{k+2} = \delta_\theta \cdot \underbrace{\rho_\theta (1 - \phi_S S_k) \theta_k}_{\theta_{k+1}}
    \label{eq:excess dynamics}
\end{equation}
Consequently, the necessity of the policy instrument $S_k$ depends on the conjunction of the lethality parameter $\delta_\theta$ and the viral growth potential $\rho_\theta$.

Let economic output $Y_k$ be a decreasing function of suppression $S_k$. The policy constraint on output is strictly binding if and only if the virus is lethal ($\delta_\theta > 0$) and transmissible ($\rho_\theta > 0$).


\begin{proof}
    We analyze the marginal effect of the suppression policy $S_k$ on the health outcome $d_{k+2}$. Differentiating Eq.~(\ref{eq:excess dynamics}) with respect to $S_k$ yields:
    \begin{equation}
        \frac{\partial d_{k+2}}{\partial S_k} = \delta_\theta \cdot \rho_\theta \cdot (-\phi_S) \cdot \theta_k = - \delta_\theta \rho_\theta \phi_S \theta_k
    \end{equation}
    The marginal health benefit of suppression (reducing deaths) is non-zero if and only if the derivative is non-zero.
    \begin{itemize}
        \item \textbf{Case $\delta_\theta = 0$ (Non-lethal):} The derivative becomes zero. Suppression saves no lives. The optimal strategy is $S_k=0$ to maximize output.
        \item \textbf{Case $\delta_\theta > 0$ (Lethal):} Given an active epidemic ($\theta_k > 0$) and effective NPIs ($\phi_S > 0$), the derivative is strictly negative. Increasing $S_k$ significantly reduces future mortality.
    \end{itemize}
    Therefore, the planner is forced to accept economic costs (set $S_k > 0$) solely because $\delta_\theta > 0$. The trade-off is conditional on the simultaneous existence of lethality and transmissibility.
\end{proof}

\section*{Analyticial Analysis: A.1 System Definition}

The nonlinear discrete-time system is defined by the state vector $x_k = (y_k, d_k, b_k, \theta_k)'$ and the control vector $u_k = (S_k, F_k^{DI}, F_k^{CP})'$. The transition equations are:
%
\begin{align}
    y_{k+1} &= \left[\rho_y + \psi \, S_k - \eta \, F_k^{CP}\right] y_k
      - \alpha_S \, S_k + \alpha_F^{DI} \, F_k^{DI}
    \label{eq:y_trans} \\[6pt]
    d_{k+1} &= \delta_\theta \, \theta_k
    \label{eq:d_trans} \\[6pt]
    b_{k+1} &= (1+r) \, b_k - \gamma_y \, y_k
      + \kappa_F^{DI} \, F_k^{DI} + \kappa_F^{CP} \, F_k^{CP} + c_k^{H}
    \label{eq:b_trans} \\[6pt]
    \theta_{k+1} &= \rho_\theta (1 - \phi_S \, S_k) \, \theta_k + \varepsilon_k
    \label{eq:theta_trans}
\end{align}
%
with exogenous health expenditures $c_k^H$ and stochastic shock $\varepsilon_k$ with $E[\varepsilon_k] = 0$.

\subsection*{A.2 Jacobian Matrices}

The linearization around a trajectory $(\bar{x}_k, \bar{u}_k)$ yields the time-varying system $\delta x_{k+1} \approx A_k \, \delta x_k + B_k \, \delta u_k$, where:
%
\begin{equation}
    A_k = \frac{\partial f}{\partial x_k} \bigg|_{(\bar{x}_k, \bar{u}_k)} =
    \begin{pmatrix}
        \rho_y + \psi \bar{S}_k - \eta \bar{F}_k^{CP} & 0 & 0 & 0 \\
        0 & 0 & 0 & \delta_\theta \\
        -\gamma_y & 0 & (1+r) & 0 \\
        0 & 0 & 0 & \rho_\theta(1 - \phi_S \bar{S}_k)
    \end{pmatrix}
    \label{eq:A_k}
\end{equation}
%
\begin{equation}
    B_k = \frac{\partial f}{\partial u_k} \bigg|_{(\bar{x}_k, \bar{u}_k)} =
    \begin{pmatrix}
        \psi \bar{y}_k - \alpha_S & \alpha_F^{DI} & -\eta \bar{y}_k \\
        0 & 0 & 0 \\
        0 & \kappa_F^{DI} & \kappa_F^{CP} \\
        -\phi_S \rho_\theta \bar{\theta}_k & 0 & 0
    \end{pmatrix}
    \label{eq:B_k}
\end{equation}

\paragraph{Rank and controllability.} Since $\text{rank}(B_k) = 3 < 4 = \dim(x)$, the mortality state $d_k$ is not directly controllable. It is reachable through the chain $S_k \to \theta_{k+1} \to d_{k+2}$ via the compound $A_k B_k$, which has a nonzero entry in the second row equal to $-\delta_\theta \phi_S \rho_\theta \bar{\theta}_k$. The full controllability matrix $\mathcal{C} = [B_k \; A_k B_k \; A_k^2 B_k \; A_k^3 B_k]$ has full row rank under the non-degeneracy conditions $\delta_\theta, \phi_S, \rho_\theta, \alpha_S, \alpha_F^{DI}, \kappa_F^{DI}, \theta_k \neq 0$.

% ============================================================
\subsection*{A.3 Three-Period Forward Iteration}
% ============================================================

We initialize at the pre-pandemic steady state $x_0 = (0, 0, 0, \theta_0)'$, where $\theta_0 > 0$ represents the initial outbreak. We iterate forward deterministically ($\varepsilon_k = 0$) for $k = 0, 1, 2$ and suppress $c_k^H$ for clarity (it enters additively and does not interact with any control).

\subsubsection*{Period 0 $\to$ 1}

From $y_0 = 0$, $b_0 = 0$:
%
\begin{align}
    y_1 &= -\alpha_S \, S_0 + \alpha_F^{DI} \, F_0^{DI}
    \label{eq:y1} \\[4pt]
    d_1 &= \delta_\theta \, \theta_0
    \label{eq:d1} \\[4pt]
    b_1 &= \kappa_F^{DI} \, F_0^{DI} + \kappa_F^{CP} \, F_0^{CP}
    \label{eq:b1} \\[4pt]
    \theta_1 &= \rho_\theta (1 - \phi_S \, S_0) \, \theta_0
    \label{eq:theta1}
\end{align}

\paragraph{Interpretation.} The persistence terms vanish because $y_0 = 0$: neither the hysteresis channel ($\psi S_0 y_0 = 0$) nor the capacity preservation channel ($\eta F_0^{CP} y_0 = 0$) is operative when the economy starts at potential. The initial output gap is determined entirely by the direct contractionary shock ($-\alpha_S S_0$) and the demand injection ($\alpha_F^{DI} F_0^{DI}$). Capacity preservation enters only through the debt equation ($\kappa_F^{CP} F_0^{CP}$)---in the first period, CP has \textit{no output effect}, only a fiscal cost. Its value materializes in subsequent periods through the persistence channel.

Mortality in period 1 is predetermined: $d_1 = \delta_\theta \theta_0$ is independent of all controls. The planner's first-period actions cannot prevent deaths that are already determined by the initial infection level---the irreducible policy lag.

\subsubsection*{Period 1 $\to$ 2}

Substituting (\ref{eq:y1})--(\ref{eq:theta1}):
%
\begin{align}
    y_2 &= \left[\rho_y + \psi S_1 - \eta F_1^{CP}\right] y_1
      - \alpha_S S_1 + \alpha_F^{DI} F_1^{DI}
    \notag \\[4pt]
    &= \left[\rho_y + \psi S_1 - \eta F_1^{CP}\right]
      \left(-\alpha_S S_0 + \alpha_F^{DI} F_0^{DI}\right)
      - \alpha_S S_1 + \alpha_F^{DI} F_1^{DI}
    \label{eq:y2}
\end{align}
%
\begin{align}
    d_2 &= \delta_\theta \, \theta_1
      = \delta_\theta \rho_\theta (1 - \phi_S S_0) \, \theta_0
    \label{eq:d2}
\end{align}
%
\begin{align}
    b_2 &= (1+r) b_1 - \gamma_y y_1
      + \kappa_F^{DI} F_1^{DI} + \kappa_F^{CP} F_1^{CP}
    \notag \\[4pt]
    &= (1+r)\left[\kappa_F^{DI} F_0^{DI} + \kappa_F^{CP} F_0^{CP}\right]
      - \gamma_y \left[-\alpha_S S_0 + \alpha_F^{DI} F_0^{DI}\right]
      + \kappa_F^{DI} F_1^{DI} + \kappa_F^{CP} F_1^{CP}
    \notag \\[4pt]
    &= (1+r)\kappa_F^{DI} F_0^{DI} + (1+r)\kappa_F^{CP} F_0^{CP}
      + \gamma_y \alpha_S S_0 - \gamma_y \alpha_F^{DI} F_0^{DI}
      + \kappa_F^{DI} F_1^{DI} + \kappa_F^{CP} F_1^{CP}
    \label{eq:b2}
\end{align}
%
\begin{align}
    \theta_2 &= \rho_\theta(1 - \phi_S S_1) \, \theta_1
      = \rho_\theta^2 (1 - \phi_S S_0)(1 - \phi_S S_1) \, \theta_0
    \label{eq:theta2}
\end{align}

\paragraph{Interpretation.}

\textit{Output gap.} The persistence channel is now active: the term $[\rho_y + \psi S_1 - \eta F_1^{CP}] \cdot y_1$ propagates the period-1 output gap forward, modulated by the current lockdown (hysteresis amplification) and current CP (recovery acceleration). Note that the period-0 CP ($F_0^{CP}$) still has no output effect---it is the \textit{current} CP ($F_1^{CP}$) that modulates the persistence of the \textit{inherited} gap. This confirms the one-period activation lag of the capacity preservation channel.

\textit{Mortality.} Period-2 mortality reflects the period-0 lockdown decision: $d_2 = \delta_\theta \rho_\theta (1 - \phi_S S_0) \theta_0$. This is the first period in which the planner's containment choice has a visible effect on mortality. The two-period lag $S_0 \to \theta_1 \to d_2$ is explicitly visible.

\textit{Debt.} Rearranging (\ref{eq:b2}) by period-0 controls:
%
\begin{align}
    b_2 &= \underbrace{\gamma_y \alpha_S \, S_0}_{\substack{\text{Indirect lockdown} \\ \text{cost via } \gamma_y}}
    + \underbrace{\left[(1+r)\kappa_F^{DI} - \gamma_y \alpha_F^{DI}\right] F_0^{DI}}_{\text{Net DI cost (constant)}}
    + \underbrace{(1+r)\kappa_F^{CP} \, F_0^{CP}}_{\substack{\text{Gross CP cost} \\ \text{(no offset in period 2)}}}
    + \underbrace{\kappa_F^{DI} F_1^{DI} + \kappa_F^{CP} F_1^{CP}}_{\text{Period-1 interventions}}
    \label{eq:b2_grouped}
\end{align}
%
The lockdown has no direct debt cost but generates an indirect cost $\gamma_y \alpha_S S_0$ through the automatic stabilizer channel. DI exhibits the fiscal dividend: its net cost is $(1+r)\kappa_F^{DI} - \gamma_y \alpha_F^{DI}$, which is below the gross cost whenever $\gamma_y \alpha_F^{DI} > 0$. Period-0 CP has no fiscal dividend visible yet in $b_2$, because its output effect (via $\rho_y^{\effd}$) only materializes in $y_2$, which feeds into $b_3$.

\subsubsection*{Period 2 $\to$ 3}

The full expansion becomes algebraically extensive. We present the key components grouped by economic mechanism.

\paragraph{Output gap.}
%
\begin{align}
    y_3 &= \left[\rho_y + \psi S_2 - \eta F_2^{CP}\right] y_2
      - \alpha_S S_2 + \alpha_F^{DI} F_2^{DI}
    \label{eq:y3}
\end{align}
%
Substituting (\ref{eq:y2}):
%
\begin{align}
    y_3 &= \left[\rho_y + \psi S_2 - \eta F_2^{CP}\right]
    \left\{
      \left[\rho_y + \psi S_1 - \eta F_1^{CP}\right]
      \left(-\alpha_S S_0 + \alpha_F^{DI} F_0^{DI}\right)
      - \alpha_S S_1 + \alpha_F^{DI} F_1^{DI}
    \right\}
    \notag \\
    &\quad - \alpha_S S_2 + \alpha_F^{DI} F_2^{DI}
    \label{eq:y3_full}
\end{align}
%
Defining $\rho_k^{\effd} \equiv \rho_y + \psi S_k - \eta F_k^{CP}$ for notational compactness:
%
\begin{equation}
    y_3 = \rho_2^{\effd} \rho_1^{\effd} (-\alpha_S S_0 + \alpha_F^{DI} F_0^{DI})
      + \rho_2^{\effd} (-\alpha_S S_1 + \alpha_F^{DI} F_1^{DI})
      + (-\alpha_S S_2 + \alpha_F^{DI} F_2^{DI})
    \label{eq:y3_compact}
\end{equation}
%
This reveals the recursive structure: the output gap in period 3 is the sum of all past net shocks $(-\alpha_S S_j + \alpha_F^{DI} F_j^{DI})$, each discounted by the product of effective persistence parameters from $j+1$ to 2. The effect of period-0 controls on $y_3$ is attenuated by $\rho_2^{\effd} \cdot \rho_1^{\effd}$---which is where CP exerts its leverage. Higher $F_1^{CP}$ and $F_2^{CP}$ reduce $\rho_1^{\effd}$ and $\rho_2^{\effd}$, causing past output gaps to dissipate faster.

\paragraph{Mortality.}
%
\begin{equation}
    d_3 = \delta_\theta \theta_2
      = \delta_\theta \rho_\theta^2 (1 - \phi_S S_0)(1 - \phi_S S_1) \, \theta_0
    \label{eq:d3}
\end{equation}
%
Mortality in period 3 depends on containment decisions in periods 0 and 1, each through the multiplicative suppression factor $(1 - \phi_S S_j)$. The period-2 lockdown $S_2$ does not yet affect $d_3$---it will affect $d_4$.

\paragraph{Debt.}
%
\begin{align}
    b_3 &= (1+r) b_2 - \gamma_y y_2
      + \kappa_F^{DI} F_2^{DI} + \kappa_F^{CP} F_2^{CP}
    \label{eq:b3_recursive}
\end{align}
%
Substituting (\ref{eq:b2}) and (\ref{eq:y2}) and collecting terms by period-0 controls:
%
\begin{align}
    \frac{\partial b_3}{\partial S_0}
      &= (1+r) \gamma_y \alpha_S
        + \gamma_y \rho_1^{\effd} \alpha_S
    \notag \\
      &= \gamma_y \alpha_S \left[(1+r) + \rho_1^{\effd}\right]
    \label{eq:db3_dS0}
\end{align}
%
\begin{align}
    \frac{\partial b_3}{\partial F_0^{DI}}
      &= (1+r)\left[(1+r)\kappa_F^{DI} - \gamma_y \alpha_F^{DI}\right]
        - \gamma_y \rho_1^{\effd} \alpha_F^{DI}
    \notag \\
      &= (1+r)^2 \kappa_F^{DI}
        - \gamma_y \alpha_F^{DI} \left[(1+r) + \rho_1^{\effd}\right]
    \label{eq:db3_dDI0}
\end{align}
%
\begin{align}
    \frac{\partial b_3}{\partial F_0^{CP}}
      &= (1+r)^2 \kappa_F^{CP}
        + \gamma_y \eta y_1 \rho_1^{\effd}
    \notag \\[2pt]
      &\quad\text{(Note: this term is nonzero only via the period-1 persistence effect)}
    \notag \\
      &= (1+r)^2 \kappa_F^{CP}
        - \gamma_y \eta |y_1| \rho_1^{\effd}
    \label{eq:db3_dCP0}
\end{align}
%
\textit{Interpretation.} The fiscal dividend of period-0 CP finally materializes in $b_3$: the term $-\gamma_y \eta |y_1| \rho_1^{\effd}$ represents the improved tax revenues and reduced mandatory expenditures that result from the faster recovery enabled by the earlier capacity preservation. This confirms the two-period activation lag of CP on the debt dimension: $F_0^{CP} \to \rho_1^{\effd} \downarrow \to y_2 \uparrow \to (-\gamma_y y_2) \downarrow \to b_3 \downarrow$.

\paragraph{Epidemiological dynamics.}
%
\begin{equation}
    \theta_3 = \rho_\theta^3 \prod_{j=0}^{2} (1 - \phi_S S_j) \cdot \theta_0
    \label{eq:theta3}
\end{equation}
%
The viral prevalence at period 3 is determined by the cumulative suppression effort across all three periods, each entering multiplicatively. This product structure implies that the marginal effectiveness of suppression in any period depends on the suppression in all other periods---lockdowns are complementary over time.

% ============================================================
\subsection*{A.4 Cumulative Effects of Period-0 Controls}
% ============================================================

The following table summarizes the effect of each period-0 control on all state variables across three periods. We assume $y_0 = 0$, $b_0 = 0$, and $y_k < 0$ for $k \geq 1$.

\begin{table}[h]
\centering
\caption{Period-0 Control Effects on State Variables}
\label{tab:control_effects}
\small
\begin{tabular}{@{}llccc@{}}
\toprule
State & Control & Period 1 & Period 2 & Period 3 \\
\midrule
\multirow{3}{*}{$y$}
  & $S_0$ & $-\alpha_S$ & $-\rho_1^{\effd} \alpha_S$ & $-\rho_2^{\effd}\rho_1^{\effd} \alpha_S$ \\
  & $F_0^{DI}$ & $+\alpha_F^{DI}$ & $+\rho_1^{\effd} \alpha_F^{DI}$ & $+\rho_2^{\effd}\rho_1^{\effd} \alpha_F^{DI}$ \\
  & $F_0^{CP}$ & $0$ & $0$\textsuperscript{a} & $0$\textsuperscript{a} \\
\midrule
\multirow{3}{*}{$d$}
  & $S_0$ & $0$ & $-\delta_\theta \rho_\theta \phi_S \theta_0$ & via $\theta_2$ \\
  & $F_0^{DI}$ & $0$ & $0$ & $0$ \\
  & $F_0^{CP}$ & $0$ & $0$ & $0$ \\
\midrule
\multirow{3}{*}{$b$}
  & $S_0$ & $0$ & $+\gamma_y \alpha_S$ & $+\gamma_y \alpha_S[(1{+}r) + \rho_1^{\effd}]$ \\
  & $F_0^{DI}$ & $+\kappa_F^{DI}$ & $(1{+}r)\kappa_F^{DI} {-} \gamma_y \alpha_F^{DI}$ & Eq.~(\ref{eq:db3_dDI0}) \\
  & $F_0^{CP}$ & $+\kappa_F^{CP}$ & $(1{+}r)\kappa_F^{CP}$ & $(1{+}r)^2\kappa_F^{CP} {-} \gamma_y \eta|y_1|\rho_1^{\effd}$ \\
\midrule
\multirow{3}{*}{$\theta$}
  & $S_0$ & $-\phi_S \rho_\theta \theta_0$ & $-\phi_S \rho_\theta^2(1{-}\phi_S S_1)\theta_0$ & cumulative \\
  & $F_0^{DI}$ & $0$ & $0$ & $0$ \\
  & $F_0^{CP}$ & $0$ & $0$ & $0$ \\
\bottomrule
\end{tabular}

\medskip
\footnotesize
\textsuperscript{a} $F_0^{CP}$ has no \textit{direct} effect on $y_2$ or $y_3$. Its effect operates indirectly through $\rho_1^{\effd}$ and $\rho_2^{\effd}$ when combined with subsequent CP, which modulates the persistence of the gap created by other controls. The partial derivative $\partial y_2 / \partial F_0^{CP} = 0$ because $y_0 = 0$ and CP enters multiplicatively on $y_k$.
\end{table}

% ============================================================
\subsection*{A.5 Optimality Conditions}
% ============================================================

The social planner minimizes the expected intertemporal cost:
%
\begin{equation}
    J = E\left\{x_N' Q_N x_N + \sum_{k=0}^{N-1} \left[x_k' Q_k x_k + u_k' R u_k\right]\right\}
    \label{eq:objective}
\end{equation}
%
with
%
\begin{equation}
    Q_k = \beta^k \begin{pmatrix} w_y & 0 & 0 & 0 \\ 0 & w_d & 0 & 0 \\ 0 & 0 & w_b & 0 \\ 0 & 0 & 0 & 0 \end{pmatrix}, \quad
    R = \begin{pmatrix} r_S & 0 & 0 \\ 0 & r_{DI} & 0 \\ 0 & 0 & r_{CP} \end{pmatrix}, \quad
    Q_N = \beta^N \begin{pmatrix} 0 & 0 & 0 & 0 \\ 0 & 0 & 0 & 0 \\ 0 & 0 & W_b & 0 \\ 0 & 0 & 0 & 0 \end{pmatrix}
    \label{eq:cost_matrices}
\end{equation}

\paragraph{Bellman equation.} The cost-to-go from period $k$ is:
%
\begin{equation}
    V_k(x_k) = \min_{u_k} \left\{x_k' Q_k x_k + u_k' R u_k + E\left[V_{k+1}(f(x_k, u_k) + G\varepsilon_k)\right]\right\}
    \label{eq:bellman}
\end{equation}
%
with terminal condition $V_N(x_N) = x_N' Q_N x_N$.

\paragraph{First-order conditions.} The necessary condition for optimality at each period $k$ is:
%
\begin{equation}
    \frac{\partial}{\partial u_k}\left[u_k' R u_k + E\left[V_{k+1}(f(x_k, u_k, \varepsilon_k))\right]\right] = 0
    \label{eq:foc_general}
\end{equation}
%
which yields:
%
\begin{equation}
    2 R u_k + B_k' \nabla_x V_{k+1}\big|_{x_{k+1} = f(x_k, u_k)} = 0
    \label{eq:foc}
\end{equation}
%
where $B_k = \partial f / \partial u_k$ is the Jacobian (\ref{eq:B_k}). Defining the costate (shadow price) vector $p_{k+1} = \nabla_x V_{k+1}(x_{k+1})$, the optimality condition becomes:
%
\begin{equation}
    u_k^* = -\frac{1}{2} R^{-1} B_k' \, p_{k+1}
    \label{eq:optimal_control}
\end{equation}

\paragraph{Control-specific optimality conditions.} Expanding (\ref{eq:optimal_control}) using the structure of $B_k$ and denoting the components of the costate vector as $p_{k+1} = (p_{k+1}^y, p_{k+1}^d, p_{k+1}^b, p_{k+1}^\theta)'$:

\medskip
\noindent\textit{Optimal lockdown intensity:}
%
\begin{equation}
    S_k^* = -\frac{1}{2 r_S}\left[
      (\psi y_k - \alpha_S) \, p_{k+1}^y
      - \phi_S \rho_\theta \theta_k \, p_{k+1}^\theta
    \right]
    \label{eq:opt_S}
\end{equation}
%
The lockdown balances two shadow prices: the output cost ($p_{k+1}^y$, weighted by the state-dependent marginal effect $\psi y_k - \alpha_S$) against the epidemiological benefit ($p_{k+1}^\theta$, weighted by the suppression effectiveness $\phi_S \rho_\theta \theta_k$). At higher $|\theta_k|$ (severe pandemic), the second term dominates, favoring stricter containment. At deeper $|y_k|$ (severe recession), the first term increases in magnitude, favoring looser containment. The political cost $r_S$ dampens overall lockdown intensity.

\medskip
\noindent\textit{Optimal demand injection:}
%
\begin{equation}
    F_k^{DI*} = -\frac{1}{2 r_{DI}}\left[
      \alpha_F^{DI} \, p_{k+1}^y
      + \kappa_F^{DI} \, p_{k+1}^b
    \right]
    \label{eq:opt_DI}
\end{equation}
%
Demand injection balances the output benefit ($\alpha_F^{DI} p_{k+1}^y$) against the debt cost ($\kappa_F^{DI} p_{k+1}^b$). Since $p_{k+1}^y < 0$ (negative output gap is costly) and $p_{k+1}^b > 0$ (debt is costly), the two terms have opposite signs: the output benefit favors more DI, the debt cost restrains it. The ratio $\alpha_F^{DI} / \kappa_F^{DI}$ governs the relative strength---confirming the ``price of stabilization in debt units'' identified in Property~3.

\medskip
\noindent\textit{Optimal capacity preservation:}
%
\begin{equation}
    F_k^{CP*} = -\frac{1}{2 r_{CP}}\left[
      -\eta y_k \, p_{k+1}^y
      + \kappa_F^{CP} \, p_{k+1}^b
    \right]
    \label{eq:opt_CP}
\end{equation}
%
Capacity preservation balances the persistence reduction benefit ($-\eta y_k \, p_{k+1}^y$) against the debt cost ($\kappa_F^{CP} p_{k+1}^b$). The output benefit term is state-dependent: $-\eta y_k > 0$ when $y_k < 0$, and its magnitude scales with $|y_k|$. In a deeper recession, the shadow value of CP increases proportionally---confirming the analytical result that CP becomes relatively more valuable as the crisis deepens.

\paragraph{Comparison of fiscal instruments.} From (\ref{eq:opt_DI}) and (\ref{eq:opt_CP}), the relative deployment of DI and CP at the optimum satisfies:
%
\begin{equation}
    \frac{F_k^{DI*}}{F_k^{CP*}} =
    \frac{r_{CP}}{r_{DI}} \cdot
    \frac{\alpha_F^{DI} \, p_{k+1}^y + \kappa_F^{DI} \, p_{k+1}^b}
         {-\eta y_k \, p_{k+1}^y + \kappa_F^{CP} \, p_{k+1}^b}
    \label{eq:ratio_DI_CP}
\end{equation}
%
Since $|{-\eta y_k}|$ is increasing in the depth of the recession while $\alpha_F^{DI}$ is constant, this ratio is \textit{decreasing} in $|y_k|$: the optimal fiscal mix shifts toward capacity preservation as the crisis deepens. In the limit of a sufficiently severe recession ($|y_k| \to \infty$), the ratio converges to $\frac{r_{CP}}{r_{DI}} \cdot \frac{\kappa_F^{DI}}{\kappa_F^{CP}}$---the optimal mix is determined entirely by the relative political costs and fiscal cost parameters, as the output benefits of CP dominate those of DI.

\paragraph{Interaction between lockdown and fiscal composition.} The optimal lockdown (\ref{eq:opt_S}) depends on $p_{k+1}^y$ and $p_{k+1}^\theta$, which are themselves functions of the entire future trajectory---including future fiscal policy. Through the costate recursion, a higher expected future CP (reducing $\rho_k^{\effd}$ in future periods) lowers the cost-to-go associated with a given output gap today, which reduces $|p_{k+1}^y|$ and thereby \textit{permits a stricter lockdown} at lower welfare cost. This formalizes the enabling role of CP: by ensuring that the economy can recover from the lockdown, CP expands the planner's feasible set on the containment dimension.

\subsection*{A.6 Summary of Structural Properties}

\begin{enumerate}
    \item \textbf{Temporal asymmetry.} Fiscal stimulus (DI) affects output within one period; containment affects mortality with a two-period lag ($S_k \to \theta_{k+1} \to d_{k+2}$); capacity preservation affects debt with a two-period lag ($F_k^{CP} \to \rho_{k+1}^{\effd} \to y_{k+2} \to b_{k+3}$).

    \item \textbf{State-dependent marginal effects.} The output cost of lockdown ($\alpha_S + \psi|y_k|$) and the output benefit of CP ($\eta|y_k|$) both scale with crisis severity. DI has constant marginal effects.

    \item \textbf{Optimal composition shift.} The ratio $F_k^{DI*} / F_k^{CP*}$ is decreasing in $|y_k|$: the optimal fiscal mix shifts from DI-heavy (mild recession) to CP-heavy (deep recession).

    \item \textbf{Fiscal dividend asymmetry.} DI generates fiscal dividends within one period (via $-\gamma_y \alpha_F^{DI}$). CP generates fiscal dividends with a two-period lag (via $\rho_k^{\effd} \to y \to \gamma_y$). Both partially offset their gross fiscal costs.

    \item \textbf{Complementarity.} CP enables stricter containment by reducing its welfare cost (through lower $|p_{k+1}^y|$). DI addresses the output gap that CP alone cannot close. The optimal policy uses both instruments simultaneously.

    \item \textbf{Rank deficiency.} $\text{rank}(B_k) = 3 < 4$: mortality cannot be controlled directly. The trilemma persists despite three control instruments, because no control independently targets a single state dimension.
\end{enumerate}

\subsection{Stability Analysis}
\label{sec:stability}

The constitutive properties established in the preceding section characterize the
\textit{mechanisms} of the pandemic trilemma. This section examines the
\textit{dynamical structure} of the system by analyzing the stability properties of
the linearized transition matrix. The analysis reveals that the planner confronts a
system with multiple simultaneously unstable modes---a configuration that transforms
the problem from multi-objective optimization into a genuine stabilization challenge.

\subsubsection{Eigenvalue Structure}

The linearized system matrix $A_k$ (equation~\ref{eq:A_k}) is block-triangular, so its
eigenvalues coincide with the diagonal entries:
%
\begin{align}
    \lambda_1^k &= \rho_y + \psi \, S_k - \eta \, F_k^{CP}
      & &\text{(output persistence)}
    \label{eq:lambda1} \\[4pt]
    \lambda_2^k &= 0
      & &\text{(mortality)}
    \label{eq:lambda2} \\[4pt]
    \lambda_3   &= 1 + r
      & &\text{(debt dynamics)}
    \label{eq:lambda3} \\[4pt]
    \lambda_4^k &= \rho_\theta (1 - \phi_S \, S_k)
      & &\text{(epidemiological dynamics)}
    \label{eq:lambda4}
\end{align}
%
A state dimension is stable if and only if its associated eigenvalue lies strictly
inside the unit circle ($|\lambda| < 1$). The mortality eigenvalue $\lambda_2 = 0$
implies that excess deaths carry no autonomous persistence: each period's mortality is
entirely determined by the preceding period's viral prevalence and does not propagate
forward independently. The remaining three eigenvalues govern the stability of the
planner's active state dimensions and merit individual analysis.

\subsubsection{Unconditionally Unstable: Public Debt}

The debt eigenvalue $\lambda_3 = 1 + r > 1$ lies outside the unit circle for any
positive real interest rate. This instability is \textit{unconditional}: it holds
regardless of the planner's policy choices, as no element of the control vector $u_k$
enters the $(3,3)$-entry of $A_k$. Existing debt compounds mechanically, and the
planner cannot alter this dynamic through any feasible intervention.

The only forces counteracting this explosive tendency operate through the off-diagonal
terms of the debt transition: the automatic stabilizer channel $-\gamma_y y_k$ and the
primary balance effect of fiscal consolidation. During the pandemic, both channels work
in the destabilizing direction---$y_k < 0$ generates additional debt through
$-\gamma_y y_k > 0$, and active fiscal support adds $\kappa_F^{DI} F_k^{DI} +
\kappa_F^{CP} F_k^{CP} > 0$. The debt state therefore accelerates away from the origin
throughout the crisis period.

The post-pandemic environment demonstrated the practical relevance of this instability.
The inflation surge of 2022--2023 triggered substantial increases in policy rates across
advanced economies \citep{jorda2023}, raising $r$ and thereby amplifying the explosive
eigenvalue. Countries that exited the pandemic with higher $b_N$ were
disproportionately exposed to this tightening---a risk that was uncertain but
foreseeable at the time of the fiscal deployment decisions. The terminal cost $W_b$ in
the objective function formalizes this concern: it disciplines the planner's
intertemporal trade-off by imposing a cost on the debt state at the planning horizon,
reflecting the present value of the post-pandemic fiscal burden under uncertain future
interest rates.

\subsubsection{Conditionally Unstable: Epidemiological Dynamics}

Without intervention ($S_k = 0$), the epidemiological eigenvalue equals $\lambda_4 =
\rho_\theta > 1$: viral prevalence grows exponentially, as each generation of
infections produces more than one subsequent generation. The planner can stabilize
this dimension by deploying containment measures of sufficient intensity. The
stabilization condition requires:
%
\begin{equation}
    \lambda_4^k < 1 \quad \Longleftrightarrow \quad
    S_k > \frac{1 - 1/\rho_\theta}{\phi_S}
    \equiv \underline{S}
    \label{eq:S_threshold}
\end{equation}
%
The threshold $\underline{S}$ represents the minimum lockdown intensity required to
reverse the epidemic trajectory---the formal counterpart of the widely cited
$R_{\text{eff}} < 1$ condition. Below this threshold, the epidemic grows despite
intervention; at or above it, infections decline. Three features of this condition are
noteworthy.

First, the threshold depends on the ratio of the virus's intrinsic transmissibility
($\rho_\theta$) to the effectiveness of containment ($\phi_S$). A more transmissible
variant (higher $\rho_\theta$) requires a stricter lockdown; more effective NPIs (higher
$\phi_S$) lower the required intensity. Second, the condition is binary: there exists a
critical level below which suppression is ineffective at reversing the epidemic's
direction, regardless of the economic and fiscal costs incurred. Third, the threshold is
time-invariant in the model's parameterization---it does not depend on the current level
of infections $\theta_k$. The \textit{level} of prevalence determines the number of
deaths (through $d_{k+2} = \delta_\theta \theta_{k+1}$), but the \textit{stability} of
the epidemiological state depends only on the structural parameters $\rho_\theta$ and
$\phi_S$.\footnote{This time-invariance is a consequence of the linearized
specification. In a nonlinear SIR framework, the effective reproduction rate depends on
the remaining susceptible population, introducing implicit time-variation. The linear
approximation is appropriate for the early phase of a pandemic where the susceptible
fraction remains close to unity.}

\subsubsection{Conditionally Unstable: Output Dynamics}

In the absence of a pandemic ($S_k = 0$, $F_k^{CP} = 0$), the output eigenvalue
equals $\lambda_1 = \rho_y < 1$: the economy is inherently stable, with output gaps
closing through natural mean reversion. This reflects the standard macroeconomic
assumption that market economies possess self-correcting mechanisms---price adjustments,
inventory cycles, and reallocation of factors of production---that gradually restore
output to its potential.

The pandemic disrupts this stability through the hysteresis channel formalized in
Property~2. Containment measures increase the output eigenvalue by $\psi S_k$:
%
\begin{equation}
    \lambda_1^k = \rho_y + \psi \, S_k - \eta \, F_k^{CP}
    \label{eq:lambda1_full}
\end{equation}
%
The critical threshold at which the output state becomes unstable is:
%
\begin{equation}
    \lambda_1^k \geq 1 \quad \Longleftrightarrow \quad
    \psi \, S_k - \eta \, F_k^{CP} \geq 1 - \rho_y
    \label{eq:hysteresis_threshold}
\end{equation}
%
When this condition is met, the output gap becomes self-reinforcing: firms exit the
market, employment relationships dissolve, human capital depreciates, and supply chains
fragment---each of these effects reducing the economy's capacity to recover, which in
turn deepens the contraction. The output state transitions from a stable, mean-reverting
process to an explosive one. I refer to this regime as the \textit{hysteresis trap}: a
state in which the structural damage inflicted by prolonged containment has eroded the
economy's self-correcting capacity to the point where recovery requires active
intervention merely to prevent further deterioration.

Capacity preservation operates as the direct counterforce to this instability. Each unit
of $F_k^{CP}$ reduces $\lambda_1^k$ by $\eta$, pulling the eigenvalue back toward the
stable region. The stabilization condition for the output state is:
%
\begin{equation}
    \lambda_1^k < 1 \quad \Longleftrightarrow \quad
    F_k^{CP} > \frac{\psi \, S_k - (1 - \rho_y)}{\eta}
    \equiv \underline{F}^{CP}(S_k)
    \label{eq:CP_threshold}
\end{equation}
%
This threshold is increasing in $S_k$: a stricter lockdown requires more capacity
preservation to prevent the output state from becoming unstable. The threshold is
positive only when $\psi S_k > 1 - \rho_y$, i.e., when the lockdown is sufficiently
stringent to overcome the economy's natural stability margin. For mild containment
measures, the economy remains self-stabilizing without active capacity support.

\subsubsection{The Competing Instabilities}

The central insight of the stability analysis is that the planner faces not one but
potentially three simultaneously unstable eigenvalues, and that the policy actions
required to stabilize one dimension systematically destabilize another. Table~\ref{tab:stability}
summarizes this structure.

\begin{table}[h]
\centering
\caption{Eigenvalue Structure and Stability Conditions}
\label{tab:stability}
\small
\begin{tabular}{@{}lcccc@{}}
\toprule
Eigenvalue & Without policy & Stability condition & Stabilizing control & Side effect \\
\midrule
$\lambda_1^k$ (Output)
  & $\rho_y < 1$ (stable)
  & $\psi S_k - \eta F_k^{CP} < 1 - \rho_y$
  & $F_k^{CP} \uparrow$
  & $b_{k+1} \uparrow$ \\[6pt]
$\lambda_3$ (Debt)
  & $(1+r) > 1$ (unstable)
  & Cannot be stabilized
  & ---
  & --- \\[6pt]
$\lambda_4^k$ (Epidemic)
  & $\rho_\theta > 1$ (unstable)
  & $S_k > \underline{S}$
  & $S_k \uparrow$
  & $\lambda_1^k \uparrow$, $y_{k+1} \downarrow$ \\
\bottomrule
\end{tabular}
\end{table}

The competing instabilities create a cascading stabilization problem:
%
\begin{enumerate}
    \item The epidemiological instability ($\lambda_4 > 1$) necessitates containment
    ($S_k > \underline{S}$).

    \item Containment destabilizes the output dimension ($\lambda_1$ increases toward
    and potentially beyond unity), necessitating capacity preservation ($F_k^{CP} >
    \underline{F}^{CP}(S_k)$).

    \item The resulting output gap ($y_k < 0$) widens the deficit through automatic
    stabilizers ($-\gamma_y y_k > 0$), necessitating demand injection ($F_k^{DI} > 0$)
    to close the gap.

    \item Both fiscal instruments augment the unconditionally unstable debt state
    ($\kappa_F^{DI} F_k^{DI} + \kappa_F^{CP} F_k^{CP} > 0$), which cannot be
    stabilized by any available control.
\end{enumerate}
%
This cascade formalizes the trilemma as a dynamical systems problem: the planner must
simultaneously manage two conditionally unstable modes and one unconditionally unstable
mode with three control instruments, where each stabilizing action on one mode
contributes to the instability of at least one other.

\subsubsection{Implications for Optimal Policy Timing}

The eigenvalue structure provides a formal rationale for the empirically documented
advantage of early intervention
\citep{balmford2020,acemoglu2021}.

\paragraph{Early lockdown.} When $\theta_k$ is small (early in the epidemic), the
absolute number of infections prevented per unit of containment is lower, but the
\textit{growth rate} that must be suppressed is the same ($\rho_\theta$ is constant).
The cost of containment, however, is state-dependent through the output channel: the
marginal output cost is $\alpha_S + \psi |y_k|$, which is smaller when $|y_k|$ is small.
Early lockdown therefore achieves the same epidemiological stabilization ($\lambda_4 < 1$)
at lower economic cost. Delay allows $\theta_k$ to grow, which---through the mortality
channel $d_{k+2} = \delta_\theta \theta_{k+1}$---increases the cumulative death toll and
simultaneously deepens the output gap (as behavioral responses and eventually stricter
containment become necessary), raising the cost of all subsequent interventions.

\paragraph{Early capacity preservation.} The hysteresis parameter $\psi$ implies that
the structural damage to the economy's recovery capacity accumulates \textit{during} the
lockdown. Each period of containment without adequate CP pushes $\lambda_1$ closer to
the instability threshold. Once $\lambda_1 \geq 1$, the output gap becomes
self-reinforcing and requires progressively larger interventions merely to prevent
further deterioration---let alone to close the gap. Early CP deployment maintains
$\lambda_1$ in the stable region, preserving the economy's capacity for autonomous
recovery once containment measures are relaxed.

\paragraph{Convexity of delay costs.} The interaction between the competing
instabilities implies that the cost of delayed intervention is \textit{convex} rather
than linear. A one-period delay in containment allows $\theta_k$ to grow by a factor
$\rho_\theta$, requiring a commensurately larger output sacrifice when containment is
eventually imposed. The deeper output gap in turn raises the cost of containment (through
$\psi |y_k|$), increases the required CP (through the threshold $\underline{F}^{CP}(S_k)$),
and accelerates debt accumulation (through $-\gamma_y y_k$). Each channel amplifies the
others, generating a superlinear relationship between delay duration and total welfare
cost. This convexity is a direct consequence of the system's multi-dimensional
instability structure and cannot be captured by models that treat health and economic
outcomes as independent trade-offs.

% ============================================================
% Two-Period Debt Effects: ∂b_{k+2}/∂F_k^{DI} and ∂b_{k+2}/∂F_k^{CP}
% ============================================================

The debt transition one period forward is:
%
\begin{equation}
    b_{k+2} = (1+r) \, b_{k+1} - \gamma_y \, y_{k+1}
      + \kappa_F^{DI} \, F_{k+1}^{DI}
      + \kappa_F^{CP} \, F_{k+1}^{CP}
      + c_{k+1}^{H}
\end{equation}
%
Period-$k$ fiscal decisions affect $b_{k+2}$ through two channels: directly via
their contribution to $b_{k+1}$, which compounds at $(1+r)$, and indirectly via
their effect on $y_{k+1}$, which feeds through the automatic stabilizer term
$-\gamma_y y_{k+1}$. Differentiating with respect to each instrument while
holding period-$(k+1)$ controls constant yields:

\paragraph{Demand Injection.}
%
\begin{equation}
    \frac{\partial b_{k+2}}{\partial F_k^{DI}}
      = \underbrace{(1+r) \, \kappa_F^{DI}}_{\substack{\text{direct fiscal cost,} \\ \text{compounded}}}
      \;-\; \underbrace{\gamma_y \, \alpha_F^{DI}}_{\substack{\text{automatic stabilizer} \\ \text{offset}}}
    \label{eq:db2_dDI}
\end{equation}
%
The two-period debt effect of DI is \textit{state-independent}: it depends only
on the structural parameters $\kappa_F^{DI}$, $\alpha_F^{DI}$, $\gamma_y$, and
$r$, but not on the current output gap or containment intensity. The first term
captures the gross fiscal cost of DI, compounded by one period of interest
accrual. The second term captures the fiscal dividend: by improving $y_{k+1}$ by
$\alpha_F^{DI}$, demand injection reduces automatic debt accumulation in period
$k+2$ by $\gamma_y \alpha_F^{DI}$. The net effect is positive under pandemic
conditions, as established in (\ref{eq:net_cost_DI}).

\paragraph{Capacity Preservation.}

Recall that $\partial y_{k+1} / \partial F_k^{CP} = -\eta \, y_k \geq 0$
(positive because $y_k \leq 0$). Substituting:
%
\begin{equation}
    \frac{\partial b_{k+2}}{\partial F_k^{CP}}
      = \underbrace{(1+r) \, \kappa_F^{CP}}_{\substack{\text{direct fiscal cost,} \\ \text{compounded}}}
      \;-\; \underbrace{\gamma_y \, \eta \, |y_k|}_{\substack{\text{automatic stabilizer} \\ \text{offset (state-dependent)}}}
    \label{eq:db2_dCP}
\end{equation}
%
The two-period debt effect of CP is \textit{state-dependent}: the automatic
stabilizer offset $\gamma_y \eta |y_k|$ scales with the depth of the recession.
In a mild downturn ($|y_k|$ small), the offset is negligible and the net fiscal
cost is approximately $(1+r) \kappa_F^{CP}$---the gross cost, compounded. As the
recession deepens, the offset grows, and the net debt effect falls. The
critical threshold at which CP becomes self-financing over two periods is:
%
\begin{equation}
    |y_k^{**}| \geq \frac{(1+r) \, \kappa_F^{CP}}{\gamma_y \, \eta}
    \label{eq:CP_self_financing}
\end{equation}
%
Beyond this threshold, the automatic stabilizer savings from capacity
preservation exceed its direct fiscal cost: each additional unit of CP
\textit{reduces} cumulative debt over two periods. This is possible because CP
does not generate output directly---it accelerates recovery from an existing
gap. When that gap is sufficiently deep, the budgetary improvement from faster
recovery exceeds the cost of the instrument.

\paragraph{Comparison and Implications.}

Juxtaposing (\ref{eq:db2_dDI}) and (\ref{eq:db2_dCP}) reveals three properties
of the planner's fiscal composition problem.

First, \textit{DI has constant net fiscal cost; CP has declining net fiscal
cost.} The two-period debt effect of DI is a fixed number determined by
structural parameters. The two-period debt effect of CP falls with the severity
of the recession and can become negative. This implies that the relative fiscal
advantage of CP over DI increases as the crisis deepens---precisely the
conditions under which strict containment is most necessary.

Second, \textit{the self-financing thresholds differ in kind.} DI is
self-financing when $\alpha_F^{DI} \geq (1+r) \kappa_F^{DI} / \gamma_y$,
which requires a multiplier above approximately 2 under standard
parameterizations---above pandemic-era estimates. CP is self-financing when
$|y_k| \geq (1+r) \kappa_F^{CP} / (\gamma_y \eta)$, a condition on the
\textit{state of the economy}, not on a structural parameter. A sufficiently
deep recession can render CP self-financing even when its effectiveness $\eta$
is modest, because the automatic stabilizer base ($|y_k|$) is large.

Third, \textit{the two instruments interact through their effect on the
self-financing condition.} CP that prevents hysteresis keeps $|y_k|$ from
growing, which---paradoxically---reduces the future automatic stabilizer offset
of CP itself. But the same prevention of hysteresis also reduces the
\textit{volume} of DI required in subsequent periods. The net effect on total
debt depends on which margin dominates: the reduced automatic stabilizer offset
per unit of CP, or the reduced volume of DI (at its fixed, higher per-unit
cost). The optimal composition trades off these two margins dynamically across
the planning horizon.


\section{Derivation of the Optimal Control Law via Dynamic Programming}
\label{app:LQR_derivation}

This appendix provides a self-contained derivation of the optimal 
linear-quadratic regulator (LQR) solution following 
\citet[Chapter~3.1]{bertsekas2017dynamic}. The derivation proceeds 
via backward induction using the principle of dynamic programming.

\subsection{Problem Statement}

Consider a discrete-time linear stochastic system evolving according to:
\begin{equation}
    \bm{x}_{k+1} = \bm{A}\bm{x}_k + \bm{B}\bm{u}_k + \bm{w}_k, 
    \quad k = 0, 1, \ldots, N-1
    \label{eq:app_system}
\end{equation}
where $\bm{x}_k \in \mathbb{R}^n$ is the state vector, 
$\bm{u}_k \in \mathbb{R}^m$ is the control vector, and 
$\bm{w}_k \in \mathbb{R}^n$ is a stochastic disturbance with 
$\mathbb{E}[\bm{w}_k] = \bm{0}$ and 
$\mathbb{E}[\bm{w}_k \bm{w}_k'] = \bm{\Sigma}_w$. The disturbances 
are assumed to be independent across time and independent of the 
initial state $\bm{x}_0$.

The social planner seeks to minimize the expected intertemporal 
quadratic cost:
\begin{equation}
    J = \mathbb{E} \left\{ \bm{x}_N' \bm{Q}_N \bm{x}_N + 
    \sum_{k=0}^{N-1} \left( \bm{x}_k' \bm{Q}_k \bm{x}_k + 
    \bm{u}_k' \bm{R} \bm{u}_k \right) \right\}
    \label{eq:app_cost}
\end{equation}
where the weighting matrices satisfy:
\begin{itemize}
    \item $\bm{Q}_k \succeq 0$ (positive semidefinite) for all 
    $k = 0, \ldots, N-1$
    \item $\bm{Q}_N \succeq 0$ (positive semidefinite)
    \item $\bm{R} \succ 0$ (positive definite)
\end{itemize}

\subsection{The Principle of Optimality}

The solution proceeds via \textit{dynamic programming}, which 
exploits Bellman's principle of optimality: an optimal policy has 
the property that, regardless of the initial state and decision, 
the remaining decisions must constitute an optimal policy with 
regard to the state resulting from the first decision.

Define the \textit{cost-to-go function} $J_k(\bm{x}_k)$ as the 
minimum expected cost from period $k$ to the terminal period $N$, 
conditional on the current state $\bm{x}_k$:
\begin{equation}
    J_k(\bm{x}_k) = \min_{\bm{u}_k, \ldots, \bm{u}_{N-1}} 
    \mathbb{E} \left\{ \bm{x}_N' \bm{Q}_N \bm{x}_N + 
    \sum_{t=k}^{N-1} \left( \bm{x}_t' \bm{Q}_t \bm{x}_t + 
    \bm{u}_t' \bm{R} \bm{u}_t \right) \,\Big|\, \bm{x}_k \right\}
    \label{eq:app_costtogo}
\end{equation}

The principle of optimality implies that the cost-to-go satisfies 
the \textit{Bellman equation}:
\begin{equation}
    J_k(\bm{x}_k) = \min_{\bm{u}_k} \mathbb{E}_{\bm{w}_k} 
    \left\{ \bm{x}_k' \bm{Q}_k \bm{x}_k + \bm{u}_k' \bm{R} \bm{u}_k 
    + J_{k+1}(\bm{A}\bm{x}_k + \bm{B}\bm{u}_k + \bm{w}_k) \right\}
    \label{eq:app_bellman}
\end{equation}

This recursive relationship allows us to solve the problem 
\textit{backward} from the terminal period $k = N$ to the initial 
period $k = 0$.

\subsection{Terminal Condition}

At the terminal period $k = N$, no further control is applied. The 
cost-to-go consists solely of the terminal penalty:
\begin{equation}
    J_N(\bm{x}_N) = \bm{x}_N' \bm{Q}_N \bm{x}_N
    \label{eq:app_terminal}
\end{equation}

This is a quadratic function of the state. We write:
\begin{equation}
    J_N(\bm{x}_N) = \bm{x}_N' \bm{P}_N \bm{x}_N 
    \quad \text{with} \quad \bm{P}_N = \bm{Q}_N
    \label{eq:app_P_terminal}
\end{equation}

\textbf{Induction Hypothesis:} We conjecture that the cost-to-go 
function is quadratic at every stage:
\begin{equation}
    J_{k+1}(\bm{x}_{k+1}) = \bm{x}_{k+1}' \bm{P}_{k+1} \bm{x}_{k+1} 
    + c_{k+1}
    \label{eq:app_hypothesis}
\end{equation}
where $\bm{P}_{k+1}$ is a symmetric positive semidefinite matrix 
and $c_{k+1}$ is a scalar constant (independent of the state). We 
will show that this structure is preserved under the backward 
recursion.

\subsection{Backward Induction Step}

Substituting the conjectured form \eqref{eq:app_hypothesis} into 
the Bellman equation \eqref{eq:app_bellman}:
\begin{equation}
    J_k(\bm{x}_k) = \min_{\bm{u}_k} \mathbb{E}_{\bm{w}_k} 
    \Big\{ \bm{x}_k' \bm{Q}_k \bm{x}_k + \bm{u}_k' \bm{R} \bm{u}_k 
    + (\bm{A}\bm{x}_k + \bm{B}\bm{u}_k + \bm{w}_k)' \bm{P}_{k+1} 
    (\bm{A}\bm{x}_k + \bm{B}\bm{u}_k + \bm{w}_k) + c_{k+1} \Big\}
    \label{eq:app_substitute}
\end{equation}

Expanding the quadratic term in the next-period state:
\begin{align}
    &(\bm{A}\bm{x}_k + \bm{B}\bm{u}_k + \bm{w}_k)' \bm{P}_{k+1} 
    (\bm{A}\bm{x}_k + \bm{B}\bm{u}_k + \bm{w}_k) \nonumber \\
    &= (\bm{A}\bm{x}_k + \bm{B}\bm{u}_k)' \bm{P}_{k+1} 
    (\bm{A}\bm{x}_k + \bm{B}\bm{u}_k) \label{eq:app_expand1} \\
    &\quad + 2(\bm{A}\bm{x}_k + \bm{B}\bm{u}_k)' \bm{P}_{k+1} 
    \bm{w}_k \label{eq:app_expand2} \\
    &\quad + \bm{w}_k' \bm{P}_{k+1} \bm{w}_k \label{eq:app_expand3}
\end{align}

Taking expectations and using $\mathbb{E}[\bm{w}_k] = \bm{0}$, the 
cross-term \eqref{eq:app_expand2} vanishes:
\begin{equation}
    \mathbb{E}\left[ 2(\bm{A}\bm{x}_k + \bm{B}\bm{u}_k)' 
    \bm{P}_{k+1} \bm{w}_k \right] = 2(\bm{A}\bm{x}_k + 
    \bm{B}\bm{u}_k)' \bm{P}_{k+1} \, \mathbb{E}[\bm{w}_k] = 0
    \label{eq:app_crossterm}
\end{equation}

The expectation of the noise term \eqref{eq:app_expand3} yields a 
constant:
\begin{equation}
    \mathbb{E}[\bm{w}_k' \bm{P}_{k+1} \bm{w}_k] = 
    \text{tr}(\bm{P}_{k+1} \bm{\Sigma}_w) =: \sigma_k^2
    \label{eq:app_noiseterm}
\end{equation}

Substituting back into \eqref{eq:app_substitute}:
\begin{align}
    J_k(\bm{x}_k) &= \bm{x}_k' \bm{Q}_k \bm{x}_k + c_{k+1} + 
    \sigma_k^2 \nonumber \\
    &\quad + \min_{\bm{u}_k} \Big\{ \bm{u}_k' \bm{R} \bm{u}_k + 
    (\bm{A}\bm{x}_k + \bm{B}\bm{u}_k)' \bm{P}_{k+1} 
    (\bm{A}\bm{x}_k + \bm{B}\bm{u}_k) \Big\}
    \label{eq:app_Jk_simplified}
\end{align}

\subsection{Minimization with Respect to the Control}

Define the function to be minimized over $\bm{u}_k$:
\begin{equation}
    f(\bm{u}_k) = \bm{u}_k' \bm{R} \bm{u}_k + 
    (\bm{A}\bm{x}_k + \bm{B}\bm{u}_k)' \bm{P}_{k+1} 
    (\bm{A}\bm{x}_k + \bm{B}\bm{u}_k)
    \label{eq:app_f_uk}
\end{equation}

Expanding the second term:
\begin{align}
    (\bm{A}\bm{x}_k + \bm{B}\bm{u}_k)' \bm{P}_{k+1} 
    (\bm{A}\bm{x}_k + \bm{B}\bm{u}_k) 
    &= \bm{x}_k' \bm{A}' \bm{P}_{k+1} \bm{A} \bm{x}_k \nonumber \\
    &\quad + 2 \bm{x}_k' \bm{A}' \bm{P}_{k+1} \bm{B} \bm{u}_k 
    \nonumber \\
    &\quad + \bm{u}_k' \bm{B}' \bm{P}_{k+1} \bm{B} \bm{u}_k
    \label{eq:app_expand_quad}
\end{align}

Substituting into \eqref{eq:app_f_uk} and collecting terms:
\begin{equation}
    f(\bm{u}_k) = \bm{u}_k' \underbrace{(\bm{R} + \bm{B}' 
    \bm{P}_{k+1} \bm{B})}_{\bm{H}} \bm{u}_k + 
    2 \bm{x}_k' \bm{A}' \bm{P}_{k+1} \bm{B} \bm{u}_k + 
    \bm{x}_k' \bm{A}' \bm{P}_{k+1} \bm{A} \bm{x}_k
    \label{eq:app_f_collected}
\end{equation}

The matrix $\bm{H} = \bm{R} + \bm{B}' \bm{P}_{k+1} \bm{B}$ is 
positive definite since $\bm{R} \succ 0$ and 
$\bm{B}' \bm{P}_{k+1} \bm{B} \succeq 0$. This guarantees that 
$f(\bm{u}_k)$ is strictly convex and admits a unique minimum.

\paragraph{First-Order Condition.}
Differentiating $f(\bm{u}_k)$ with respect to $\bm{u}_k$ and 
setting the gradient to zero:
\begin{equation}
    \frac{\partial f}{\partial \bm{u}_k} = 
    2(\bm{R} + \bm{B}' \bm{P}_{k+1} \bm{B}) \bm{u}_k + 
    2 \bm{B}' \bm{P}_{k+1} \bm{A} \bm{x}_k = \bm{0}
    \label{eq:app_FOC}
\end{equation}

Solving for the optimal control $\bm{u}_k^*$:
\begin{equation}
    \boxed{\bm{u}_k^* = -\bm{K}_k \bm{x}_k}
    \label{eq:app_optimal_control}
\end{equation}
where the \textit{feedback gain matrix} is defined as:
\begin{equation}
    \boxed{\bm{K}_k = (\bm{R} + \bm{B}' \bm{P}_{k+1} \bm{B})^{-1} 
    \bm{B}' \bm{P}_{k+1} \bm{A}}
    \label{eq:app_gain}
\end{equation}

\paragraph{Second-Order Condition.}
The Hessian of $f(\bm{u}_k)$ is:
\begin{equation}
    \frac{\partial^2 f}{\partial \bm{u}_k \partial \bm{u}_k'} = 
    2(\bm{R} + \bm{B}' \bm{P}_{k+1} \bm{B}) \succ 0
    \label{eq:app_SOC}
\end{equation}
confirming that the solution \eqref{eq:app_optimal_control} is 
indeed a minimum.

\subsection{Derivation of the Riccati Equation}

Substituting the optimal control \eqref{eq:app_optimal_control} 
back into \eqref{eq:app_Jk_simplified}, we obtain the minimized 
cost-to-go. After algebraic manipulation, the result takes the 
quadratic form:
\begin{equation}
    J_k(\bm{x}_k) = \bm{x}_k' \bm{P}_k \bm{x}_k + c_k
    \label{eq:app_Jk_quadratic}
\end{equation}

To derive the recursion for $\bm{P}_k$, substitute 
$\bm{u}_k^* = -\bm{K}_k \bm{x}_k$ into the deterministic part of 
\eqref{eq:app_Jk_simplified}:
\begin{align}
    &\bm{u}_k^{*'} \bm{R} \bm{u}_k^* + 
    (\bm{A}\bm{x}_k + \bm{B}\bm{u}_k^*)' \bm{P}_{k+1} 
    (\bm{A}\bm{x}_k + \bm{B}\bm{u}_k^*) \nonumber \\
    &= \bm{x}_k' \bm{K}_k' \bm{R} \bm{K}_k \bm{x}_k + 
    \bm{x}_k' (\bm{A} - \bm{B}\bm{K}_k)' \bm{P}_{k+1} 
    (\bm{A} - \bm{B}\bm{K}_k) \bm{x}_k
    \label{eq:app_substitute_optimal}
\end{align}

Combining with the state cost $\bm{x}_k' \bm{Q}_k \bm{x}_k$:
\begin{equation}
    \bm{P}_k = \bm{Q}_k + \bm{K}_k' \bm{R} \bm{K}_k + 
    (\bm{A} - \bm{B}\bm{K}_k)' \bm{P}_{k+1} (\bm{A} - \bm{B}\bm{K}_k)
    \label{eq:app_Pk_form1}
\end{equation}

This expression can be simplified. Expanding and using the 
definition of $\bm{K}_k$:
\begin{align}
    \bm{P}_k &= \bm{Q}_k + \bm{A}' \bm{P}_{k+1} \bm{A} - 
    \bm{A}' \bm{P}_{k+1} \bm{B} \bm{K}_k - 
    \bm{K}_k' \bm{B}' \bm{P}_{k+1} \bm{A} + 
    \bm{K}_k' (\bm{R} + \bm{B}' \bm{P}_{k+1} \bm{B}) \bm{K}_k
    \label{eq:app_Pk_expand}
\end{align}

Using $\bm{K}_k = (\bm{R} + \bm{B}' \bm{P}_{k+1} \bm{B})^{-1} 
\bm{B}' \bm{P}_{k+1} \bm{A}$, we have:
\begin{equation}
    (\bm{R} + \bm{B}' \bm{P}_{k+1} \bm{B}) \bm{K}_k = 
    \bm{B}' \bm{P}_{k+1} \bm{A}
    \label{eq:app_K_identity}
\end{equation}

Substituting \eqref{eq:app_K_identity} into 
\eqref{eq:app_Pk_expand} and simplifying yields the 
\textit{Discrete-Time Riccati Equation}:
\begin{equation}
    \boxed{\bm{P}_k = \bm{A}' \bm{P}_{k+1} \bm{A} - 
    \bm{A}' \bm{P}_{k+1} \bm{B} 
    (\bm{R} + \bm{B}' \bm{P}_{k+1} \bm{B})^{-1} 
    \bm{B}' \bm{P}_{k+1} \bm{A} + \bm{Q}_k}
    \label{eq:app_riccati}
\end{equation}

Equivalently, using the gain matrix $\bm{K}_k$:
\begin{equation}
    \bm{P}_k = \bm{A}' \bm{P}_{k+1} \bm{A} - 
    \bm{A}' \bm{P}_{k+1} \bm{B} \bm{K}_k + \bm{Q}_k
    \label{eq:app_riccati_alt}
\end{equation}

The recursion for the constant term is:
\begin{equation}
    c_k = c_{k+1} + \text{tr}(\bm{P}_{k+1} \bm{\Sigma}_w)
    \label{eq:app_ck_recursion}
\end{equation}
with terminal condition $c_N = 0$.

\subsection{Algorithm Summary}

The complete backward-forward algorithm consists of two phases:

\paragraph{Phase 1: Backward Recursion (Offline).}
\begin{enumerate}
    \item Initialize: $\bm{P}_N = \bm{Q}_N$
    \item For $k = N-1, N-2, \ldots, 0$:
    \begin{enumerate}
        \item Compute the gain matrix:
        \begin{equation*}
            \bm{K}_k = (\bm{R} + \bm{B}' \bm{P}_{k+1} \bm{B})^{-1} 
            \bm{B}' \bm{P}_{k+1} \bm{A}
        \end{equation*}
        \item Update the Riccati matrix:
        \begin{equation*}
            \bm{P}_k = \bm{A}' \bm{P}_{k+1} \bm{A} - 
            \bm{A}' \bm{P}_{k+1} \bm{B} \bm{K}_k + \bm{Q}_k
        \end{equation*}
        \item Store $\bm{K}_k$
    \end{enumerate}
\end{enumerate}

\paragraph{Phase 2: Forward Simulation (Online).}
\begin{enumerate}
    \item Initialize: Observe initial state $\bm{x}_0$
    \item For $k = 0, 1, \ldots, N-1$:
    \begin{enumerate}
        \item Apply optimal control: $\bm{u}_k^* = -\bm{K}_k \bm{x}_k$
        \item System evolves: $\bm{x}_{k+1} = \bm{A}\bm{x}_k + 
        \bm{B}\bm{u}_k^* + \bm{w}_k$
        \item Observe $\bm{x}_{k+1}$
    \end{enumerate}
\end{enumerate}

\subsection{Certainty Equivalence Principle}

A fundamental property of the linear-quadratic-Gaussian (LQG) 
framework is the \textit{Certainty Equivalence Principle}. This 
states that:

\begin{enumerate}
    \item The optimal feedback gain $\bm{K}_k$ depends only on the 
    system matrices $(\bm{A}, \bm{B})$ and cost matrices 
    $(\bm{Q}_k, \bm{R})$, but \textbf{not} on the noise covariance 
    $\bm{\Sigma}_w$.
    
    \item The optimal control law is identical to that of the 
    deterministic problem (where $\bm{w}_k \equiv \bm{0}$).
    
    \item The stochastic disturbances affect only the 
    \textit{expected cost}, not the \textit{optimal policy}. 
    Specifically, the expected cost under optimal control is:
    \begin{equation}
        J^* = \bm{x}_0' \bm{P}_0 \bm{x}_0 + 
        \sum_{k=0}^{N-1} \text{tr}(\bm{P}_{k+1} \bm{\Sigma}_w)
        \label{eq:app_expected_cost}
    \end{equation}
    where the second term represents the irreducible cost of 
    uncertainty.
\end{enumerate}

\noindent The certainty equivalence property arises because: (i) the system 
dynamics are linear in both state and control, (ii) the cost 
function is quadratic, and (iii) the disturbances have zero mean. 
Under these conditions, the cross-terms between the deterministic 
and stochastic components vanish in expectation, allowing the 
optimization to proceed as if the future disturbances were known 
to equal their expected value of zero.

\vspace{5mm}

\noindent\textit{This derivation follows \citet[Chapter~3.1]{bertsekas2017dynamic}. 
For extensions to infinite-horizon problems, time-varying systems, 
and output feedback, see anderson2007optimal.}



\end{document}





